### 2015-04-17 Article.tex
first commit

\documentclass[conference]{IEEEtran}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Package
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ifpdf}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[french, english]{babel}
\usepackage{cite}
\usepackage{caption}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{array}

\usepackage{float}
\usepackage{multicol}
\usepackage[]{algorithm2e}

%
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Title
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Correlation among smart grid actors}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Authors
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Nicolas Gensollen, Monique Becker, Vincent Gauthier and Michel Marot  \\
\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
Telecom SudParis/Institut Mines Telecom\\
Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Abstract
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

\end{abstract}

\IEEEpeerreviewmaketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section I: Introduction
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}

Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences. Nevertheless, stability is a general concept and can be studied from several points of view. A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines. As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance, that enable scientists to get a better understanding of the structure of such systems. It appears that power grids are particular systems compared to well-known large graphs like the internet, the world wide web, or social networks. That is, they exhibit different characteritics, suggesting that they tend to behave differently. A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities. Therefore, meaningfull properties like power law degree distributions, prerential attachment, or hub oriented attacks do not apply with power grids.

Actually, considering only the underlying infrastructure when abstracting the power grid in a graph generally gives unrealistic results since they are mostly dynamical systems transporting electricity. As a second step, this dynamical component needs to be modeled using the physical laws governing the electricity behavior like the famous Kirchoff's laws. Besides, it is known that frequency synchronization in the grid is a necessary condition for stability. A nice solution abstracting this uses network models of coupled oscillators. For example, extended Kuramoto models using a swing equation based dynamcic provide interesting results.

Understanding power grids seems like a reasonable strarting point for studying so-called \textit{smart grids} that gained a lot of attention recently. As there exists many good articles presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point. In this paper, we rather concentrate on a different approach to the stability concept. We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads. In other words, an agent can produce and consume energy at any time. His production can be of course used to met his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production on a market. Such an agent model is known as a "\textit{prosumer}" (and will be called accordingly in this paper).

Since the number of prosumers in a system may be large it seems unrealistic to have a communication topology where a central agent coordinates all the others so they can inject power whenever they are over-producing. Furthermore, as the production component of these prosumers is supposed to come mainly from renewables (wind and solar power for instance), this over-producing state might be rather unstable. Just imagine how strong the impact of a simple cloud passing in the sky on the agent production will be. It has been suggested that energy diversification and geographical expansion are stabilizing factors for the production. Moreover, forming coalitions of agents in order to achieve a goal, a popular approach in game theory, could be of good help for solving both problems. The system operator would only have to communicate with a relatively small number of aggregators/coalitions, that in turn will use some internal communication protocol to manage the involved agents. Such coalitional and hierarchical concepts are often refered to as virtual power plants (VPP) in the smart grid litterature. The second point raised above is the central topic of the present paper : given N prosumers, what coalitions should be formed so that this over-production state has a high probability of being stable over the next period of interest ?

We will see that variability in the coalitions productions can be quantified to a certain extend by the correlation among the agents forming the coalitions. Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it will help us decide both what coalitions to form and how much they should sell. More precisely, we will build a framework in which the system operator has the possibility of fixing some entrance conditions on the market both in terms of stability and sufficient production. Based on predictions, coalitions can decide wether or not to enter and what power quantity they are willing to provide. Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or lying) will be exposed to financial penalities.

The paper is organized as follows, section 2 will give a brief overview of the related litterature, section 3 will clarify how we generated realistic prosumer production traces based on weather data. In section 4, we will define most of the notations and explain why correlation between prosumer is a quantity of interest for our objective. Based on the conclusions of section 4, section 5 shows how we approached the problem in a complex system fashion. Finally, section 6 will provide some results both on performance of the method and resilience of the coalitions formed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section III: Generating realistic prosumer patterns
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Generating realistic prosumer patterns}
\label{sec:data}

An essential component of the smart grid is the smart meter which makes the interface between the end user and the rest of the system. Smart meters coupled with sensors measure quantities of interest like instantaneous consumption, receive informations from the grid (electricity price for instance), and take actions accordingly (demand side management program). Smart meters are currently and gradually deployed, and will probably provide interesting datasets to work on. Unfortunately, at the time this paper was written, production and consumption data for prosumers over a large region were not yet available to our knowledge. Some interesting experiments are notwithstanding being conducted and data are progressively made public (see for instance the ISSDA experiment in Ireland).

We denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t. It can be expanded in $ P_{i}(t) = \bar{P}_{i}(t) - P_{i}^{D}(t) $, where $ \bar{P}_{i}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t. In other words, $ P_{i}(t) $ represents the instantaneous power that agent i is willing to sell at time t. As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ \bar{P}_{i} $ and $ P_{i}^{D} $.

For a prosumer i, it is possible to write both quantities as a sum over the distributed energy ressource ($ DER_{i} $) and loads ($ load_{i} $) of i : $ \bar{P}_{i}(t) = \sum_{k \in DER_{i}} P_{k}(t) $ and $ P_{i}^{D}(t) = \sum_{k \in load_{i}} P_{k}(t) $. For simplicity, in this paper we only consider windturbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):  $ \bar{P}_{i}(t) = \sum_{k \in WT_{i}} P_{k}(t) + \sum_{k \in PV_{i}} P_{k}(t) $.

We denote by $ \nu_{i}(t) $ and $ \xi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at agent i location and at time t, so that :
\begin{equation}
 \bar{P}_{i}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\xi_{i}(t) ) 
\end{equation}
Where $ \mathcal{F}_{WT} $ (resp. $ \mathcal{F}_{PV} $) is the power curve for the windturbines (resp. photovoltaic panels). We made here the implicit assumption that all windturbines (resp. photovoltaic panels) have the same power curve. More details about power curves and their approximations can be found in [].

Weather quantities like wind speed appear thus as alternative data for generating the $ P_{i} $ series. Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their granularity keeps increasing. The idea for generating the $ P_{i} $ series is pictured in the first block of the process diagram. Basically, the space is discretized into zones around the chosen weather stations. A simplifying assumption is that weather quantities are constant over their zone. That is, if prosumer i and j are positioned on the same zone, their are exposed to the same weather.

Note that a prosumer i is defined by his zone $ Z_{i} $ as well as the sets $ DER_{i} $ and $ load_{i} $. That is, a prosumer can be configured to represent anything from a single windturbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) throug more complex combinations. In practice, we use random configurations for the agents.

In the rest of the paper, we use french weather data (see www.infoclimat.fr) starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section II: Related Work
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{sec:related}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section III: Model
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Notations}
\label{sec:notations}

As explained in section 2, we have a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulated his extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012. Our objective is now to form groups of prosumers (the coalitions) so that the global power production resulting from the superposition of individual's extra-productions be sufficiently predictable. A parallel objectiv is to avoid very large coalitions as they would provide poor ameliorations for the communication load. Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the extra-production of coalition S at time t.

Suppose now that coalition S has to suggest a production value $ P^{CRCT} $ to enter the market. This means that, during the time S is on the market, it will have to inject in the grid exactly $ P^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalities if it deviates. Obviously, the actual extra-production will not be constant at this value and will oscillate due to various reasons. If S always produces more than $ P^{CRCT} $, it will never have to pay penalities, but it is losing some gains since it could have annonced a higher contract value. If the production oscillates around $ P^{CRCT} $, by using batteries or demand side management techniques, S could be able to maintain its production to the contract value at any time. Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalities. We can see that there is a return over risk tradeoff here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalities.

Let us illustrate the rest of the notations and concepts with a simple example. We consider only two agents i and j such that the distribution of their extra-production can be approximated by normal distributions : $ P_{i} \sim \mathcal{N}(\mu_{i}, \sigma_{i} ) $ and $ P_{j} \sim \mathcal{N}(\mu_{j}, \sigma_{j} ) $. This is only for explanation purposes as it is of course rather unrealistic in real situations where the distributions are skewed. Using simple statistics, we can write the distribution of the coalition $ \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :

\begin{equation}
\left\{ \begin{array}{lll}
		\mu_{ij} = \mu_{i} + \mu_{j} \\
		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij} \sigma_{i} \sigma_{j} }
\end{array} \right.
\end{equation}

$ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $. If the coalition $ \{i,j\}$ proposes a contract value $ P^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P^{CRCT} $ is critical. Indeed, in this situation, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid. The probability that $ \{i,j\}$ is underproducing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's credibility on the market. A well-known result for normal distributions is that the cumulative distribution function can be written as :
\begin{equation}
Pr[P_{ij} \leq P^{CRCT}] = \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P^{CRCT} - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] 
\end{equation}

The amount of risks a given coalition is willing to take depends on a lot of things, among wich its capacity to compensate for under-producing (using batteries, backup generators...). Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works. In order to keep the present paper in a reasonable length, we simplify the contract value selection problem a little bit by giving some responsabilities to a third party named the grid operator. The role of the grid operator is to constrain the market entry with an absolute low bound on the contract value and a upper limit relative to the coalition. In other words, the grid operator restrict the market to coalitions able to propose both sufficiently high and sufficiently credible contract values. More formally, let $ \phi \in [0,1] $ be a threshold fixed by the grid operator as a maximum value for the probability of under-producing. The highest contract value that a coalition can propose is thus $ P^{CRCT} $ such that $ Pr[P_{ij} \leq P^{CRCT}] = \phi $. In the gaussian example, this translates by coalition $ \{i,j\}$ announcing :
\begin{equation}
P^{CRCT} = \mu_{ij} + \sqrt{2} \sigma_{ij} erf^{-1}(2 \phi - 1 )
\end{equation}

This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator, and it is computable by any coalition. In order not to overload the market with unrealistically small coalitions, the grid operator also specify a lower bound $ P^{MIN} $ on the contract values. We thus characterized a valid coalition as one satisfying the two conditions :

\begin{equation}
\left\{ \begin{array}{lll}
			Pr[P_{ij} \leq P^{CRCT}] \leq \phi \\
			P^{CRCT} \geq P^{MIN}
\end{array} \right.
\end{equation}

The gaussian assumption of this small example is convenient as it allows us to write all conditions in closed forms. Nevertheless, such assumption is rather unrealistic in practice. In the following, we thus release this assumption and work with observed distributions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section V: Coalition Formation
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Coalition Formation}
\label{sec:forming}

Now that we defined the notions of contract values and valid coalitions, we can face the problem of forming the "right" coalitions given a pool of agents. In this paper, we consider the "right" coalitions as being the ones that maximize some utility notion that, basically, indicates how much stable power can be injected in the grid. This section aims at formalizing this utility notion, while explaining how this utility will be optimized in a greedy fashion over the correlation structure of the agents.

\subsection{Utility function}

We designed our model in a way that coalitions are remunerated proportionnally to their contract values $ \mathcal{U}(S) \propto P_{S}^{CRCT} $. That is, if $ \lambda $ is the unitary price rate for electricity, a coalition S injecting $ P_{S}^{CRCT} $ in the grid during a period $ [t_{0},t_{k}] $ earns :

\begin{equation}
\mathcal{U}(S ) = \int_{t_{0}}^{t_{k}} \lambda P_{S}^{CRCT} dt = P_{S}^{CRCT} \lambda \Delta_{t}
\end{equation}

Since $ \lambda $ appears, in this simple model, just a multiplicative constant, we consider for simplicity $ \lambda = 1 $ in the following. The utility of S is now the amount of energy S plans to inject during its contract period. Since we focused on power quantities from the begining of the present paper, and because the power injected by S in the grid is supposed to be constant over time, we simplify further the utility to its contract value : $ \mathcal{U}(S) = P_{S}^{CRCT} $. Although concise, this formulation suffers a major drawback. It is indeed not a concave function of the coalition length, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.

Such a model, that virtually allows infinitely large coalitions, is in practice not realistic. There are indeed costs (communication costs for instance) that increase with the coalitions sizes. We take this observation into account by rescaling the utility of a coalitions by its size in term of number of agents :

\begin{equation}
\mathcal{U}(S) = \left\{ \begin{array}{lll}
							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT} }{P_{S}^{MAX}},\ if\ S\ is\ valid, \\
							0,\ if\ S\ is\ not\ valid
						 \end{array}
				  \right.
\end{equation}

Where $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P_{S}^{MAX} $ is a normalizing factor. $ P_{S}^{MAX} $ is the maximum production possible, which only occurs when all consumptions are null and all generators produce their maximum power.

Based on $ \mathcal{U} $, the marginal contribution of an agent i can be expressed as $ \delta_{S}(i) = \mathcal{U}(S+\{i\}) - \mathcal{U}(S) $. A coalition S has thus an interest in adding an additional agent i if this marginal contribution is positive :

\begin{equation}
\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT} \geq P_{S}^{CRCT} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha} \dfrac{P_{S + \{i\}}^{MAX}}{P_{S}^{MAX}}
\end{equation}

Clearly, if $ \alpha = 0 $, whatever the size of the coalition, additional agents are added as long as they increase the contract value. Values of $\alpha $ greater than zero implies that additional agents must improve the contract value by a certain coefficient that decreases and converges to one as the size of the coalition is growing. For instance, if $\alpha = 1$ and $ |S| = 1 $, an additional agent must first double the contract value to be included, the next agent will then have to increase it by a factor of $ 1.5 $, and so on... Obviously, $ \alpha $ impacts directly the sizes of the coalitions formed. In the following, we consider the case where $ \alpha = 1 $.

As can be pointed out, the purpose of $ \mathcal{U} $ is not a study of coalitions stability against player defection, wich game theory provides a lot of tools for. This is indeed a problem on its own. The redistribution of a coalition's utility in terms of individual payoffs will therefore not be considered directly in this paper. $ \mathcal{U} $ can be rather interpreted as a measure of how good a given coalition is according to our criteria since it favors small coalitions with a good production to risk ratio.

\subsection{Representing the correlation structure}

The previous section explained that coalitions with small variances in their production probability distributions are more likely to afford high contracts. Furthermore, the gaussian example highlighted that the correlation structure of the agents plays an important role in finding stable coalitions. Usually, this correlation structure is formalized with a covariance matrix or a correlation matrix that contains all the correlation coefficients between the agents : $ M = (\rho_{ij})_{\forall i,j \in \mathcal{A}^{2}}$.

Since the work of Mantanega, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the timeseries in a correlation graph where the weight of an edge between two timeseries (nodes) is the metric value for these series. In the following, we use two opposite distance metrics :

\begin{equation}
\left\{ \begin{array}{lll}
			d_{ij}^{1} = 1 - \rho_{ij}^{2}, \\
			d_{ij}^{2} = \rho_{ij}^{2} = 1 - d_{ij}^{1}
\end{array} \right.
\end{equation}

Clearly, $ d^{1} $ (resp. $ d^{2} $) maps two correlated series as close points (resp. distant) while two uncorrelated series are distant (resp. close). These metrics enable us to compute a correlation graph $ G_{1} = (\mathcal{A}, E_{1}) $ and a "de-correlation" graph $ G_{2} = (\mathcal{A}, E_{2} ) $, where $ E_{1} $ is considered different from $ E_{2} $ since the weights are different.

Because the distance metrics can be computed for all pairs, these graphs are complete and of little use as is. Historically, the approach used by Mantanega was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering. Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph. Another filtering approach by mean of a threshold $ \epsilon $ on the edge weights solves this problem and allow one to exhibit clusters of correlated series.

Selecting the right filter $ \epsilon $ is an important point since it affects the correlation structure the coalition formation algorithm will be working on. Unfortunately, there seem to be no clear consensus in the litterature on how to select such a threshold. For our purpose, we have a number of coalitions $ N_{COAL} $ that we want to generate from cliques in the graph, we need therefore at least $ N_{COAL} $ cliques of a given size to start. Besides, we consider coalitions as disjoint sets in this section, meaning that the starting cliques must be non overlapping. We thus select the optimal threshold as :

\begin{equation}
\epsilon^{\star} = min_{ \epsilon \in [0,1]} \left\{ \epsilon\ s.t.\ |\Theta_{k}(G_{2}^{\epsilon})| \geq N_{COAL} \right\}
\end{equation} 
Where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G. In other words we select $ \epsilon $ as the smallest threshold posible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k. If the set is empty, it means that either $ N_{COAL} $ or k is to large.

Correlation can be seen as cosine of angles in $ L^{2} $, hence even if there is no strict transivity relation for correlation, there is, to a certain extent, some partial notion of it. More precisely, if a, b, and c are three items such that $ \rho_{ab} > \delta $ and $ \rho_{bc} > \delta $, then we know, by the cosine addition formula\footnote{ $ cos(a+b) = cos(a)cos(b) - sin(a)sin(b) $ }, that $ \rho_{ac} > 2 \delta^{2} - 1 $. That is, if a and b are strongly correlated ($\rho_{ab} > 0.9 $) and b and c are also strongly correlated ($\rho_{bc} > 0.9 $), then there is a high probability for a and c of being strongly correlated ($\rho_{ac} > 0.62 $). This is one of the reasons why searching for clusters in correlation graphs, that is, clustering according to the correlation, makes sense.

Nevertheless de-correlation seems like a more complex concept than correlation in the sense that there is not even a partial notion of transitivity when it comes to it. As expected, the clustering coefficients of $ G_{1} $ is much higher than the one of $ G_{2} $. This can be seen as another formulation of Onnela's study on the structural roles of weak and strong links on correlation graphs. Strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters. Searching for clusters in $ G_{2} $ and hoping that this strategy will provide a nice coalition structure of internally uncorrelated coalitions seems thus pointless.

Consider now a clique in  $ G_{2} $, which is a complete subgraph of $ G_{2} $. This is indeed a structure of interest for our purpose. Since there is a link for every pairs of nodes, we know, by construction, that a clique has a mean correlation and a maximum correlation less than $ \epsilon $. More formally, let

\begin{equation}
\left\{ \begin{array}{lll}
			\bar{\rho}_{S} = \dfrac{2}{|S|(|S|-1)} \sum_{i \in S} \sum_{j \in S, j>i} \rho_{ij}, \\ 
			\rho_{S}^{>} = MAX_{(i,j) \in S^{2}}(\rho_{ij})
\end{array} \right.			
\end{equation}

be the mean correlation of coalition S and the maximum correlation value in S. If S is a clique in $ G_{2} $ and $ \epsilon \in [0,1] $ is the threshold used for filtering $ G_{2} $, then $ \bar{\rho}_{S} \leq \dfrac{2}{|S|(|S|-1)} \sum_{i \in S} \sum_{j \in S, j>i} \epsilon $, that is, $ \bar{\rho}_{S} \leq \epsilon $, and $ \rho_{S}^{>} \leq \epsilon $.

Because of this de-correlation property, cliques in $ G_{2} $ appear as good candidates for coalitions. Nevertheless, by doing so, coalitions are often small and only formed based on the underlying correlation structure of the agents, and not on power generation capacity. It is possible that adding agents to these coalitions has the combined effect of increasing the production while decreasing its stability. The question revolves around measuring the benefits of this production surplus compared to the disadvantage of a higher unstability. Hopefully, this can be quantified exactly with the marginal benefit developped in the previous section.

\subsection{Coalition formation algorithm}

The algorithm takes inputs from :
\begin{itemize}
	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
\end{itemize} 
The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $. Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds. The next step is a local greedy improvement over the correlation structure represented by  $ G_{2}^{\epsilon^{\star}} $. Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique. This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing. The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing :

\begin{algorithm}
 \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
 \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
 Compute $ G_{2}^{\epsilon^{\star}} $ \;
 Find the $ N_{COAL} $ cliques in $ G_{2}^{\epsilon^{\star}} $\;
 \While{$ \mathcal{U}(CS) $ is improving}{
 	\For{each clique}{
 		Find $ i^{\star} $ \;
 		\If{ $ \delta_{clique}(i^{\star}) >= 0 $ }{
 			$ clique = clique \cup \{i^{\star} \} $ \;
 			}
   		}
  	}
 \caption{Algorithm}
\end{algorithm}

\section{Results}

The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities, and therefore, high contract values for relatively small sizes. Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the optimal coalitions set is very low. Actually, it is not obvious that a strict optimum exists at all. Besides, there is no, to our knowledge, state of the art algorithm that aggregate uncorrelated series in an optimum way.

In order to have an idea about the algorithm quality, we compare its results with :
\begin{itemize}
\item \textbf{Random :} This algorithm is a random search over the coalition structures space. It analyses a given number of structures and returns the best that it has encountered.
\item \textbf{Correlated :} This is the complete opposite of our algorithm. It basically uses the correlation graph $ G_{1} $ and performs community detection on it. The resulting coalitions have thus very high internal correlations. We thus expect this algorithm to perform very bad compared to the others.
\item \textbf{K-means :} This is simply a K-means clustering of the agents $ P_{i} $'s series. The goal of presenting this algorithm is to show that K-means do not consider correlations explicitely and is therefore not well suited for our purpose.
\end{itemize}

The coalitions formed according to these algorithms will be compared using two criteria : the normalized global utility of the coalitions formed $ \bar{\mathcal{U}}(CS) \in [0,1] $, and the percentage of valid coalitions that satisfy the grid conditions $ \tau(CS) \in [0,1] $ :
\begin{equation}
	\bar{\mathcal{U}}(CS) = \dfrac{1}{N_{COAL}} \sum_{S \in CS} \mathcal{U}(S)
\end{equation}
\begin{equation}
	\tau(CS) = \dfrac{|\{ S \in CS,\ s.t\ \mathcal{U}(S) > 0 \}|}{N_{COAL}}
\end{equation}

%Let $ P^{MARKET} = \sum_{S \in CS} P_{S}^{CRCT} $ be the production on the market, where CS is the coalition structure formed $ CS = \{S_{1},...S_{N_{COAL}} \} $, and $ N_{COAL} $ is the number of coalitions in the structure. First of all, $ P^{MARKET} $ depends explicitely on $ N_{COAL} $ such that, forming more coalitions induces an increase of the production. However, for a finite pool of agents, we expect an optimal number of coalitions $ N_{COAL}^{\star} $ that maximizes the market production. The reason for this expectation is that, for a given number of agents, more coalitions causes necessarily less agents in the coalitions, and therefore less diversity. We expect thus a point where coalitions formed start to fail the market entrance, and deteriorate the market production.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Section VI: Conclusion
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Bibliography
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{IEEEtran}  
\bibliography{Article}

\end{document}

### 2015-04-21 Article.tex
update

@@
 \usepackage[]{algorithm2e}
+\usepackage{hyperref}
 
@@
 A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines.
-As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance, that enable scientists to get a better understanding of the structure of such systems.
-It appears that power grids are particular systems compared to well-known large graphs like the internet, the world wide web, or social networks.
+As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance.
+These metrics enable scientists to get a better understanding of the structure of such complex systems.
+It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks.
 That is, they exhibit different characteritics, suggesting that they tend to behave differently.
 A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities.
-Therefore, meaningfull properties like power law degree distributions, prerential attachment, or hub oriented attacks do not apply with power grids.
+Therefore, meaningfull properties like power law degree distributions, prerential attachment, or hub oriented attacks do not seem to apply with power grids.
 
-Actually, considering only the underlying infrastructure when abstracting the power grid in a graph generally gives unrealistic results since they are mostly dynamical systems transporting electricity.
-As a second step, this dynamical component needs to be modeled using the physical laws governing the electricity behavior like the famous Kirchoff's laws.
-Besides, it is known that frequency synchronization in the grid is a necessary condition for stability.
-A nice solution abstracting this uses network models of coupled oscillators.
-For example, extended Kuramoto models using a swing equation based dynamcic provide interesting results.
+Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results.
+Power grids are ideed dynamical systems governed by physical laws.
+Concepts such as shortest path or centrality metrics have thus to be extended to take into account the particular nature of power grids.
+Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph.
+It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability.
+A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic.
+It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state.
 
@@
 As there exists many good articles presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
-In this paper, we rather concentrate on a different approach to the stability concept.
-We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads.
-In other words, an agent can produce and consume energy at any time.
-His production can be of course used to met his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production on a market.
+In this paper, we rather concentrate on a different approach to the system stability.
+We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads, such that an agent can produce and consume energy at any time.
+His production can be, of course, used to meet his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production to the grid.
 Such an agent model is known as a "\textit{prosumer}" (and will be called accordingly in this paper).
 
-Since the number of prosumers in a system may be large it seems unrealistic to have a communication topology where a central agent coordinates all the others so they can inject power whenever they are over-producing.
-Furthermore, as the production component of these prosumers is supposed to come mainly from renewables (wind and solar power for instance), this over-producing state might be rather unstable.
-Just imagine how strong the impact of a simple cloud passing in the sky on the agent production will be.
-It has been suggested that energy diversification and geographical expansion are stabilizing factors for the production.
-Moreover, forming coalitions of agents in order to achieve a goal, a popular approach in game theory, could be of good help for solving both problems.
-The system operator would only have to communicate with a relatively small number of aggregators/coalitions, that in turn will use some internal communication protocol to manage the involved agents.
-Such coalitional and hierarchical concepts are often refered to as virtual power plants (VPP) in the smart grid litterature.
-The second point raised above is the central topic of the present paper : given N prosumers, what coalitions should be formed so that this over-production state has a high probability of being stable over the next period of interest ?
+The constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
+A solution for organizing such a system consists in a market environement, where participating entities announce a production capacity for an upcoming period of time.
+These entities thus commit to injecting exactly at any time the contracted amount of power under financial penalities if they fail.
+Such a market system necessitates obviously some communication architecture to be viable.
+Since the number of prosumers in a system may be large, it seems unrealistic to have a flat centralized topology where a single agent coordinates all the others individually.
+Moreover, taking into account that the production component of these prosumers is supposed to come mainly from renewables, the over-producing state of the prosumers, and therefore the power injected in the grid, might be rather unstable.
+Imagine indeed how strong the impact of a simple cloud on a single agent production will be.
+
+It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production.
+Considering virtual coalitions of prosumers as potential providers, and allowing these aggregations to enter the market, could potentially solve these problems.
+Indeed, the system operator would only have to communicate with a relatively small number of aggregators, that in turn will use some internal communication protocol to manage the involved agents.
+Moreover, if the aggregation has been done properly, one expect a more stable and predictable energy production for the coalition than for a single agent.
+This diversification idea is ideed a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
 
@@
 Finally, section 6 will provide some results both on performance of the method and resilience of the coalitions formed.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section II: Related Work
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Related Work}
+\label{sec:related}
+
+The intermittent nature of renewables such as wind or solar power introduces new challenges in control design.
+Besides, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present.
+Unfortunately, these moments do not necessarily coincide with the consumption peak hours.
+A first approach to remedy this situation by acting on the consumption component is the use of dynamic ectricity prices sent to the end users.
+Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles.
+In periods where the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
+On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
+
+Another popular approach on the production side, is to combined renewable generators with storage devices, such that these are charged when there is a surplus of production, and discharged when the consumption exceeds the production.
+Although simple, this idea causes numerous challenges.
+Depending on the size of the system considered, centralized or decentralized control is desirable.
+In[], the authors introduce a distributed energy management system with a high penetration of renewables such that power is scheduled in a distributed fashion.
+In [] the optimal storage capacity problem is adressed.
+There is indeed an interesting tradeoff between the costs of the equipments and the expected availability of power.
+The authors develop a framework that enables them to exhibit a Pareto front of efficient solutions.
+
+Although there are great progresses in stabilizing the production by demand side managment or storage architectures, the need for good prediction techniques is more important than ever.
+Understanding the statistical properties of wind or solar irradiance are already well-established areas of research.
+There exists indeed a very interesting litterature on techniques aiming at predicting wind or solar irradiance profiles for future periods with error margins.
+In this paper, we adopt a slightly different approach : we suppose that the agents have historical records of their production and consumption, and we seek aggregations of agents with high expected production and low variability (wich is sometimes called risk in the present paper).
+
+Optimization of expected returns to risk is a traditional goal in finance.
+It is indeed well-known, that the more risk one is willing to take, the higher his potential gains.
+On the contrary, when investing exclusively on low rik assets, one should expect relatively small gains.
+This tradeoff is formalized in the Markowitz' portfolio theory.
+More precisely, given a set of assets for which we have some historic data of returns, the objective is to find a linear combination of these assets (the so-called portfolio) which maximizes the expected value while minimizing the variance of the portfolio's return.
+Markowitz's answer is a set of efficient portolios that all optimize in some sense this tradeoff.
+If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
+
+This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk as the variance of the production.
+However, one of the assumptions in the portfolio theory is that returns are normaly distributed random variables (or at least that the joint distribution is elliptic).
+This is a strong assumption which reveals unrealistic in most cases for our purposes when looking at real data.
+
+Nevertheless, one of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly (as we will see in section X) the variances of the portfolios.
+Since the work of Mantanega, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph where the weight of an edge between two timeseries (nodes) is the metric value for these series.
+
+Because the metric can be computed for all pairs, these graphs are complete and of little use as is.
+Historically, the approach used by Mantanega was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering.
+Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph.
+Another filtering approach by mean of a threshold $ \epsilon $ on the edge weights solves this problem and allows one to exhibit clusters of correlated series.
+In the following we will refer to these filtered graphs as $ \epsilon $-graph.
+
+An efficient portfolio of assets and a clustering of assets based on the correlations are obviously two very distinct things.
+And none of them is exactly what we seek.
+We will see in section X, how we combined both theories in order to form sufficiently producing coalitions with an acceptable risk level.
 
@@
 We denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t.
-It can be expanded in $ P_{i}(t) = \bar{P}_{i}(t) - P_{i}^{D}(t) $, where $ \bar{P}_{i}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
-In other words, $ P_{i}(t) $ represents the instantaneous power that agent i is willing to sell at time t.
-As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ \bar{P}_{i} $ and $ P_{i}^{D} $.
+That is, $ P_{i}(t) = P_{i}^{P}(t) - P_{i}^{D}(t) $, where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
+In other words, $ P_{i}(t) $ represents the instantaneous surplus of power that agent i is willing to sell at time t.
+As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ P_{i}^{P} $ and $ P_{i}^{D} $.
 
-For a prosumer i, it is possible to write both quantities as a sum over the distributed energy ressource ($ DER_{i} $) and loads ($ load_{i} $) of i : $ \bar{P}_{i}(t) = \sum_{k \in DER_{i}} P_{k}(t) $ and $ P_{i}^{D}(t) = \sum_{k \in load_{i}} P_{k}(t) $.
-For simplicity, in this paper we only consider windturbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):  $ \bar{P}_{i}(t) = \sum_{k \in WT_{i}} P_{k}(t) + \sum_{k \in PV_{i}} P_{k}(t) $.
+For a prosumer i, it is possible to write both quantities as a sum over the distributed energy ressources ($ DER_{i} $) and loads ($ load_{i} $) of i :
 
-We denote by $ \nu_{i}(t) $ and $ \xi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at agent i location and at time t, so that :
 \begin{equation}
- \bar{P}_{i}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\xi_{i}(t) ) 
+P_{i}^{P}(t) = \sum_{k \in DER_{i}} P_{k}(t)
+\end{equation}
+\begin{equation}
+P_{i}^{D}(t) = \sum_{k \in load_{i}} P_{k}(t)
+\end{equation}
+
+For simplicity, in this paper we only consider windturbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
+
+\begin{equation}
+P_{i}^{D}(t) = \sum_{k \in WT_{i}} P_{k}(t) + \sum_{k \in PV_{i}} P_{k}(t)
+\end{equation}
+
+We denote by $ \nu_{i}(t) $ and $ \xi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
+\begin{equation}
+ P_{i}^{P}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\xi_{i}(t) ) 
 \end{equation}
@@
 photovoltaic panels) have the same power curve.
+The model can be easily extended to muplitiple power curves accounting for different types of generators.
 More details about power curves and their approximations can be found in [].
 
-Weather quantities like wind speed appear thus as alternative data for generating the $ P_{i} $ series.
-Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their granularity keeps increasing.
-The idea for generating the $ P_{i} $ series is pictured in the first block of the process diagram.
-Basically, the space is discretized into zones around the chosen weather stations.
-A simplifying assumption is that weather quantities are constant over their zone.
-That is, if prosumer i and j are positioned on the same zone, their are exposed to the same weather.
+Weather quantities like wind speed or solar irradiance appear thus as alternative data for generating the $ P_{i} $ series.
+Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
+A quantity like wind speed depends indeed both on time and location.
+In the following, time is discretized into slots and space into zones.
+A zone is simply a portion of the considered space for which we have weather data.
+Therefore, if prosumers i and j are positioned on the same zone, they are exposed to the same weather.
+Adding some noise can easily be done though not considered in this paper.
+The process for generating the $ P_{i} $ series is pictured in the first block of the process diagram (see figure \ref{fig:process}).
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section II: Related Work
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Related Work}
-\label{sec:related}
+\begin{figure}
+\includegraphics[scale=.45]{Fig2.pdf}
+\caption{Process diagram}
+\label{fig:process}
+\end{figure}
 
@@
 
-As explained in section 2, we have a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulated his extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
-Our objective is now to form groups of prosumers (the coalitions) so that the global power production resulting from the superposition of individual's extra-productions be sufficiently predictable.
-A parallel objectiv is to avoid very large coalitions as they would provide poor ameliorations for the communication load.
+As explained in section \ref{sec:data}, we have a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulated his extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
+Based on these historical values, our objective is now to form groups of prosumers (the so-called coalitions) so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
 Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the extra-production of coalition S at time t.
 
-Suppose now that coalition S has to suggest a production value $ P^{CRCT} $ to enter the market.
-This means that, during the time S is on the market, it will have to inject in the grid exactly $ P^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalities if it deviates.
-Obviously, the actual extra-production will not be constant at this value and will oscillate due to various reasons.
-If S always produces more than $ P^{CRCT} $, it will never have to pay penalities, but it is losing some gains since it could have annonced a higher contract value.
-If the production oscillates around $ P^{CRCT} $, by using batteries or demand side management techniques, S could be able to maintain its production to the contract value at any time.
+Suppose now that coalition S has to suggest a production value $ P_{S}^{CRCT} $ to enter the market.
+This means that, during the time S is on the market, it will have to inject in the grid exactly $ P_{S}^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalities if it deviates.
+Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittencies in the production and consumption.
+If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalities, but it is losing some gains since it could have annonced a higher contract value.
+If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques (see section \ref{sec:related}), S could be able to maintain its production to the contract value at any time.
 Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalities.
@@
 $ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
-If the coalition $ \{i,j\}$ proposes a contract value $ P^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P^{CRCT} $ is critical.
+If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
 Indeed, in this situation, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid.
@@
 More formally, let $ \phi \in [0,1] $ be a threshold fixed by the grid operator as a maximum value for the probability of under-producing.
-The highest contract value that a coalition can propose is thus $ P^{CRCT} $ such that $ Pr[P_{ij} \leq P^{CRCT}] = \phi $.
+The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
 In the gaussian example, this translates by coalition $ \{i,j\}$ announcing :
 \begin{equation}
-P^{CRCT} = \mu_{ij} + \sqrt{2} \sigma_{ij} erf^{-1}(2 \phi - 1 )
+P_{S}^{CRCT \star} = \mu_{ij} + \sqrt{2} \sigma_{ij} erf^{-1}(2 \phi - 1 )
 \end{equation}
@@
 
-The gaussian assumption of this small example is convenient as it allows us to write all conditions in closed forms.
+The gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
 Nevertheless, such assumption is rather unrealistic in practice.
@@
 
-Since the work of Mantanega, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the timeseries in a correlation graph where the weight of an edge between two timeseries (nodes) is the metric value for these series.
 In the following, we use two opposite distance metrics :
@@
 close).
-These metrics enable us to compute a correlation graph $ G_{1} = (\mathcal{A}, E_{1}) $ and a "de-correlation" graph $ G_{2} = (\mathcal{A}, E_{2} ) $, where $ E_{1} $ is considered different from $ E_{2} $ since the weights are different.
+These metrics enable us to compute a correlation graph $ G_{1} = (\mathcal{A}, E_{1}) $ and a "de-correlation" graph $ G_{2} = (\mathcal{A}, E_{2} ) $.
+For any i and j, the weight of the edge $ e_{ij} $ is $ d_{ij}^{1} $ in $ G_{1} $ and $ d_{ij}^{2} $ in $ G_{2} $.
 
-Because the distance metrics can be computed for all pairs, these graphs are complete and of little use as is.
-Historically, the approach used by Mantanega was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering.
-Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph.
-Another filtering approach by mean of a threshold $ \epsilon $ on the edge weights solves this problem and allow one to exhibit clusters of correlated series.
-
-Selecting the right filter $ \epsilon $ is an important point since it affects the correlation structure the coalition formation algorithm will be working on.
-Unfortunately, there seem to be no clear consensus in the litterature on how to select such a threshold.
-For our purpose, we have a number of coalitions $ N_{COAL} $ that we want to generate from cliques in the graph, we need therefore at least $ N_{COAL} $ cliques of a given size to start.
-Besides, we consider coalitions as disjoint sets in this section, meaning that the starting cliques must be non overlapping.
+Selecting the right filter $ \epsilon $ is an important point since it affects the correlation structure which affects in turn the formation of the coalitions.
+Unfortunately, there seems to be no clear consensus in the litterature on how to select such a threshold.
+In our situation, we want to generate $ N_{COAL} $ coalitions from cliques in the graph.
+We need therefore at least $ N_{COAL} $ cliques of a given size to start.
+Besides, since we consider coalitions as disjoint sets, the starting cliques should be non overlapping.
 We thus select the optimal threshold as :
@@
 \begin{equation}
+\label{epsilon_star}
 \epsilon^{\star} = min_{ \epsilon \in [0,1]} \left\{ \epsilon\ s.t.\ |\Theta_{k}(G_{2}^{\epsilon})| \geq N_{COAL} \right\}
 \end{equation}
+
 Where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
 In other words we select $ \epsilon $ as the smallest threshold posible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
-If the set is empty, it means that either $ N_{COAL} $ or k is to large.
+The existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
+The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
 
@@
 This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
-The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing :
+The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
+See the details in algorithm.
 
 \begin{algorithm}
+\label{alg:algo1}
  \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
@@
  		\If{ $ \delta_{clique}(i^{\star}) >= 0 $ }{
- 			$ clique = clique \cup \{i^{\star} \} $ \;
+ 			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
  			}
@@
   	}
- \caption{Algorithm}
+ \caption{Percolation algorithm}
+\end{algorithm}
+
+\begin{algorithm}
+\label{alg:algo2}
+ \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ maxiter $}
+ \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
+ $ curriter \leftarrow 0 $ \;
+ $ CS^{\star} \leftarrow \emptyset $\;
+ \While{$ curriter < maxiter $}{
+ 	$ CS \leftarrow SelectRandomCS() $\;
+ 	\If{ $ \mathcal{U}(CS) > \mathcal{U}(CS^{\star}) $ }{
+ 		$ CS^{\star } \leftarrow CS $\;
+ 		}
+ 	$ curriter \leftarrow  curriter + 1 $\;
+  	}
+  return $ CS^{\star} $
+ \caption{Random algorithm}
+\end{algorithm}
+
+\begin{algorithm}
+\label{alg:algo3}
+ \KwData{$P_{i}$ series,\\ Desired number of coalitions $ N_{COAL} $,\\ search step size $ \beta << 1 $}
+ \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
+ $ \epsilon \leftarrow 1 $ \;
+ $ CS \leftarrow \emptyset $\;
+ \While{$ |CS| < N_{COAL} $}{
+ 	Compute $ G_{1}^{\epsilon} $ \;
+ 	$ CS \leftarrow computeClusters( G_{1}^{\epsilon} ) $\;
+ 	\If{ $ |CS| = N_{COAL} $ }{
+ 		return CS\;
+ 		}
+ 	\Else{ 
+ 	$ \epsilon \leftarrow \epsilon - \beta $\;
+ 	}
+  	}
+ \caption{Correlated algorithm}
 \end{algorithm}
@@
 
-In order to have an idea about the algorithm quality, we compare its results with :
+In order to have an idea about the algorithm's quality, we compare its results with :
 \begin{itemize}
-\item \textbf{Random :} This algorithm is a random search over the coalition structures space. It analyses a given number of structures and returns the best that it has encountered.
+\item \textbf{Random :} This algorithm is basically a random search over the coalition structures space.
+It analyses a given number of structures and returns the best that it has encountered so far.
+See algorithm.
 \item \textbf{Correlated :} This is the complete opposite of our algorithm.
@@
 We thus expect this algorithm to perform very bad compared to the others.
+See algorithm.
 \item \textbf{K-means :} This is simply a K-means clustering of the agents $ P_{i} $'s series.

### 2015-04-22 Article.tex
update

@@
 In the gaussian example, this translates by coalition $ \{i,j\}$ announcing :
+
 \begin{equation}
+\label{eq:P_star}
 P_{S}^{CRCT \star} = \mu_{ij} + \sqrt{2} \sigma_{ij} erf^{-1}(2 \phi - 1 )
@@
 
+\begin{figure}
+\includegraphics[scale=.6]{gaussian_P_star.pdf}
+\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for gaussian distributions (see equation \ref{eq:P_star}).
+Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5.
+Under a grid policy of $ \phi = 0.1 $, it is able to announce a contract value of $ P_{S}^{CRCT \star} = 4.36 $.
+The same coalition in term of expected production ($\mu = 5 $), but with a higher variance ($ \sigma = 1.5 $, cyan curve with circles) can only afford a smaller contract value of $ P_{S}^{CRCT \star} = 3.07 $.
+The red curve with diamonds stands for a coalition with a higher expected production $(\mu = 7)$, but with a very high unpredictability ($ \sigma = 5 $).
+For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units.
+Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line)	.
+}
+\label{fig:Gaussian}
+\end{figure}
+
 This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator, and it is computable by any coalition.
+Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $.
+For illustration, the range of $ \phi $ values is shown from 0 to 1, but in practice, only small values of $ \phi $ really make sense : $ \phi = 1 $ for instance means that coalitions can announce absolutely anything since the probability of producing less than any contract value is necessarily less than one by trivial definition of a probability.
+As visible on figure \ref{fig:Gaussian}, coalitions with high expected production but presenting a high unpredictability are penalized and can only afford small contracts.
+
 In order not to overload the market with unrealistically small coalitions, the grid operator also specify a lower bound $ P^{MIN} $ on the contract values.
@@
 
+On figure \ref{fig:Gaussian}, $ P^{MIN} $ is fixed to 2 units for illustration purpose.
+For $ \phi = 0.1 $, only blue triangles and cyan circles coalitions are valid while red diamonds coalition is not.
+
 The gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
 Nevertheless, such assumption is rather unrealistic in practice.
-In the following, we thus release this assumption and work with observed distributions.
+In the following, we keep the same framework but release this gaussian assumption.
+We therefore work with any observed distributions drawn from the data.
 
@@
 There are indeed costs (communication costs for instance) that increase with the coalitions sizes.
-We take this observation into account by rescaling the utility of a coalitions by its size in term of number of agents :
+We take this observation into account by rescaling the utility of a coalition by its size in term of number of agents :
 
@@
 
+Clearly, as $ \alpha $ impacts the sizes of the coalitions, it should not be selected randomly.
+In order to have an initial guess, we use a mean field approximation in a gaussian case (as in section \ref{sec:notations}).
+In this section, $ \bar{x} $ denote the mean approximation of quantity x.
+Recall that we have N agents and we wish to form $ N_{COAL} $ coalitions.
+We denote by $ \bar{N} = \lfloor \dfrac{N}{N_{COAL}} \rfloor $ the expected mean number of agents in a coalition.
+If $ \mathcal{U}(S) $ favors coalitions of size $ \bar{N} $, then :
+
+\begin{equation}
+\left[ \dfrac{\partial{ U}}{ \partial{|S|}} \right]_{|S| = \bar{N}} = 0
+\end{equation}
+
+Solving this equation for $ \alpha $ in the mean field approximation for the gaussian case yields :
+\footnotesize
+\begin{equation}
+\alpha^{\star}_{\bar{N}} = \dfrac{0.7 \bar{\sigma}(\bar{\rho}-1)erf^{-1}(2 \phi - 1)}{\bar{\mu}\sqrt{\bar{N}(\bar{\rho}\bar{N}-\bar{\rho}+1)}+1.4 \bar{\sigma} erf^{-1} (2 \phi -1 ) (\bar{\rho}\bar{N}-\bar{\rho} + 1)}
+\end{equation}
+\normalsize
 Clearly, if $ \alpha = 0 $, whatever the size of the coalition, additional agents are added as long as they increase the contract value.

### 2015-04-23 Article.tex
update

@@
 Nevertheless, such assumption is rather unrealistic in practice.
-In the following, we keep the same framework but release this gaussian assumption.
-We therefore work with any observed distributions drawn from the data.
+In the following, we keep the same framework but release this gaussian assumption unless the contrary is specified.
+This assumption will indeed be convenient for computing some parameter estimates.
 
@@
 
-Where $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P_{S}^{MAX} $ is a normalizing factor.
+Where parameter $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P_{S}^{MAX} $ is a normalizing factor.
 $ P_{S}^{MAX} $ is the maximum production possible, which only occurs when all consumptions are null and all generators produce their maximum power.
@@
 \delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT} \geq P_{S}^{CRCT} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha} \dfrac{P_{S + \{i\}}^{MAX}}{P_{S}^{MAX}}
+\label{eq:marginal_benefit}
 \end{equation}
 
-Clearly, as $ \alpha $ impacts the sizes of the coalitions, it should not be selected randomly.
-In order to have an initial guess, we use a mean field approximation in a gaussian case (as in section \ref{sec:notations}).
-In this section, $ \bar{x} $ denote the mean approximation of quantity x.
-Recall that we have N agents and we wish to form $ N_{COAL} $ coalitions.
-We denote by $ \bar{N} = \lfloor \dfrac{N}{N_{COAL}} \rfloor $ the expected mean number of agents in a coalition.
-If $ \mathcal{U}(S) $ favors coalitions of size $ \bar{N} $, then :
+In order to understand how this utility function behaves and how it relates to the $ \alpha $ parameter, we use again the gaussian case (as in section \ref{sec:notations}).
+Since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
+Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
+In the following $ \bar{N} $ denotes the mean number of agents in a coalition.
+If we wish that $ \mathcal{U}(S) $ tends to form coalitions of size approximately of the order of $ \bar{N} $, then :
 
@@
 
-Solving this equation for $ \alpha $ in the mean field approximation for the gaussian case yields :
+Solving this equation for $ \alpha $ in the mean approximation for the gaussian case yields :
 \footnotesize
@@
 \alpha^{\star}_{\bar{N}} = \dfrac{0.7 \bar{\sigma}(\bar{\rho}-1)erf^{-1}(2 \phi - 1)}{\bar{\mu}\sqrt{\bar{N}(\bar{\rho}\bar{N}-\bar{\rho}+1)}+1.4 \bar{\sigma} erf^{-1} (2 \phi -1 ) (\bar{\rho}\bar{N}-\bar{\rho} + 1)}
+\label{eq:alpha_star}
 \end{equation}
 \normalsize
-Clearly, if $ \alpha = 0 $, whatever the size of the coalition, additional agents are added as long as they increase the contract value.
+
+Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $.
+These nice curves are of course only valid in the simplified example considered here, but they will provide some guidance when using real data.
+
+\begin{figure}
+\includegraphics[scale=.6]{mean_field.pdf}
+\caption{Gaussian mean approximation.
+Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}).
+Subplot \textbf{b} dislays the corresponding utility functions for different values of $ \alpha $.
+Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
+Finally, the red curve with squares has an optimum size of 15 agents.
+}
+\label{fig:mean_approx}
+\end{figure}
+
+%Clearly, if $ \alpha = 0 $, whatever the size of the coalition, additional agents are added as long as they increase the contract value.
 Values of $\alpha $ greater than zero implies that additional agents must improve the contract value by a certain coefficient that decreases and converges to one as the size of the coalition is growing.
@@
 The previous section explained that coalitions with small variances in their production probability distributions are more likely to afford high contracts.
-Furthermore, the gaussian example highlighted that the correlation structure of the agents plays an important role in finding stable coalitions.
+Furthermore, the gaussian approximation highlighted that the correlation structure of the agents plays an important role in finding stable coalitions.
 Usually, this correlation structure is formalized with a covariance matrix or a correlation matrix that contains all the correlation coefficients between the agents : $ M = (\rho_{ij})_{\forall i,j \in \mathcal{A}^{2}}$.
@@
 be the mean correlation of coalition S and the maximum correlation value in S.
-If S is a clique in $ G_{2} $ and $ \epsilon \in [0,1] $ is the threshold used for filtering $ G_{2} $, then $ \bar{\rho}_{S} \leq \dfrac{2}{|S|(|S|-1)} \sum_{i \in S} \sum_{j \in S, j>i} \epsilon $, that is, $ \bar{\rho}_{S} \leq \epsilon $, and $ \rho_{S}^{>} \leq \epsilon $.
+If S is a clique in $ G_{2} $ and $ \epsilon \in [0,1] $ is the threshold used for filtering $ G_{2} $, then $ \bar{\rho}_{S} \leq 2/(|S|(|S|-1)) \sum_{i \in S} \sum_{j \in S, j>i} \epsilon $, that is, $ \bar{\rho}_{S} \leq \epsilon $, and $ \rho_{S}^{>} \leq \epsilon $.
 
-Because of this de-correlation property, cliques in $ G_{2} $ appear as good candidates for coalitions.
-Nevertheless, by doing so, coalitions are often small and only formed based on the underlying correlation structure of the agents, and not on power generation capacity.
-It is possible that adding agents to these coalitions has the combined effect of increasing the production while decreasing its stability.
+Figure \ref{fig:histo_cliques} shows the distributions of the utility values for cliques of size 3 (triangles) in $ G_{2}^{\epsilon^{\star}} $ and for all the other possible triplets of agents.
+It is clearly visible that cliques tend to exhibit higher utilities because of their de-correlation property.
+Choosing cliques in $ G_{2}^{\epsilon^{\star}} $ as coalitions seems therefore appealing.
+Nevertheless, the quality of the results seems to decrease as the sizes of the cliques increase.
+Indeed, the larger the desired cliques, the more dense $ G_{2}^{\epsilon^{\star}} $ becomes (see equation \ref{epsilon_star}).
+There is a point where cliques results more from noisy edges than true de-correlation relationships, which decreases the quality of the results.
+
+Directly mapping cliques to coalitions by this de-correlation oriented approach is thus not sufficient.
+It is indeed possible that adding agents to these cliques has the combined effect of increasing the expected production while decreasing its stability.
 The question revolves around measuring the benefits of this production surplus compared to the disadvantage of a higher unstability.
-Hopefully, this can be quantified exactly with the marginal benefit developped in the previous section.
+This can be quantified by the marginal benefit of equation \ref{eq:marginal_benefit}.
+
+\begin{figure}
+\includegraphics[scale=.6]{cliques_histo.pdf}
+\caption{Histograms of utility values for coalitions of size 3.
+Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets.
+As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
+\label{fig:histo_cliques}
+\end{figure}
 
@@
 The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
-See the details in algorithm.
+See the details in algorithm \ref{alg:algo1}.
 
 \begin{algorithm}
-\label{alg:algo1}
  \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
@@
  \caption{Percolation algorithm}
+ \label{alg:algo1}
 \end{algorithm}
@@
 \begin{algorithm}
-\label{alg:algo2}
  \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ maxiter $}
@@
  \caption{Random algorithm}
+ \label{alg:algo2}
 \end{algorithm}
@@
 \begin{algorithm}
-\label{alg:algo3}
  \KwData{$P_{i}$ series,\\ Desired number of coalitions $ N_{COAL} $,\\ search step size $ \beta << 1 $}
@@
  \caption{Correlated algorithm}
+\label{alg:algo3}
 \end{algorithm}
@@
 It analyses a given number of structures and returns the best that it has encountered so far.
-See algorithm.
+See algorithm \ref{alg:algo2}.
 \item \textbf{Correlated :} This is the complete opposite of our algorithm.
@@
 We thus expect this algorithm to perform very bad compared to the others.
-See algorithm.
+See algorithm \ref{alg:algo3}.
 \item \textbf{K-means :} This is simply a K-means clustering of the agents $ P_{i} $'s series.
@@
 \end{equation}
+
+Before running the algorithm, we need to calibrate the utility function by choosing the $ \alpha $ parameter.
+In this paper, neither the communication network nor the electrical grid are explicitely considered.
+Thus, we do not have any technical constraints on coalition sizes even if we designed the utitility such that these kind of constraints could be taken into account.
+We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
+Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
+Using equation \ref{eq:alpha_star} to select $ \alpha $ based on the mean quantities seems to give acceptable results for the utility function.
+
+\begin{figure}
+\includegraphics[scale=.6]{real_utility.pdf}
+\caption{Utility of random coalitions depending on their size.
+Lines shows mean utility values and colored areas stands for standard deviations.
+The slim blue noisy curve shows the real mean utility, and the thick red curve is its smoothed version by applying a Savitzky-Golay filter.
+On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 10 agents coalitions.}
+\label{fig:real_utility}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=.6]{real_utility_2.pdf}
+\caption{Utility of random coalitions depending on their size.
+Blue dots shows real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
+On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
+\label{fig:real_utility2}
+\end{figure}
 

### 2015-04-29 Article.tex
update

@@
 Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
-Nevertheless, stability is a general concept and can be studied from several points of view.
 A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines.
@@
 It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks.
-That is, they exhibit different characteritics, suggesting that they tend to behave differently.
+That is, they exhibit different characteritics, suggesting that they may behave differently.
 A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities.
-Therefore, meaningfull properties like power law degree distributions, prerential attachment, or hub oriented attacks do not seem to apply with power grids.
+Therefore, properties like power law degree distributions, prerential attachment, or hub oriented attacks do not necessarily apply with power grids.
 
@@
 The constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
-A solution for organizing such a system consists in a market environement, where participating entities announce a production capacity for an upcoming period of time.
+An envisaged solution for organizing such a system consists in a market environement, where participating entities announce a production capacity for an upcoming period of time.
 These entities thus commit to injecting exactly at any time the contracted amount of power under financial penalities if they fail.
@@
 Since the number of prosumers in a system may be large, it seems unrealistic to have a flat centralized topology where a single agent coordinates all the others individually.
-Moreover, taking into account that the production component of these prosumers is supposed to come mainly from renewables, the over-producing state of the prosumers, and therefore the power injected in the grid, might be rather unstable.
-Imagine indeed how strong the impact of a simple cloud on a single agent production will be.
+Some articles suggest that a fully decentralized architecture will probably be too complicated to achieve, and that some kind of hierarchical hybrid topology seems more likely to be viable.
+
+Moreover, taking into account that the production component of a prosumer is supposed to come mainly from renewables, its over-producing state, and therefore the power he injects in the grid, might be rather unstable.
 
 It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production.
-Considering virtual coalitions of prosumers as potential providers, and allowing these aggregations to enter the market, could potentially solve these problems.
+Considering virtual coalitions of prosumers as potential providers, and allowing these aggregations to enter the market, could potentially solve some of these problems.
 Indeed, the system operator would only have to communicate with a relatively small number of aggregators, that in turn will use some internal communication protocol to manage the involved agents.
-Moreover, if the aggregation has been done properly, one expect a more stable and predictable energy production for the coalition than for a single agent.
+Moreover, if the aggregation has been done properly, one should expect a more stable and predictable energy production for the coalition than for a single agent.
 This diversification idea is ideed a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
@@
 
-This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk as the variance of the production.
+This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk / volatility as the variance of the production.
 However, one of the assumptions in the portfolio theory is that returns are normaly distributed random variables (or at least that the joint distribution is elliptic).
@@
 And none of them is exactly what we seek.
-We will see in section X, how we combined both theories in order to form sufficiently producing coalitions with an acceptable risk level.
+We will see in section X, how we combined both theories in order to form sufficiently producing coalitions with an acceptable volatility level.
 
@@
 An essential component of the smart grid is the smart meter which makes the interface between the end user and the rest of the system.
-Smart meters coupled with sensors measure quantities of interest like instantaneous consumption, receive informations from the grid (electricity price for instance), and take actions accordingly (demand side management program).
+Smart meters coupled with sensors measure quantities of interest (like instantaneous consumption), receive informations from the grid (electricity price for instance), and take actions accordingly (demand side management program).
 Smart meters are currently and gradually deployed, and will probably provide interesting datasets to work on.
@@
 
-We denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t.
+In the following, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t.
 That is, $ P_{i}(t) = P_{i}^{P}(t) - P_{i}^{D}(t) $, where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
@@
 Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
-A quantity like wind speed depends indeed both on time and location.
+A quantity like wind speed depends both on time and location.
 In the following, time is discretized into slots and space into zones.
@@
 Indeed, in this situation, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid.
-The probability that $ \{i,j\}$ is underproducing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's credibility on the market.
+The probability that $ \{i,j\}$ is underproducing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's quality.
 A well-known result for normal distributions is that the cumulative distribution function can be written as :
@@
 
+Where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
+
 The amount of risks a given coalition is willing to take depends on a lot of things, among wich its capacity to compensate for under-producing (using batteries, backup generators...).
@@
 In order to keep the present paper in a reasonable length, we simplify the contract value selection problem a little bit by giving some responsabilities to a third party named the grid operator.
-The role of the grid operator is to constrain the market entry with an absolute low bound on the contract value and a upper limit relative to the coalition.
-In other words, the grid operator restrict the market to coalitions able to propose both sufficiently high and sufficiently credible contract values.
+The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
 More formally, let $ \phi \in [0,1] $ be a threshold fixed by the grid operator as a maximum value for the probability of under-producing.
@@
 For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units.
-Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line)	.
+Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line).
 }
@@
 
-This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator, and it is computable by any coalition.
+This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator.
 Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $.
@@
 \left\{ \begin{array}{lll}
-			Pr[P_{ij} \leq P^{CRCT}] \leq \phi \\
-			P^{CRCT} \geq P^{MIN}
+			Pr[P_{ij} \leq P_{S}^{CRCT}] \leq \phi \\
+			P_{S}^{CRCT} \geq P^{MIN}
 \end{array} \right.
@@
 Although concise, this formulation suffers a major drawback.
-It is indeed not a concave function of the coalition length, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.
+It is indeed not a concave function, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.
 
@@
 There are indeed costs (communication costs for instance) that increase with the coalitions sizes.
-We take this observation into account by rescaling the utility of a coalition by its size in term of number of agents :
+We take this observation into account by rescaling the utility of a coalition S by its size in term of number of agents ($|S|$):
 
@@
 \mathcal{U}(S) = \left\{ \begin{array}{lll}
-							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT} }{P_{S}^{MAX}},\ if\ S\ is\ valid, \\
+							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT} }{P^{MAX}},\ if\ S\ is\ valid, \\
 							0,\ if\ S\ is\ not\ valid
@@
 
-Where parameter $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P_{S}^{MAX} $ is a normalizing factor.
-$ P_{S}^{MAX} $ is the maximum production possible, which only occurs when all consumptions are null and all generators produce their maximum power.
+Where parameter $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
+$ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
 
@@
 \begin{equation}
-\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT} \geq P_{S}^{CRCT} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha} \dfrac{P_{S + \{i\}}^{MAX}}{P_{S}^{MAX}}
+\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT} \geq P_{S}^{CRCT} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
 \label{eq:marginal_benefit}
@@
 
-In order to understand how this utility function behaves and how it relates to the $ \alpha $ parameter, we use again the gaussian case (as in section \ref{sec:notations}).
-Since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
-Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
-In the following $ \bar{N} $ denotes the mean number of agents in a coalition.
+If $ \alpha $ is set to zero, agents are added as long as they increase the contract value of the coalition.
+If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
+Since tunning the utility function with $ \alpha $ is not necessarily convenient, we relate $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $.
 If we wish that $ \mathcal{U}(S) $ tends to form coalitions of size approximately of the order of $ \bar{N} $, then :
@@
 \left[ \dfrac{\partial{ U}}{ \partial{|S|}} \right]_{|S| = \bar{N}} = 0
+\label{eq:derivative}
 \end{equation}
 
-Solving this equation for $ \alpha $ in the mean approximation for the gaussian case yields :
+In order to get an estimator for $ \alpha $, we solve equation \ref{eq:derivative} in a gaussian case (as in section \ref{sec:notations}).
+Furthermore, since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
+Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
+Solving equation \ref{eq:derivative} for $ \alpha $ in these conditions yields :
+
 \footnotesize
@@
 The redistribution of a coalition's utility in terms of individual payoffs will therefore not be considered directly in this paper.
-$ \mathcal{U} $ can be rather interpreted as a measure of how good a given coalition is according to our criteria since it favors small coalitions with a good production to risk ratio.
+$ \mathcal{U} $ can be rather interpreted as a measure of how good a given coalition is according to our criteria since it favors coalitions with good production to risk ratios.
 
@@
 
-The previous section explained that coalitions with small variances in their production probability distributions are more likely to afford high contracts.
+\subsubsection{Filtering the correlation graph}
+
+The previous section explained that coalitions with small variances in their production probability distributions are more likely to afford better contracts.
 Furthermore, the gaussian approximation highlighted that the correlation structure of the agents plays an important role in finding stable coalitions.
@@
 
-Selecting the right filter $ \epsilon $ is an important point since it affects the correlation structure which affects in turn the formation of the coalitions.
+Selecting the right filter $ \epsilon $ for the $ \epsilon $-graph is an important point since it affects the correlation structure which affects in turn the formation of the coalitions.
 Unfortunately, there seems to be no clear consensus in the litterature on how to select such a threshold.
-In our situation, we want to generate $ N_{COAL} $ coalitions from cliques in the graph.
-We need therefore at least $ N_{COAL} $ cliques of a given size to start.
-Besides, since we consider coalitions as disjoint sets, the starting cliques should be non overlapping.
-We thus select the optimal threshold as :
+We will see later in this section that cliques in $ G_{2} $ are potential seeds for the coalitions.
+Since we want to generate $ N_{COAL} $ coalitions, we need at least $ N_{COAL} $ cliques of a given size to start.
+Besides, since we consider coalitions as disjoint, the starting cliques should be non overlapping.
+We thus select our optimal threshold for $ G_{2} $ as :
 
@@
 Where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
-In other words we select $ \epsilon $ as the smallest threshold posible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
+In other words we select $ \epsilon^{\star} $ as the smallest threshold posible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
 The existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
 The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
+\\
+
+\subsubsection{Cliques as de-correlated starting seeds}
 
@@
 That is, if a and b are strongly correlated ($\rho_{ab} > 0.9 $) and b and c are also strongly correlated ($\rho_{bc} > 0.9 $), then there is a high probability for a and c of being strongly correlated ($\rho_{ac} > 0.62 $).
-This is one of the reasons why searching for clusters in correlation graphs, that is, clustering according to the correlation, makes sense.
 
 Nevertheless de-correlation seems like a more complex concept than correlation in the sense that there is not even a partial notion of transitivity when it comes to it.
-As expected, the clustering coefficients of $ G_{1} $ is much higher than the one of $ G_{2} $.
+Therefore, the clustering coefficients of $ G_{1}^{\epsilon} $ is much higher than the one of $ G_{2}^{\epsilon} $.
 This can be seen as another formulation of Onnela's study on the structural roles of weak and strong links on correlation graphs.
 Strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters.
-Searching for clusters in $ G_{2} $ and hoping that this strategy will provide a nice coalition structure of internally uncorrelated coalitions seems thus pointless.
+Searching for clusters in $ G_{2}^{\epsilon} $ and hoping that this strategy will provide a nice coalition structure of internally uncorrelated coalitions seems thus pointless.
 
-Consider now a clique in  $ G_{2} $, which is a complete subgraph of $ G_{2} $.
+Consider now a clique in  $ G_{2}^{\epsilon} $, which is a complete subgraph of $ G_{2}^{\epsilon} $.
 This is indeed a structure of interest for our purpose.
 Since there is a link for every pairs of nodes, we know, by construction, that a clique has a mean correlation and a maximum correlation less than $ \epsilon $.
-More formally, let
 
-\begin{equation}
-\left\{ \begin{array}{lll}
-			\bar{\rho}_{S} = \dfrac{2}{|S|(|S|-1)} \sum_{i \in S} \sum_{j \in S, j>i} \rho_{ij}, \\ 
-			\rho_{S}^{>} = MAX_{(i,j) \in S^{2}}(\rho_{ij})
-\end{array} \right.			
-\end{equation}
-
-be the mean correlation of coalition S and the maximum correlation value in S.
-If S is a clique in $ G_{2} $ and $ \epsilon \in [0,1] $ is the threshold used for filtering $ G_{2} $, then $ \bar{\rho}_{S} \leq 2/(|S|(|S|-1)) \sum_{i \in S} \sum_{j \in S, j>i} \epsilon $, that is, $ \bar{\rho}_{S} \leq \epsilon $, and $ \rho_{S}^{>} \leq \epsilon $.
+\begin{figure}
+\includegraphics[scale=.6]{cliques_histo.pdf}
+\caption{Histograms of utility values for coalitions of size 3.
+Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets.
+As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
+\label{fig:histo_cliques}
+\end{figure}
 
@@
 This can be quantified by the marginal benefit of equation \ref{eq:marginal_benefit}.
-
-\begin{figure}
-\includegraphics[scale=.6]{cliques_histo.pdf}
-\caption{Histograms of utility values for coalitions of size 3.
-Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets.
-As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
-\label{fig:histo_cliques}
-\end{figure}
 
@@
 
-The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities, and therefore, high contract values for relatively small sizes.
+The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities.
 Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the optimal coalitions set is very low.
 Actually, it is not obvious that a strict optimum exists at all.
-Besides, there is no, to our knowledge, state of the art algorithm that aggregate uncorrelated series in an optimum way.
+Besides, there is, to our knowledge, no state of the art algorithm that aggregate uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems in finance).
 
@@
 \item \textbf{Correlated :} This is the complete opposite of our algorithm.
-It basically uses the correlation graph $ G_{1} $ and performs community detection on it.
+It uses the correlation graph $ G_{1} $ and performs a community detection.
 The resulting coalitions have thus very high internal correlations.
@@
 See algorithm \ref{alg:algo3}.
-\item \textbf{K-means :} This is simply a K-means clustering of the agents $ P_{i} $'s series.
-The goal of presenting this algorithm is to show that K-means do not consider correlations explicitely and is therefore not well suited for our purpose.
 \end{itemize}
 
-The coalitions formed according to these algorithms will be compared using two criteria : the normalized global utility of the coalitions formed $ \bar{\mathcal{U}}(CS) \in [0,1] $, and the percentage of valid coalitions that satisfy the grid conditions $ \tau(CS) \in [0,1] $ :
-\begin{equation}
-	\bar{\mathcal{U}}(CS) = \dfrac{1}{N_{COAL}} \sum_{S \in CS} \mathcal{U}(S)
-\end{equation}
-\begin{equation}
-	\tau(CS) = \dfrac{|\{ S \in CS,\ s.t\ \mathcal{U}(S) > 0 \}|}{N_{COAL}}
-\end{equation}
-
 Before running the algorithm, we need to calibrate the utility function by choosing the $ \alpha $ parameter.
+Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
 In this paper, neither the communication network nor the electrical grid are explicitely considered.
-Thus, we do not have any technical constraints on coalition sizes even if we designed the utitility such that these kind of constraints could be taken into account.
+Thus, we do not have any technical constraints on coalition sizes even if we designed the utitility such that these could be taken into account.
 We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
@@
 Using equation \ref{eq:alpha_star} to select $ \alpha $ based on the mean quantities seems to give acceptable results for the utility function.
+
+Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
+The transition from unvalid to valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
+After this transition, coalition's utilities improve slowly up to a maximum point.
 
@@
 \label{fig:real_utility2}
+\end{figure}
+
+\begin{figure}
+\includegraphics[scale=.6]{search.pdf}
+\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
+\label{fig:search}
+\end{figure}
+
+Figure \ref{fig:coalitions} shows the coalitions formed with the considered algorithms.
+A coalition is represented by a marker in the contract value / volatility space, and the color and shape of a marker indicates by which algorithm the coalition has been formed.
+Besides, the size of a marker is proportional to the coalition size in number of agents.
+The coalitions of correlated agents are clearly of poor quality according to our criteria since they can only afford small production's contracts, and with a very high volatility (compared to the other coalitions on the plot).
+On figure \ref{fig:coalitions}, the de-correlated coalitions and the randomized ones are closer to the bottom right corner indicating a much better quality.
+We can see that the de-correlated coalitions are globally more located toward the bottom right corner, meaning that there are less disparities between coalitions than with the random case.
+Besides, we can see that the coalitions seem quite balanced in term of the number of agents that they contain.
+
+\begin{figure}
+\includegraphics[scale=.6]{./figures/coalitions/coalitions.pdf}
+\caption{Coalitions formed in the (contract value, volatility) space.
+The color map indicates qualities of portions of the plane.
+The closer to red the better (high contract values with small volatility).
+On the opposite, blue areas shows poor quality (small contract values with high volatility).
+Blue dots stand for the un-correlated coalitions, red triangles for random search, and green squares for correlated coalitions.}
+\label{fig:coalitions}
 \end{figure}

### 2015-05-04 Article.tex
update

@@
 
+A key point for the coalitions, besides stability and productivity, is their resilience.
+The resilience of a system can be roughly described as its ability to perform its tasks when subject to failures of its components.
+Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when node failures occur.
+According to our model, the injected power of a coalition structure ($ P_{CS}^{inject} $) is nothing more than the sum of the contract values of its coalitions :
+
+\begin{equation}
+P_{CS}^{inject} = \sum_{S \in CS} P_{S}^{CRCT}
+\label{eq:inject}
+\end{equation}
+
+In this section the following scenario is played : a coalition structure is formed with a given algorithm and its injected power is considered as a reference value.
+Fonctional prosumers are then randomly selected and removed from their respective coalitions.
+Each time a prosumer is removed, the contract value of its coalition is recomputed, and therefore $ P_{CS}^{inject} $ is also recomputed according to equation \ref{eq:inject}.
+We denote by $x$ the number of nodes removed, and for the sake of comparison between algorithms, we observe the percentage of initially injected power as a function of $ x $ :
+
+\begin{equation}
+\tau_{CS}^{inject}(x) = \dfrac{P_{CS}^{inject}(x)}{P_{CS}^{inject}(0)} 
+\label{eq:perc}
+\end{equation}
+
+Figure \ref{fig:resilience} exhibits how the percentage of initially injected power (see equation \ref{eq:perc}) evolves with random failures.
+The three curves show the averaged results for the three coalition structures obtained with the three studied algorithms.
+The x axis stands for the number of nodes that failed.
+For $ x = 0 $, coalitions are all intact, and thus inject $ 100\% $ of their initial capacities.
+When a node fails, the production capacity of its former coalition drops, forcing the latter to lower its contract value, which in turn decreases the injected power.
+There is a point where the coalition cannot provide a sufficient contract value ($P_{S}^{CRCT} < P^{MIN} $) and has to leave the market.
+These coalition failures have a strong impact as they cut down the injected power to a bigger extent than a single node failure.
+As visible on figure \ref{fig:resilience}, the de-correlated coalitions achieve a more resilient production on the market in the sense that their market production is less impacted by the failures compared with random and correlated coalitions.
+
+\begin{figure}
+\includegraphics[scale=.6]{./figures/resilience/resilience.pdf}
+\caption{Percentage of initially injected power (see equation \ref{eq:perc}) when nodes fail randomly. The curves are averaged over 100 realizations.}
+\label{fig:resilience}
+\end{figure}
+
 %Let $ P^{MARKET} = \sum_{S \in CS} P_{S}^{CRCT} $ be the production on the market, where CS is the coalition structure formed $ CS = \{S_{1},...S_{N_{COAL}} \} $, and $ N_{COAL} $ is the number of coalitions in the structure.

### 2015-05-11 Article.tex
update

--- 2015-04-29 Article.tex
@@
 
+Because agents are susceptible to fail for diverse reasons, the impact on the system's stability is critical.
+The propensy of a system to undertake these failures is often refered to as its resilience.
+Once formed, we also investigate in this paper the resilience of the coalitions when prosumers fail.
+Despite the fact that loosing agents is always detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
+
 The paper is organized as follows, section 2 will give a brief overview of the related litterature, section 3 will clarify how we generated realistic prosumer production traces based on weather data.
@@
 However, one of the assumptions in the portfolio theory is that returns are normaly distributed random variables (or at least that the joint distribution is elliptic).
-This is a strong assumption which reveals unrealistic in most cases for our purposes when looking at real data.
+In this paper, we chose to use normal distributions only for explanation purposes, or for estimating some parameters.
+Therefore, all the results are computed based on observed distributions (sometines far from gaussians).
 
-Nevertheless, one of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly (as we will see in section X) the variances of the portfolios.
+One of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly (as we will see in section X) the variances of the portfolios.
 Since the work of Mantanega, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph where the weight of an edge between two timeseries (nodes) is the metric value for these series.
@@
 Unfortunately, at the time this paper was written, production and consumption data for prosumers over a large region were not yet available to our knowledge.
-Some interesting experiments are notwithstanding being conducted and data are progressively made public (see for instance the ISSDA experiment in Ireland).
+Some interesting experiments are notwithstanding being conducted and data are progressively made public [].
 
-In the following, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t.
-That is, $ P_{i}(t) = P_{i}^{P}(t) - P_{i}^{D}(t) $, where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
+In the following, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
+
+\begin{equation}
+P_{i}(t) = P_{i}^{P}(t) - P_{i}^{D}(t)
+\end{equation}
+
+Where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
 In other words, $ P_{i}(t) $ represents the instantaneous surplus of power that agent i is willing to sell at time t.
@@
 Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
-A quantity like wind speed depends both on time and location.
+A quantity like wind speed depends both on time and location (we neglect height since this information is rarely present in the data).
 In the following, time is discretized into slots and space into zones.
@@
 Therefore, if prosumers i and j are positioned on the same zone, they are exposed to the same weather.
-Adding some noise can easily be done though not considered in this paper.
+Adding some intra-zone noise can easily be done though not considered in this paper.
 The process for generating the $ P_{i} $ series is pictured in the first block of the process diagram (see figure \ref{fig:process}).
-
-Note that a prosumer i is defined by his zone $ Z_{i} $ as well as the sets $ DER_{i} $ and $ load_{i} $.
-That is, a prosumer can be configured to represent anything from a single windturbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) throug more complex combinations.
-In practice, we use random configurations for the agents.
-
-In the rest of the paper, we use french weather data (see www.infoclimat.fr) starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
 
@@
 \end{figure}
+
+Note that a prosumer i is defined by his zone $ Z_{i} $ as well as the sets $ DER_{i} $ and $ load_{i} $.
+That is, a prosumer can be configured to represent anything from a single windturbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) through more complex combinations.
+In practice, we use random configurations for the agents.
+
+In the rest of the paper, we use french weather data (see www.infoclimat.fr) starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
 
@@
 
-As explained in section \ref{sec:data}, we have a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulated his extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
+This section provides most of the notations and introduces important concepts for the rest of the paper.
+As explained in section \ref{sec:data}, we consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
 Based on these historical values, our objective is now to form groups of prosumers (the so-called coalitions) so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
@@
 This is only for explanation purposes as it is of course rather unrealistic in real situations where the distributions are skewed.
-Using simple statistics, we can write the distribution of the coalition $ \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :
+Using simple statistics, we can write the distribution of the coalition $ S = \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :
 
@@
 If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
-Indeed, in this situation, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid.
+Indeed, in this kind of situations, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid.
 The probability that $ \{i,j\}$ is underproducing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's quality.
@@
 \begin{equation}
-Pr[P_{ij} \leq P^{CRCT}] = \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P^{CRCT} - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] 
+Pr[P_{ij} \leq P_{S}^{CRCT}] = \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{S}^{CRCT} - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] 
 \end{equation}
@@
 
-The amount of risks a given coalition is willing to take depends on a lot of things, among wich its capacity to compensate for under-producing (using batteries, backup generators...).
+The amount of risks a given coalition is willing to take depends on a lot of things, among wich its capacity to compensate for under-producing (using batteries, backup generators...), or the risk aversion of the aggregator.
 Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works.
@@
 The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
-More formally, let $ \phi \in [0,1] $ be a threshold fixed by the grid operator as a maximum value for the probability of under-producing.
+More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the grid operator as a maximum value for the probability of under-producing.
 The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
@@
 \label{eq:P_star}
-P_{S}^{CRCT \star} = \mu_{ij} + \sqrt{2} \sigma_{ij} erf^{-1}(2 \phi - 1 )
+P_{S}^{CRCT \star} = \mu_{ij} - \sqrt{2} \sigma_{ij} erf^{-1}( 1 - 2 \phi )
 \end{equation}
@@
 For illustration, the range of $ \phi $ values is shown from 0 to 1, but in practice, only small values of $ \phi $ really make sense : $ \phi = 1 $ for instance means that coalitions can announce absolutely anything since the probability of producing less than any contract value is necessarily less than one by trivial definition of a probability.
-As visible on figure \ref{fig:Gaussian}, coalitions with high expected production but presenting a high unpredictability are penalized and can only afford small contracts.
+As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts.
 
@@
 
-Before running the algorithm, we need to calibrate the utility function by choosing the $ \alpha $ parameter.
+Before running the algorithms, we need to calibrate the utility function by choosing the $ \alpha $ parameter.
 Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
@@
 A coalition is represented by a marker in the contract value / volatility space, and the color and shape of a marker indicates by which algorithm the coalition has been formed.
-Besides, the size of a marker is proportional to the coalition size in number of agents.
-The coalitions of correlated agents are clearly of poor quality according to our criteria since they can only afford small production's contracts, and with a very high volatility (compared to the other coalitions on the plot).
-On figure \ref{fig:coalitions}, the de-correlated coalitions and the randomized ones are closer to the bottom right corner indicating a much better quality.
-We can see that the de-correlated coalitions are globally more located toward the bottom right corner, meaning that there are less disparities between coalitions than with the random case.
-Besides, we can see that the coalitions seem quite balanced in term of the number of agents that they contain.
+Besides, as the size of a marker is proportional to the number of agents in the coalition, we can see that the utility function results in approximatively balanced coalitions.
+The coalitions of correlated agents (green squares) are clearly of poor quality according to our criteria since they can only afford small production contracts, and with a very high volatility (compared to the other coalitions on the plot).
+
+On figure \ref{fig:coalitions}, the de-correlated coalitions (blue dots) and the random ones (red triangles) are closer to the bottom right corner indicating a much better quality.
+We can see that, despite its simplicity, the random algorithm is able to achieve correct results thanks to the diversification inherent to such algorithm.
+Nevertheless, based on the utility, coalitions resulting from this random search are less good than the de-correlated ones.
+This can be understood on figure \ref{fig:coalitions} as the de-correlated coalitions are globally more located toward the bottom right corner, meaning that there are less disparities between coalitions than with the random case.
 
@@
 \label{fig:coalitions}
+\end{figure}
+
+A key point for the coalitions, besides stability and productivity, is their resilience.
+The resilience of a system can be roughly described as its ability to perform its tasks when subject to failures of its components.
+Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when node failures occur.
+According to our model, the grid operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $.
+As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market.
+We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold :
+
+\begin{equation}
+\mathcal{R}_{S} = 1 - Pr[P_{S} < P^{MIN}]
+\end{equation}
+
+And we extend this measure to the coalition stuctures :
+
+\begin{equation}
+\mathcal{R}_{CS} = \prod_{S \in CS} \left( 1 - Pr[ P_{S} < P^{MIN} ] \right)
+\label{eq:resilience}
+\end{equation}
+
+We consider that prosumers fail randomly, and we denote by $ \psi \in [0,1] $ the fraction of agents that failed.
+Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolve according to $ \psi $.
+On the top subplot, $ P^{MIN} $ was volontarily selected relatively low such that the resiliences of the three structures fit on the same figure.
+When the $ P^{MIN} $ requirement increases, the differences between the algorithms also increase as visible on the bottom subplot of figure \ref{fig:resilience}.
+The de-correlated coalitions seem to achieve a more resilient production on the market in the sense that they are able to sustain a higher fraction of node failures.
+
+\begin{figure}
+\includegraphics[scale=.6]{./figures/resilience_both.pdf}
+\caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
+\label{fig:resilience}
 \end{figure}

### 2015-05-12 Article.tex
update

@@
 \begin{abstract}
+Achieving a successful energetic transition became a major goal for the 21st century in most developped countries.
+The reconciliation of several research areas and their recent focus on the same smart grid objective has led to major technical improvements.
+For instance, allowing large penetrations of renewables into the electricity production seemed more like a dream yesterday, but appears more and more doable as research progresses.
+The inherent volatility of some of these energy sources poses indeed various problems in such a constrained environment.
+It is often supposed that a large number of these small generators will be decentralized down to the distribution networks and even be owned by individuals.
+By enabling bi-directional flows, this vision revolutionises the nature of the simple todays's consumer into some futuristic agent that both consumes and produces.
+It is often considered that these individual productions will be used only to meet the agent own demand (in real time or through batteries).
 
+In this paper, we consider agents that tend to produce more than they consume and seek to sell their surplus of production.
+More precisely, we rely on aggregations of these agents that can be seen as virtual power plants with production and consumption components that are both variable.
+Because these virtual power plants are complex aggregations of different types of production means and consumption habits, the volatilities of the plants productions are deeply impacted by the correlation between the agents.
+We propose here an algorithm that takes the underlying correlation structure of the agents into account while forming the coalitions.
+We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
 \end{abstract}

### 2015-05-18 Article.tex
update

@@
 It is often supposed that a large number of these small generators will be decentralized down to the distribution networks and even be owned by individuals.
-By enabling bi-directional flows, this vision revolutionises the nature of the simple todays's consumer into some futuristic agent that both consumes and produces.
-It is often considered that these individual productions will be used only to meet the agent own demand (in real time or through batteries).
+By enabling bi-directional flows, this vision revolutionises the nature of the simple today's consumer into some futuristic agent that can both consume and produce.
+Such an agent can obviously reduce its electricity bill by seeking energetic independance through demand side management programs coupled with storage.
+Nevertheless, as it could be the case that an agent's production exceeds its consumption and storage capacity, its ability to sell its extra-production appears as a way to save energy.
 
-In this paper, we consider agents that tend to produce more than they consume and seek to sell their surplus of production.
+In this paper, we thus consider agents that tend to produce more than they consume and seek to sell their surplus of production.
 More precisely, we rely on aggregations of these agents that can be seen as virtual power plants with production and consumption components that are both variable.
@@
 
+In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator.
+By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions.
+The power that a coalition is able to propose on the market is therefore related to both quantities.
+As the correlations between the prosumers that form these coalitions impact directly their volatilities, we seek uncorrelated aggregations of agents.
+We used a greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements.
+We compare the results with a random search and an opposite strategy that clusters correlated agents together.
+We showed that the coalitions resulting from our algorithm are able to provide more stable power to the grid because they tend to have globally better production over volatility ratios while the others are generally more disparate.
+
+Because in real situations, agents are prone to failure, the resilience of the coalitions is an important criterium.
+We therefore studied how the coalitions are able to remain on the market when their agents fail randomly one by one.
+We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-05-21 Article.tex
update

@@
 The reconciliation of several research areas and their recent focus on the same smart grid objective has led to major technical improvements.
-For instance, allowing large penetrations of renewables into the electricity production seemed more like a dream yesterday, but appears more and more doable as research progresses.
+For instance, allowing a large penetration of renewables into the electricity production grid seemed more like a dream yesterday, but appears more and more doable as research progresses.
 The inherent volatility of some of these energy sources poses indeed various problems in such a constrained environment.
-It is often supposed that a large number of these small generators will be decentralized down to the distribution networks and even be owned by individuals.
-By enabling bi-directional flows, this vision revolutionises the nature of the simple today's consumer into some futuristic agent that can both consume and produce.
-Such an agent can obviously reduce its electricity bill by seeking energetic independance through demand side management programs coupled with storage.
+It is often supposed that a large number of these small generators will be decentralized down to the distribution network and even be owned by individuals.
+By enabling bi-directional flows, this vision revolutionises the nature of today simple consumers into futuristic agents that can both consume and produce.
+Such agents can obviously reduce their electricity bills by seeking energetic independance through demand side management programs coupled with storage.
 Nevertheless, as it could be the case that an agent's production exceeds its consumption and storage capacity, its ability to sell its extra-production appears as a way to save energy.
@@
 Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
-A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines.
+A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Rosas-Casals2007} \cite{Pahwa}.
 As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance.
@@
 It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks.
-That is, they exhibit different characteritics, suggesting that they may behave differently.
+That is, they exhibit different characteritics, suggesting that they may behave differently \cite{Rosas-Casals2007}.
 A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities.
@@
 
-Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results.
+Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}.
 Power grids are ideed dynamical systems governed by physical laws.
 Concepts such as shortest path or centrality metrics have thus to be extended to take into account the particular nature of power grids.
-Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph.
+Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph \cite{Pahwa}.
 It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability.
@@
 
-Understanding power grids seems like a reasonable strarting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
+Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
 As there exists many good articles presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
@@
 An envisaged solution for organizing such a system consists in a market environement, where participating entities announce a production capacity for an upcoming period of time.
-These entities thus commit to injecting exactly at any time the contracted amount of power under financial penalities if they fail.
-Such a market system necessitates obviously some communication architecture to be viable.
-Since the number of prosumers in a system may be large, it seems unrealistic to have a flat centralized topology where a single agent coordinates all the others individually.
-Some articles suggest that a fully decentralized architecture will probably be too complicated to achieve, and that some kind of hierarchical hybrid topology seems more likely to be viable.
+If the contract is bought, the concerned entity commits to injecting exactly at any time the contracted amount of power under financial penalities if it fails.
 
-Moreover, taking into account that the production component of a prosumer is supposed to come mainly from renewables, its over-producing state, and therefore the power he injects in the grid, might be rather unstable.
+In this paper, we consider that the production of a prosumer comes from renewables.
+The over-producing state of the agents, and thus the power they inject in the grid, might be therefore rather unstable.
+It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production.
+Besides, it has been argued that considering coalitions of prosumers as potential providers, and allowing these aggregations to enter the market as virtual power plants could have a stabilizing effect.
+Indeed, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
+One thus expect a more stable and predictable energy production for a coalition than for a single agent.
+However, all possible aggregations are not equal in terms of risks and expected productions.
+This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
 
-It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production.
-Considering virtual coalitions of prosumers as potential providers, and allowing these aggregations to enter the market, could potentially solve some of these problems.
-Indeed, the system operator would only have to communicate with a relatively small number of aggregators, that in turn will use some internal communication protocol to manage the involved agents.
-Moreover, if the aggregation has been done properly, one should expect a more stable and predictable energy production for the coalition than for a single agent.
-This diversification idea is ideed a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
-
-We will see that variability in the coalitions productions can be quantified to a certain extend by the correlation among the agents forming the coalitions.
-Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it will help us decide both what coalitions to form and how much they should sell.
-More precisely, we will build a framework in which the system operator has the possibility of fixing some entrance conditions on the market both in terms of stability and sufficient production.
-Based on predictions, coalitions can decide wether or not to enter and what power quantity they are willing to provide.
+We will see that variability in the aggregated productions can be quantified to a certain extend by the correlation among the agents forming the coalitions.
+Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell.
+More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different agents aggregations.
+We build a framework in which the system operator has the possibility of fixing some entrance conditions on the market both in terms of stability and sufficient production.
+Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide.
 Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or lying) will be exposed to financial penalities.
@@
 Because agents are susceptible to fail for diverse reasons, the impact on the system's stability is critical.
-The propensy of a system to undertake these failures is often refered to as its resilience.
+The propensy of a system to undertake these failures is often refered to as its resilience \cite{Pahwa}.
 Once formed, we also investigate in this paper the resilience of the coalitions when prosumers fail.
-Despite the fact that loosing agents is always detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
+Despite the fact that loosing agents is usually detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
 
-The paper is organized as follows, section 2 will give a brief overview of the related litterature, section 3 will clarify how we generated realistic prosumer production traces based on weather data.
-In section 4, we will define most of the notations and explain why correlation between prosumer is a quantity of interest for our objective.
-Based on the conclusions of section 4, section 5 shows how we approached the problem in a complex system fashion.
-Finally, section 6 will provide some results both on performance of the method and resilience of the coalitions formed.
+The paper is organized as follows, section \ref{sec:related} gives a brief overview of the related litterature, section \ref{sec:data} clarifies how we generated realistic prosumer production traces based on weather data.
+In section \ref{sec:notations}, we define most of the notations and explain why correlation between prosumers is a quantity of interest for our objective.
+Based on the conclusions of section \ref{sec:notations}, we develop in section \ref{sec:utility} a utility function quantifying how much power a coalition can announce on the market given an accepted risk level.
+We then develop, in section \ref{sec:forming} a greedy optimization algorithm that uses decorrelated cliques as inputs and improves their utility over a correlation-constrained environment.
+Finally, section \ref{sec:results} provides some results both on performance of the method and resilience of the coalitions formed.
 
@@
 The intermittent nature of renewables such as wind or solar power introduces new challenges in control design.
-Besides, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present.
+On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present.
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours.
@@
 
-Another popular approach on the production side, is to combined renewable generators with storage devices, such that these are charged when there is a surplus of production, and discharged when the consumption exceeds the production.
+Another popular approach on the production side, is to combine renewable generators with storage devices, such that these are charged when there is a surplus of production, and discharged when the consumption exceeds the production.
 Although simple, this idea causes numerous challenges.
@@
 
-Although there are great progresses in stabilizing the production by demand side managment or storage architectures, the need for good prediction techniques is more important than ever.
+Although there are great progresses in stabilizing the production by demand side management or storage architectures, the need for good prediction techniques is more important than ever.
 Understanding the statistical properties of wind or solar irradiance are already well-established areas of research.
 There exists indeed a very interesting litterature on techniques aiming at predicting wind or solar irradiance profiles for future periods with error margins.
-In this paper, we adopt a slightly different approach : we suppose that the agents have historical records of their production and consumption, and we seek aggregations of agents with high expected production and low variability (wich is sometimes called risk in the present paper).
+In this paper, we adopt a slightly different approach : we suppose that the agents have historical records of their production and consumption, and we seek aggregations of agents with high expected production and low variability.
 
@@
 On the contrary, when investing exclusively on low rik assets, one should expect relatively small gains.
-This tradeoff is formalized in the Markowitz' portfolio theory.
+This tradeoff is formalized in the Markowitz' portfolio theory \cite{Markowitz1959}.
 More precisely, given a set of assets for which we have some historic data of returns, the objective is to find a linear combination of these assets (the so-called portfolio) which maximizes the expected value while minimizing the variance of the portfolio's return.
@@
 This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk / volatility as the variance of the production.
-However, one of the assumptions in the portfolio theory is that returns are normaly distributed random variables (or at least that the joint distribution is elliptic).
+However, it is assume in the portfolio theory \cite{Markowitz1959} that returns are normaly distributed (or at least that the joint distribution is elliptic).
 In this paper, we chose to use normal distributions only for explanation purposes, or for estimating some parameters.
@@
 
-One of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly (as we will see in section X) the variances of the portfolios.
-Since the work of Mantanega, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph where the weight of an edge between two timeseries (nodes) is the metric value for these series.
+One of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly the variances of the portfolios.
+Since the work of \cite{Mantegna1999}, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph where the weight of an edge between two series (nodes) is the metric value for these series.
 
 Because the metric can be computed for all pairs, these graphs are complete and of little use as is.
-Historically, the approach used by Mantanega was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering.
+Historically, the approach used by \cite{Mantegna1999} was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering.
 Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph.
@@
 %
-% Section V: Coalition Formation
+% Section V: Utility Function
 %
@@
 
-\section{Coalition Formation}
-\label{sec:forming}
+\section{Utility Function}
+\label{sec:utility}
 
-Now that we defined the notions of contract values and valid coalitions, we can face the problem of forming the "right" coalitions given a pool of agents.
-In this paper, we consider the "right" coalitions as being the ones that maximize some utility notion that, basically, indicates how much stable power can be injected in the grid.
-This section aims at formalizing this utility notion, while explaining how this utility will be optimized in a greedy fashion over the correlation structure of the agents.
-
-\subsection{Utility function}
-
-We designed our model in a way that coalitions are remunerated proportionnally to their contract values $ \mathcal{U}(S) \propto P_{S}^{CRCT} $.
-That is, if $ \lambda $ is the unitary price rate for electricity, a coalition S injecting $ P_{S}^{CRCT} $ in the grid during a period $ [t_{0},t_{k}] $ earns :
+In this section, we use the notions of contract values and valid coalitions developped in section \ref{sec:notations} in order to design a proper utility function.
+The contract basically indicates the rate at which a coalition has to inject power in the grid.
+It seems then natural that coalitions are remunerated proportionnally to their contract values $ gain(S) \propto P_{S}^{CRCT \star} $.
+More precisely, if $ \lambda $ is the unitary price rate for electricity, a coalition S injecting $ P_{S}^{CRCT \star} $ in the grid during a period $ [t_{0},t_{k}] $ earns :
 
 \begin{equation}
-\mathcal{U}(S ) = \int_{t_{0}}^{t_{k}} \lambda P_{S}^{CRCT} dt = P_{S}^{CRCT} \lambda \Delta_{t}
+gain(S) = \int_{t_{0}}^{t_{k}} \lambda P_{S}^{CRCT \star} dt = P_{S}^{CRCT \star} \int_{t_{0}}^{t_{k}} \lambda dt
 \end{equation}
 
-Since $ \lambda $ appears, in this simple model, just a multiplicative constant, we consider for simplicity $ \lambda = 1 $ in the following.
-The utility of S is now the amount of energy S plans to inject during its contract period.
-Since we focused on power quantities from the begining of the present paper, and because the power injected by S in the grid is supposed to be constant over time, we simplify further the utility to its contract value : $ \mathcal{U}(S) = P_{S}^{CRCT} $.
-Although concise, this formulation suffers a major drawback.
-It is indeed not a concave function, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.
+(since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
+Using $ gain(S) $ directly as a utility function suffers a major drawback.
+It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.
 
-Such a model, that virtually allows infinitely large coalitions, is in practice not realistic.
+Such a model, that virtually allows infinitely large coalitions and contract values, is in practice not realistic.
 There are indeed costs (communication costs for instance) that increase with the coalitions sizes.
@@
 \mathcal{U}(S) = \left\{ \begin{array}{lll}
-							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT} }{P^{MAX}},\ if\ S\ is\ valid, \\
+							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT \star} }{P^{MAX}},\ if\ S\ is\ valid, \\
 							0,\ if\ S\ is\ not\ valid
@@
 
-Where parameter $ \alpha $ controls to what extent the size of the coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
+Where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
 $ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
@@
 \begin{equation}
-\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT} \geq P_{S}^{CRCT} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
+\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT \star} \geq P_{S}^{CRCT \star} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
 \label{eq:marginal_benefit}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section VI: Coalition Formation
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Coalition Formation}
+\label{sec:forming}
+Section \ref{sec:notations} explained how contract values for the coalitions are computed, and in section \ref{sec:utility} we related this quantity to the utility and gains of a coalition.
+As the computation time of this quantity is not negligable, and is increasing in the sizes of the aggregations, we explain in this section the heuristic we used to compute the coalitions.
+
 \subsection{Representing the correlation structure}
+As seen in section \ref{sec:notations}, the variance of the aggregated production impacts directly the contract values, and depends on the covariances between the agents productions.
+We argue here that, by having some representation of the correlation structure between the agents, the search landscape for high utility coalitions could be reduced, such that good coalitions are more likely to be found quickly.
+Usually, this correlation structure is formalized with a covariance matrix or a correlation matrix that contains all the correlation coefficients between the agents : $ M = (\rho_{ij})_{\forall i,j \in \mathcal{A}^{2}}$.
+By using a metric to map this matrix in a weighted adjacency matrix (see section \ref{sec:related}), it is possible to obtain a graph representation of the correlation relationships between the agents.
 
-\subsubsection{Filtering the correlation graph}
-
-The previous section explained that coalitions with small variances in their production probability distributions are more likely to afford better contracts.
-Furthermore, the gaussian approximation highlighted that the correlation structure of the agents plays an important role in finding stable coalitions.
-Usually, this correlation structure is formalized with a covariance matrix or a correlation matrix that contains all the correlation coefficients between the agents : $ M = (\rho_{ij})_{\forall i,j \in \mathcal{A}^{2}}$.
-
-In the following, we use two opposite distance metrics :
+In the following, we use two opposite distance metrics for this mapping :
 
@@
 
-Selecting the right filter $ \epsilon $ for the $ \epsilon $-graph is an important point since it affects the correlation structure which affects in turn the formation of the coalitions.
+In both cases, we want to keep only the edges which weights are located in the lower tail of the distance distributions.
+In other words, we want to compute the $ \epsilon$-graphs of $ G_{1} $ and $ G_{2} $ such that only meaningful edges remain.
+Selecting the right filter $ \epsilon $ is thus an important point since it affects the landscape search for the coalition formation.
 Unfortunately, there seems to be no clear consensus in the litterature on how to select such a threshold.
@@
 
-\subsubsection{Cliques as de-correlated starting seeds}
-
+\subsection{Cliques as de-correlated starting seeds}
 Correlation can be seen as cosine of angles in $ L^{2} $, hence even if there is no strict transivity relation for correlation, there is, to a certain extent, some partial notion of it.
@@
 Therefore, the clustering coefficients of $ G_{1}^{\epsilon} $ is much higher than the one of $ G_{2}^{\epsilon} $.
-This can be seen as another formulation of Onnela's study on the structural roles of weak and strong links on correlation graphs.
+This can be seen as another formulation of \cite{Garas2008} on the structural roles of weak and strong links on financial correlation graphs.
 Strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters.
@@
 
-\subsection{Coalition formation algorithm}
+\subsection{Algorithm}
 
 The algorithm takes inputs from :
+
 \begin{itemize}
@@
 \end{itemize}
+
 The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
 Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
-The next step is a local greedy improvement over the correlation structure represented by  $ G_{2}^{\epsilon^{\star}} $.
+The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
 Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
@@
  		Find $ i^{\star} $ \;
- 		\If{ $ \delta_{clique}(i^{\star}) >= 0 $ }{
+ 		\If{ $ \delta_{clique}(i^{\star}) \geq 0 $ }{
  			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
+ 			}
+ 		\If{ $ \exists j \in clique,\ s.t\ \delta_{clique}(j) < 0 $}{
+ 			$ clique \leftarrow clique - \{j \} $ \;
  			}
@@
   	}
- \caption{Percolation algorithm}
+ \caption{Local greedy optimization algorithm}
  \label{alg:algo1}
@@
 \begin{algorithm}
- \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ maxiter $}
+ \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ Loop^{max} $}
  \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- $ curriter \leftarrow 0 $ \;
+ $ N^{loop} \leftarrow 0 $ \;
  $ CS^{\star} \leftarrow \emptyset $\;
- \While{$ curriter < maxiter $}{
+ \While{$ N^{loop} < Loop^{max} $}{
  	$ CS \leftarrow SelectRandomCS() $\;
@@
  		}
- 	$ curriter \leftarrow  curriter + 1 $\;
+ 	$ N^{loop} \leftarrow  N^{loop} + 1 $\;
   	}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Section VII: Results
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
+\label{sec:results}
 
 The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities.
-Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the optimal coalitions set is very low.
+Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the global optimal coalitions set is very low.
 Actually, it is not obvious that a strict optimum exists at all.
-Besides, there is, to our knowledge, no state of the art algorithm that aggregate uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems in finance).
+Besides, there is, to our knowledge, no state of the art algorithm that aggregate uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems).
 
@@
 Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
-Using equation \ref{eq:alpha_star} to select $ \alpha $ based on the mean quantities seems to give acceptable results for the utility function.
+Using equation \ref{eq:alpha_star} to estimate $ \alpha $ based on the mean quantities and gaussian approximations seems to give acceptable results for the utility function behavior.
 
@@
 \begin{equation}
-\mathcal{R}_{S} = 1 - Pr[P_{S} < P^{MIN}]
+\mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}]
 \end{equation}
@@
 As the correlations between the prosumers that form these coalitions impact directly their volatilities, we seek uncorrelated aggregations of agents.
-We used a greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements.
-We compare the results with a random search and an opposite strategy that clusters correlated agents together.
-We showed that the coalitions resulting from our algorithm are able to provide more stable power to the grid because they tend to have globally better production over volatility ratios while the others are generally more disparate.
+We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation.
+A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results.
+We compare these results with a random search, and an opposite strategy that clusters correlated agents together.
+We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
+Because they tend to have globally better production over volatility ratios, these coalitions will tend to use les storage and spill less energy than more unstable ones.
+We plan to study these benefits for the control of the aggregations in future works.
 
-Because in real situations, agents are prone to failure, the resilience of the coalitions is an important criterium.
+Because in real situations, agents are prone to failure, resilience is also an important criterium for the quality of the aggregations.
 We therefore studied how the coalitions are able to remain on the market when their agents fail randomly one by one.

### 2015-05-22 Article.tex
update

@@
 Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
-A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Rosas-Casals2007} \cite{Pahwa}.
+A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Pagani2012} \cite{Rosas-Casals2007} \cite{Pahwa}.
 As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance.
-These metrics enable scientists to get a better understanding of the structure of such complex systems.
+These metrics enable scientists to get a better understanding of the structure of such complex systems \cite{Pagani2012}.
 It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks.
@@
 It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability.
-A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic.
-It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state.
+A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic \cite{Filatrella2008}.
+It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state \cite{Dorfler2013}.
 
 Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
-As there exists many good articles presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
+As there exists many good articles \cite{Ramchurn} presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
 In this paper, we rather concentrate on a different approach to the system stability.
@@
 His production can be, of course, used to meet his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production to the grid.
-Such an agent model is known as a "\textit{prosumer}" (and will be called accordingly in this paper).
+Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
 
@@
 On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present.
-Unfortunately, these moments do not necessarily coincide with the consumption peak hours.
+Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
 A first approach to remedy this situation by acting on the consumption component is the use of dynamic ectricity prices sent to the end users.
-Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles.
+Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles \cite{Milligan2010}.
 In periods where the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.

### 2015-06-22 Article.tex
intro modif

@@
 
-Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
+%Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
 A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Pagani2012} \cite{Rosas-Casals2007} \cite{Pahwa}.
@@
 
-Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}.
+%Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}.
 Power grids are ideed dynamical systems governed by physical laws.
@@
 
-Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
+Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
+One natural condition for stability revolves around sustaining a power generation that meets the demand at any time.
+If one is larger than the other, the system deviates from its synchronous state.
+If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures.
+With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption.
+Relatively small deviations from this schedule can then be supported almost on real time by using fast power plants, interconnections with border countries, as well as regulator entities.
+Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid.
+Operations are decided on a market where contracts are created based on information from the grid.
+Such a market environment appears necessary in order to improve the reliability of the whole system, and is already being used with large renewable plants in the production portfolio.
+
+The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
+This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution network.
+Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time.
+On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
+These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...), which predictions can be difficult and uncertain.
+Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition.
+The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.
+
+%Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
 As there exists many good articles \cite{Ramchurn} presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
-In this paper, we rather concentrate on a different approach to the system stability.
-We consider a set of agents that are equiped with distributed energy ressources (DER) as well as electrical loads, such that an agent can produce and consume energy at any time.
-His production can be, of course, used to meet his own demand, but in cases where he is over-producing, we consider that he has the possibility to sell his extra-production to the grid.
+
+In this paper, we focus on these nodes in the distribution network that can produce and consume electricity.
+More precisely, we consider a set of agents that can be seen as aggregations of renewable distributed energy ressources (DER) and electrical loads.
+The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
 Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
@@
 The constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
-An envisaged solution for organizing such a system consists in a market environement, where participating entities announce a production capacity for an upcoming period of time.
-If the contract is bought, the concerned entity commits to injecting exactly at any time the contracted amount of power under financial penalities if it fails.
+In this paper, the envisaged solution for organizing such a multi-producers system consists in a market environement, where participating entities announce an estimated production capacity for an upcoming period of time.
+If a contract is bought, the concerned entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
 
@@
 
-We will see that variability in the aggregated productions can be quantified to a certain extend by the correlation among the agents forming the coalitions.
+We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
 Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell.
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different agents aggregations.
-We build a framework in which the system operator has the possibility of fixing some entrance conditions on the market both in terms of stability and sufficient production.
+We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market.
 Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide.
-Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or lying) will be exposed to financial penalities.
+Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.
 
@@
 Once formed, we also investigate in this paper the resilience of the coalitions when prosumers fail.
-Despite the fact that loosing agents is usually detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
+Despite the fact that losing agents is usually detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
 
@@
 Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles \cite{Milligan2010}.
-In periods where the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
+In periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
 On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
@@
 This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk / volatility as the variance of the production.
-However, it is assume in the portfolio theory \cite{Markowitz1959} that returns are normaly distributed (or at least that the joint distribution is elliptic).
+However, it is assumed in the portfolio theory \cite{Markowitz1959} that returns are normaly distributed.
 In this paper, we chose to use normal distributions only for explanation purposes, or for estimating some parameters.
@@
 Weather quantities like wind speed or solar irradiance appear thus as alternative data for generating the $ P_{i} $ series.
-Fortunately, these kind of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
+Fortunately, these kinds of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
 A quantity like wind speed depends both on time and location (we neglect height since this information is rarely present in the data).
@@
 Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
-Finally, the red curve with squares has an optimum size of 15 agents.
+Finally, the red curve with squares has an optimal size of 15 agents.
 }
@@
 
-As can be pointed out, the purpose of $ \mathcal{U} $ is not a study of coalitions stability against player defection, wich game theory provides a lot of tools for.
+As can be pointed out, the purpose of $ \mathcal{U} $ is not a study of coalitions stability against player defection, which game theory provides a lot of tools for.
 This is indeed a problem on its own.
@@
 
-Before running the algorithms, we need to calibrate the utility function by choosing the $ \alpha $ parameter.
+Before running the algorithms, we need to calibrate the utility function by choosing the value of the $ \alpha $ parameter.
 Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
@@
 By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions.
-The power that a coalition is able to propose on the market is therefore related to both quantities.
+The power that a coalition is able to propose on the market is therefore related to production and stability.
 As the correlations between the prosumers that form these coalitions impact directly their volatilities, we seek uncorrelated aggregations of agents.
@@
 We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
-Because they tend to have globally better production over volatility ratios, these coalitions will tend to use les storage and spill less energy than more unstable ones.
+Because they tend to have globally better production over volatility ratios, these coalitions will tend to use less storage and spill less energy than more unstable ones.
 We plan to study these benefits for the control of the aggregations in future works.

### 2015-06-23 Article.tex
intro updates

@@
 
-Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
-One natural condition for stability revolves around sustaining a power generation that meets the demand at any time.
+Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences.
+One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand.
 If one is larger than the other, the system deviates from its synchronous state.
@@
 With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption.
-Relatively small deviations from this schedule can then be supported almost on real time by using fast power plants, interconnections with border countries, as well as regulator entities.
+Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities.
 Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid.
-Operations are decided on a market where contracts are created based on information from the grid.
-Such a market environment appears necessary in order to improve the reliability of the whole system, and is already being used with large renewable plants in the production portfolio.
+It is often assumed that the granularity of these prices is meant to increase in the future as a way to reverberate the production conditions on the end users.
+The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market.
+Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
+These kind of settings are already being used at the transport level with large renewable plants in the production portfolio, and they appear as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.
 
 The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
-This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution network.
+This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
 Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time.
@@
 
-In this paper, we focus on these nodes in the distribution network that can produce and consume electricity.
-More precisely, we consider a set of agents that can be seen as aggregations of renewable distributed energy ressources (DER) and electrical loads.
+In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
+More precisely, we consider a set of agents that can be seen as aggregators of renewable distributed energy ressources (DER) and electrical loads.
 The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
@@
 
-The constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
-In this paper, the envisaged solution for organizing such a multi-producers system consists in a market environement, where participating entities announce an estimated production capacity for an upcoming period of time.
-If a contract is bought, the concerned entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
+As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
+Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges.
+In this paper, the envisaged solution for organizing such a multi-producers system consists in extending the kind of market environement discussed in the previous section.
+Participating aggregators estimate a production capacity for an upcoming period of time and announce it on the market.
+If a contract between the grid and a participant is bought, the latter commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
 
 In this paper, we consider that the production of a prosumer comes from renewables.
-The over-producing state of the agents, and thus the power they inject in the grid, might be therefore rather unstable.
-It has been suggested that energy diversification, geographical expansion, as well as using storage devices are envisageable techniques for  stabilizing the production.
-Besides, it has been argued that considering coalitions of prosumers as potential providers, and allowing these aggregations to enter the market as virtual power plants could have a stabilizing effect.
-Indeed, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
+Contracts are therefore based on predictions and come with a given amount of risk due to the intermittent nature of most renewables.
+Using storage as buffers in order to reduce riks is a popular approach.
+Actually, a whole branch of the smart grid litterature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
+Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable.
+This is clearly unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.
+
+Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities.
+Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side, since an aggregator can be considered as a single node by the level above it.
+On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
 One thus expect a more stable and predictable energy production for a coalition than for a single agent.
-However, all possible aggregations are not equal in terms of risks and expected productions.
+Nevertheless, all coalitions are obviously not equally stable, which means that special attention should be paid to the aggregation step.
+
+Furthermore, in our settings, since the prosumers have both a consumption and a production component, each depending on location and time, the diversification of the aggregations seems not trivial.
 This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
@@
 Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell.
-More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different agents aggregations.
+More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
 We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market.
@@
 \begin{figure}
-\includegraphics[scale=.6]{real_utility.pdf}
-\caption{Utility of random coalitions depending on their size.
-Lines shows mean utility values and colored areas stands for standard deviations.
-The slim blue noisy curve shows the real mean utility, and the thick red curve is its smoothed version by applying a Savitzky-Golay filter.
-On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 10 agents coalitions.}
-\label{fig:real_utility}
-\end{figure}
-
-\begin{figure}
 \includegraphics[scale=.6]{real_utility_2.pdf}

### 2015-06-25 Article.tex
some updates

@@
 Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid.
-It is often assumed that the granularity of these prices is meant to increase in the future as a way to reverberate the production conditions on the end users.
+It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users.
 The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market.
 Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
-These kind of settings are already being used at the transport level with large renewable plants in the production portfolio, and they appear as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.
+These kind of settings are already being used at the transport level of current power grid with large renewable plants in the production portfolio.
+Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.
 
@@
 This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
-Nodes that were simply pure loads yesterday could behave tomorow as generators or loads depending on time.
+Nodes that were simply pure loads yesterday could behave tomorow as generators or loads.
 On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
-These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...), which predictions can be difficult and uncertain.
+These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...).
 Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition.
@@
 Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
+The primary objective of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
+While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are linked to the quality and stability of the injections.
+The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts.
+This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation relationships come into play.
 
-As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state obiously forbid any anarchic system where prosumers inject power whenever and however they want.
+As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state forbid any anarchic system where prosumers inject power without proper control.
 Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges.
@@
 One thus expect a more stable and predictable energy production for a coalition than for a single agent.
-Nevertheless, all coalitions are obviously not equally stable, which means that special attention should be paid to the aggregation step.
+Nevertheless, all coalitions are obviously not equally stable or productive, which means that special attention should be paid to the aggregation step.
 
@@
 We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
-Understanding the correlation reliationships among the agents can thus be illuminating in the sense that it gives an indication about what coalitions to form and how much they should sell.
+Understanding the correlation reliationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
@@
 
-\subsection{Algorithm}
-
-The algorithm takes inputs from :
-
-\begin{itemize}
-	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
-	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
-	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
-\end{itemize}
-
-The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
-Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
-The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
-Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
-This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
-The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
-See the details in algorithm \ref{alg:algo1}.
-
-\begin{algorithm}
- \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- Compute $ G_{2}^{\epsilon^{\star}} $ \;
- Find the $ N_{COAL} $ cliques in $ G_{2}^{\epsilon^{\star}} $\;
- \While{$ \mathcal{U}(CS) $ is improving}{
- 	\For{each clique}{
- 		Find $ i^{\star} $ \;
- 		\If{ $ \delta_{clique}(i^{\star}) \geq 0 $ }{
- 			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
- 			}
- 		\If{ $ \exists j \in clique,\ s.t\ \delta_{clique}(j) < 0 $}{
- 			$ clique \leftarrow clique - \{j \} $ \;
- 			}
-   		}
-  	}
- \caption{Local greedy optimization algorithm}
- \label{alg:algo1}
-\end{algorithm}
-
-\begin{algorithm}
- \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ Loop^{max} $}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- $ N^{loop} \leftarrow 0 $ \;
- $ CS^{\star} \leftarrow \emptyset $\;
- \While{$ N^{loop} < Loop^{max} $}{
- 	$ CS \leftarrow SelectRandomCS() $\;
- 	\If{ $ \mathcal{U}(CS) > \mathcal{U}(CS^{\star}) $ }{
- 		$ CS^{\star } \leftarrow CS $\;
- 		}
- 	$ N^{loop} \leftarrow  N^{loop} + 1 $\;
-  	}
-  return $ CS^{\star} $
- \caption{Random algorithm}
- \label{alg:algo2}
-\end{algorithm}
-
-\begin{algorithm}
- \KwData{$P_{i}$ series,\\ Desired number of coalitions $ N_{COAL} $,\\ search step size $ \beta << 1 $}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- $ \epsilon \leftarrow 1 $ \;
- $ CS \leftarrow \emptyset $\;
- \While{$ |CS| < N_{COAL} $}{
- 	Compute $ G_{1}^{\epsilon} $ \;
- 	$ CS \leftarrow computeClusters( G_{1}^{\epsilon} ) $\;
- 	\If{ $ |CS| = N_{COAL} $ }{
- 		return CS\;
- 		}
- 	\Else{ 
- 	$ \epsilon \leftarrow \epsilon - \beta $\;
- 	}
-  	}
- \caption{Correlated algorithm}
-\label{alg:algo3}
-\end{algorithm}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities.
-Nevertheless, as it comprises mainly of a greedy optimization based on local improvements, the probability that the algorithm finds the global optimal coalitions set is very low.
-Actually, it is not obvious that a strict optimum exists at all.
-Besides, there is, to our knowledge, no state of the art algorithm that aggregate uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems).
+As it comprises mainly of a greedy optimization based on local improvements, there is no guarantee that the algorithm finds the global optimum.
+Since there is, to our knowledge, no state of the art algorithm that aggregates uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems), we compare the results with :
 
-In order to have an idea about the algorithm's quality, we compare its results with :
 \begin{itemize}
@@
 \begin{figure}
-\includegraphics[scale=.6]{./figures/coalitions/coalitions.pdf}
+\includegraphics[scale=.5]{./figures/coalitions/coalitions6.pdf}
 \caption{Coalitions formed in the (contract value, volatility) space.
@@
 
+\appendix
+\subsection{Algorithms}
+
+The algorithm takes inputs from :
+
+\begin{itemize}
+	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
+	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
+	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
+\end{itemize}
+
+The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
+Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
+The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
+Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
+This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
+The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
+See the details in algorithm \ref{alg:algo1}.
+
+\begin{algorithm}
+ \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
+ \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
+ Compute $ G_{2}^{\epsilon^{\star}} $ \;
+ Find the $ N_{COAL} $ cliques in $ G_{2}^{\epsilon^{\star}} $\;
+ \While{$ \mathcal{U}(CS) $ is improving}{
+ 	\For{each clique}{
+ 		Find $ i^{\star} $ \;
+ 		\If{ $ \delta_{clique}(i^{\star}) \geq 0 $ }{
+ 			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
+ 			}
+ 		\If{ $ \exists j \in clique,\ s.t\ \delta_{clique}(j) < 0 $}{
+ 			$ clique \leftarrow clique - \{j \} $ \;
+ 			}
+   		}
+  	}
+ \caption{Local greedy optimization algorithm}
+ \label{alg:algo1}
+\end{algorithm}
+
+\begin{algorithm}
+ \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ Loop^{max} $}
+ \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
+ $ N^{loop} \leftarrow 0 $ \;
+ $ CS^{\star} \leftarrow \emptyset $\;
+ \While{$ N^{loop} < Loop^{max} $}{
+ 	$ CS \leftarrow SelectRandomCS() $\;
+ 	\If{ $ \mathcal{U}(CS) > \mathcal{U}(CS^{\star}) $ }{
+ 		$ CS^{\star } \leftarrow CS $\;
+ 		}
+ 	$ N^{loop} \leftarrow  N^{loop} + 1 $\;
+  	}
+  return $ CS^{\star} $
+ \caption{Random algorithm}
+ \label{alg:algo2}
+\end{algorithm}
+
+\begin{algorithm}
+ \KwData{$P_{i}$ series,\\ Desired number of coalitions $ N_{COAL} $,\\ search step size $ \beta << 1 $}
+ \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
+ $ \epsilon \leftarrow 1 $ \;
+ $ CS \leftarrow \emptyset $\;
+ \While{$ |CS| < N_{COAL} $}{
+ 	Compute $ G_{1}^{\epsilon} $ \;
+ 	$ CS \leftarrow computeClusters( G_{1}^{\epsilon} ) $\;
+ 	\If{ $ |CS| = N_{COAL} $ }{
+ 		return CS\;
+ 		}
+ 	\Else{ 
+ 	$ \epsilon \leftarrow \epsilon - \beta $\;
+ 	}
+  	}
+ \caption{Correlated algorithm}
+\label{alg:algo3}
+\end{algorithm}
+
+\subsection{Net production series}
+
+Data were collected from [] (similar data can be found at [] for the United States). The variables used in the simulation are :
+
+\begin{itemize}
+\item Average wind speed (in $m.s^{-1}$)
+\item Nebulosity (integer in $[0,8] $)
+\item Temperature (in degree celsius)
+\end{itemize}
+
+Power curves are functions that, for a given type of generator, map the input to the output power.
+For windturbines and solar arrays these functions are well studied and approximations have been proposed.
+For the wind turbines, the power curve is thus specified by 4 values :
+
+\begin{itemize}
+\item Cut-in-speed : The wind speed at which the turbine first starts to rotate and generates power.
+\item Rated-output-power : The maximum power that the turbine can generate.
+\item Rated-output-speed : The wind speed at which the turbine attains its rated output power.
+\item Cut-out-speed : The speed at which the turbine is turned off as not to damage the rotor.
+\end{itemize}
+
+The most interesting part is the increase of output power when the wind speed is in the cut-in-speed rated-output-speed range.
+Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the litterature.
+
+The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data.
+As we mainly collected nebulosity series, we used the Helios model described in [].
+This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps.
+As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see [] for more details) :
+
+\begin{equation}
+\left\{ \begin{array}{lll}
+			\Psi{real}(t) = \Psi_{perfect}(t) \eta(t) \\
+			\eta(t) = 1-0.75 \left( \dfrac{N(t)}{8} \right)^{3.4}
+\end{array} \right.
+\end{equation}
+
+Where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
+
+Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :
+
+\begin{equation}
+\mathcal{F}_{PV}(\Psi{real}(t)) =  S_{PV} \Psi{real}(t)  e_{PV}
+\end{equation}
+
+Where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
+The very simple form of this power curve is due to some simplifications in order not to overload the model.
+For instance, it does not take into account angles and orientations degradations.
+These could be incorporated if needed by changing the power curve in the simulations.
+
+data, helios model, power curves, scema
+
 \end{document}

### 2015-06-26 Article.tex
update

@@
 \usepackage{hyperref}
-
+\usepackage{fancyhdr}
 %
@@
 \begin{document}
+\pagestyle{plain}
+\pagenumbering{arabic}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We denote by $ \nu_{i}(t) $ and $ \xi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
+We denote by $ \nu_{i}(t) $ and $ \Psi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
 \begin{equation}
- P_{i}^{P}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\xi_{i}(t) ) 
+ P_{i}^{P}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\Psi_{i}(t) ) 
 \end{equation}
@@
 The model can be easily extended to muplitiple power curves accounting for different types of generators.
-More details about power curves and their approximations can be found in [].
+More details about power curves and their approximations can be found in the appendix and in \cite{Lydia2014}.
 
@@
 
-In the rest of the paper, we use french weather data (see www.infoclimat.fr) starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
+In the rest of the paper, we use french weather data \cite{Infoclimat} starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
 
@@
 
+\subsection{Algorithm}
+The algorithm takes inputs from :
+
+\begin{itemize}
+	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
+	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
+	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
+\end{itemize}
+
+The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
+Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
+The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
+Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
+This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
+The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
+See the details in algorithm \ref{alg:algo1} in the appendix.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{itemize}
-\item \textbf{Random :} This algorithm is basically a random search over the coalition structures space.
-It analyses a given number of structures and returns the best that it has encountered so far.
-See algorithm \ref{alg:algo2}.
+\item \textbf{Random sampling of coalitions :} Coalitions are formed randomly without any other constraint that the desired size.
+This enables us to have an idea about the distributions of utility values for coalitions of a given size.
+\item \textbf{Random sampling of coalition structures :} Coalition structures are sampled randomly by shuffling and random divisions of the agents.
+Algorithm \ref{alg:algo2} uses such a sampling and returns the highest utility coalition structure sampled.
+This algorithm will be refered to as "\textit{random}" in the following.
 \item \textbf{Correlated :} This is the complete opposite of our algorithm.
@@
 
-Figure \ref{fig:coalitions} shows the coalitions formed with the considered algorithms.
-A coalition is represented by a marker in the contract value / volatility space, and the color and shape of a marker indicates by which algorithm the coalition has been formed.
-Besides, as the size of a marker is proportional to the number of agents in the coalition, we can see that the utility function results in approximatively balanced coalitions.
-The coalitions of correlated agents (green squares) are clearly of poor quality according to our criteria since they can only afford small production contracts, and with a very high volatility (compared to the other coalitions on the plot).
+Figure \ref{fig:coalitions} shows the coalitions formed with the considered algorithms in the contract value / volatility space.
+The color map in the background indicates regions where we expect high utilities (red) and the ones where we expect very poor utility values (blue).
+The bottom right corner, with high contract values and low volatilities, is therefore the region where we wish to form our coalitions.
+A single coalition is represented by a marker, and the color and shape of a marker indicates by which algorithm the coalition has been formed.
+Besides, the sizes of the coalitions are indicated on the markers, and the marker size is also proportional to the coalition size.
+We can see that the utility function results in approximatively balanced coalitions.
+Small yellow markers indicates the gravity centers of their respective coalition structures.
+The coalitions of correlated agents (green squares) are clearly of poor quality according to our criteria since they can only afford small production contracts, and with a very high volatility.
 
-On figure \ref{fig:coalitions}, the de-correlated coalitions (blue dots) and the random ones (red triangles) are closer to the bottom right corner indicating a much better quality.
-We can see that, despite its simplicity, the random algorithm is able to achieve correct results thanks to the diversification inherent to such algorithm.
-Nevertheless, based on the utility, coalitions resulting from this random search are less good than the de-correlated ones.
-This can be understood on figure \ref{fig:coalitions} as the de-correlated coalitions are globally more located toward the bottom right corner, meaning that there are less disparities between coalitions than with the random case.
+On figure \ref{fig:coalitions}, the decorrelated coalitions (blue dots) are closer to the bottom right corner indicating a much better quality in term of productivity over volatility ratio.
+The black dotted line indicates the mean values for the random coalitions sampling technique.
+Each small dot stands for the mean position of all sampled coalitions of this given size.
+Variances are not indicated for readalibility, but are usually quite large since this sampling only take the size as a constraint.
+Moreover, note that only a subset of the points are labeled with the size they represents as not to overload the figure.
+We can see that as coalitions get larger, they tend to increase on average their contract values, but at the price of a higher volatility.
+
+On figure \ref{fig:coalitions}, the results of the random coalition structure sampling is shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
+Since the center of the ellipses stands for the mean and each ellipse adds one standard deviation, more than $ 99\% $ of the sampled gravity centers are within the largest ellipse.
+The small yellow dot below the ellipses indicates the gravity center of our solution.
+It is thus visible that our greedy graph based algorithm is able to find a quite good coalition structure in terms of volatility and contract values.
 
 \begin{figure}
-\includegraphics[scale=.5]{./figures/coalitions/coalitions6.pdf}
+\includegraphics[scale=.6]{./figures/coalitions/coalitions6.pdf}
 \caption{Coalitions formed in the (contract value, volatility) space.
@@
 On the opposite, blue areas shows poor quality (small contract values with high volatility).
-Blue dots stand for the un-correlated coalitions, red triangles for random search, and green squares for correlated coalitions.}
+Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions.
+The smaller yellow markers stands for the gravity centers of the coalition structures.
+The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increase (a subset of the points are labeled by the size of the coalition they represent).
+Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition.
+As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse add 1 standard deviation).
+For comparison with our algorithm where the utility function favors balanced coalitions, we also constrained loosely the sizes of the coalitions in the random process.
+}
 \label{fig:coalitions}
@@
 A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results.
-We compare these results with a random search, and an opposite strategy that clusters correlated agents together.
+We compare these results with random samplings, and an opposite strategy that clusters correlated agents together.
 We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
@@
 \subsection{Algorithms}
-
-The algorithm takes inputs from :
-
-\begin{itemize}
-	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
-	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
-	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
-\end{itemize}
-
-The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
-Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
-The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
-Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
-This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
-The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
-See the details in algorithm \ref{alg:algo1}.
 
@@
  \While{$ N^{loop} < Loop^{max} $}{
- 	$ CS \leftarrow SelectRandomCS() $\;
+ 	$ CS \leftarrow SelectRandomCS( ) $\;
  	\If{ $ \mathcal{U}(CS) > \mathcal{U}(CS^{\star}) $ }{
@@
 
-Data were collected from [] (similar data can be found at [] for the United States). The variables used in the simulation are :
+Data were collected from \cite{Infoclimat} (similar data can be found at \cite{NCDC} for the United States). The variables used in the simulation are :
 
@@
 
+\subsubsection{Wind power curve}
 Power curves are functions that, for a given type of generator, map the input to the output power.
-For windturbines and solar arrays these functions are well studied and approximations have been proposed.
-For the wind turbines, the power curve is thus specified by 4 values :
+For windturbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Dans2007}.
+For the wind turbines, the power curve can be specified by 4 values :
 
@@
 The most interesting part is the increase of output power when the wind speed is in the cut-in-speed rated-output-speed range.
-Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the litterature.
+Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the litterature \cite{Lydia2014}.
 
+\subsubsection{Solar power curve}
 The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data.
-As we mainly collected nebulosity series, we used the Helios model described in [].
+As we mainly collected nebulosity series, we used the Helios model described in \cite{Dans2007}.
 This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps.
-As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see [] for more details) :
+As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Dans2007} for more details) :
 
@@
 
-data, helios model, power curves, scema
+\subsubsection{Consumption}
+Modeling electric consumption has already been widely tackled in the litterature.
+Models can be basically divided in two main categories : Top-down and bottom-up approaches.
+Top-down techniques take aggregated consumption data as inputs and try to estimate individual consumption patterns while bottom-up methods use a fine modeling of users consumptions as to obtain realistic aggregated consumption curves.
+In this paper, we used a bottom-up model since the end user, or relatively small aggregations of end users, are in our interest.
+The main objective was to capture both daily patterns and seasonal variations of the consumptions.
+We assumed an additive model where the consumption of an agent is the sum of a seasonal heating term that depends on the outside temperature and an electronic consumption term that only depends on the hour of the day.
+By denoting $ \tau(t) $ the outside temperature at timestamp t, we can express the consumption $ P_{i}^{D}(t) $ of agent i at time t :
+
+\begin{equation}
+P_{i}^{D}(t) = \mathcal{F}_{i}^{heat}(\tau(t), t) + \mathcal{F}_{i}^{elec}(t)
+\end{equation}
+
+Where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
+In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification.
+By using termodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :
+
+\begin{equation}
+\mathcal{F}_{i}^{heat}(\tau(t), t) = \dfrac{B_{i}}{R_{i}} \left[ T_{i} - \tau(t) \right]
+\end{equation}
+
+Where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
+
+We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers.
+We also denote by $ \omega_{i}(t) = \{ \omega_{i}(t_{0}),...,\omega_{i}(t_{24}) \} $ the vector of the average fraction of $ \Omega_{i} $ used for each hour.
+We can therefore write :
+
+\begin{equation}
+\mathcal{F}_{i}^{elec}(t) = \Omega_{i} ( \omega_{i}(t) + \epsilon )
+\end{equation}
+
+Where $ \epsilon $ is a noise term.
+The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors.
+Business or residential areas for instance can be easily distinguised with this kind of model.
 

### 2015-07-22 Article.tex
abstract modif

@@
 \begin{abstract}
-Achieving a successful energetic transition became a major goal for the 21st century in most developped countries.
+%Achieving a successful energetic transition became a major goal for the 21st century in most developped countries.
 The reconciliation of several research areas and their recent focus on the same smart grid objective has led to major technical improvements.
@@
 
-In this paper, we thus consider agents that tend to produce more than they consume and seek to sell their surplus of production.
-More precisely, we rely on aggregations of these agents that can be seen as virtual power plants with production and consumption components that are both variable.
-Because these virtual power plants are complex aggregations of different types of production means and consumption habits, the volatilities of the plants productions are deeply impacted by the correlation between the agents.
-We propose here an algorithm that takes the underlying correlation structure of the agents into account while forming the coalitions.
+Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
+It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with a large penetration of small renewable generators and a proper efficient information system.
+All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
+In this paper, we study the formation of coalitions of these futuristic agents that both produce and consume.
+These coalitions, which can be seen as aggregators of the electricity market, aim at selling the aggregated surplus of production of the agents they control.
+We rely on real weather data sampled accross stations of a given territory in order to simulate realistic production and consumption patterns for each agent.
+This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
+As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
+We propose here an algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
 We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.

### 2015-07-23 Article.tex
some updates

@@
 In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
-More precisely, we consider a set of agents that can be seen as aggregators of renewable distributed energy ressources (DER) and electrical loads.
+More precisely, we consider a set of agents that own both renewable distributed energy ressources (DER) and electrical loads.
 The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
@@
 The intermittent nature of renewables such as wind or solar power introduces new challenges in control design.
-On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the natural ressource is present.
+On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the ressource is present.
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
@@
 
-An efficient portfolio of assets and a clustering of assets based on the correlations are obviously two very distinct things.
-And none of them is exactly what we seek.
-We will see in section X, how we combined both theories in order to form sufficiently producing coalitions with an acceptable volatility level.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 An essential component of the smart grid is the smart meter which makes the interface between the end user and the rest of the system.
-Smart meters coupled with sensors measure quantities of interest (like instantaneous consumption), receive informations from the grid (electricity price for instance), and take actions accordingly (demand side management program).
+Smart meters coupled with sensors measure quantities of interest (like instantaneous consumption), receive informations from the grid (electricity prices for instance), and take actions accordingly (demand side management programs).
 Smart meters are currently and gradually deployed, and will probably provide interesting datasets to work on.
 Unfortunately, at the time this paper was written, production and consumption data for prosumers over a large region were not yet available to our knowledge.
-Some interesting experiments are notwithstanding being conducted and data are progressively made public [].
+Some interesting experiments are notwithstanding being conducted and data are progressively made public \cite{ISSDA}.
 
@@
 Indeed, the larger the desired cliques, the more dense $ G_{2}^{\epsilon^{\star}} $ becomes (see equation \ref{epsilon_star}).
-There is a point where cliques results more from noisy edges than true de-correlation relationships, which decreases the quality of the results.
+There is a point where cliques results more from noisy edges than true de-correlation, which decreases the quality of the results.
 
@@
 \subsubsection{Wind power curve}
-Power curves are functions that, for a given type of generator, map the input to the output power.
+Power curves are functions that, for a given type of generator, map some input quantity to the output power produced.
 For windturbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Dans2007}.

### 2015-07-23 Article.tex
intro simplification

@@
 Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
-The primary objective of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
+A key point of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
 While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are linked to the quality and stability of the injections.
@@
 
-As beautiful as this distributed participatory production scenario seems to be, the constraints necessary for the electrical grid to remain in a stable state forbid any anarchic system where prosumers inject power without proper control.
-Actually the architecture and control of such a distributed grid creates numerous interdisciplinary challenges.
-In this paper, the envisaged solution for organizing such a multi-producers system consists in extending the kind of market environement discussed in the previous section.
-Participating aggregators estimate a production capacity for an upcoming period of time and announce it on the market.
-If a contract between the grid and a participant is bought, the latter commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
-
-In this paper, we consider that the production of a prosumer comes from renewables.
-Contracts are therefore based on predictions and come with a given amount of risk due to the intermittent nature of most renewables.
+More precisely, in our model, any entity that whishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
+If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
+Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
 Using storage as buffers in order to reduce riks is a popular approach.
@@
 Nevertheless, all coalitions are obviously not equally stable or productive, which means that special attention should be paid to the aggregation step.
-
-Furthermore, in our settings, since the prosumers have both a consumption and a production component, each depending on location and time, the diversification of the aggregations seems not trivial.
-This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
+Recall that prosumers have both a consumption and a production component, each depending on location and time, meaning that there are complex underlying correlations between the agents.
+This is indeed a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
 
@@
 We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market.
-Based on predictions, coalitions can decide wether or not to enter and what power quantities they are willing to provide.
-Of course, coalitions failing at fulfilling their obligations during the contracted period of time (because of bad forecasts, unexpected rare events, or intentional false information) will be exposed to financial penalities.
+We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
 
-Because agents are susceptible to fail for diverse reasons, the impact on the system's stability is critical.
-The propensy of a system to undertake these failures is often refered to as its resilience \cite{Pahwa}.
-Once formed, we also investigate in this paper the resilience of the coalitions when prosumers fail.
+Furthermore, because agents are susceptible to fail for diverse reasons, the propensy of a system to undertake these failures is critical \cite{Pahwa}.
+We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
 Despite the fact that losing agents is usually detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.

### 2015-07-27 Article.tex
related work updates

@@
 Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid.
-It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users.
+It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users \cite{Jiang2014}.
 The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market.
@@
 A key point of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
-While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are linked to the quality and stability of the injections.
+While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are related to the quality and stability of the injections.
 The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts.
-This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation relationships come into play.
+This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation come into play.
 
@@
 Recall that prosumers have both a consumption and a production component, each depending on location and time, meaning that there are complex underlying correlations between the agents.
-This is indeed a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
+This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
 
-We will see that variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
+We will see that the variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
 Understanding the correlation reliationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
@@
 
-Furthermore, because agents are susceptible to fail for diverse reasons, the propensy of a system to undertake these failures is critical \cite{Pahwa}.
+Furthermore, because agents are susceptible to fail for diverse reasons, the propensy of the system to undertake these failures is critical \cite{Pahwa}.
 We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
-Despite the fact that losing agents is usually detrimental to the coalitions, we will see that, with our algorithm, coalitions tend to be less impacted by random failures.
+Despite the fact that losing agents is usually detrimental to the coalitions, we will see that the coalitions formed with our algorithm tend to be less impacted by random failures.
 
@@
 
-The intermittent nature of renewables such as wind or solar power introduces new challenges in control design.
-On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the ressource is present.
+As explained in the section above, allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
+Indeed, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the ressource is present.
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
-A first approach to remedy this situation by acting on the consumption component is the use of dynamic ectricity prices sent to the end users.
-Together with demand side management techniques implemented on the smart meters, they consitute a first tool for shaping the consumption profiles \cite{Milligan2010}.
-In periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
+One possible solution consists in shifting some of the delayable loads through dynamic pricing.
+Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production spilled as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
+Schematically, in periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
 On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
 
-Another popular approach on the production side, is to combine renewable generators with storage devices, such that these are charged when there is a surplus of production, and discharged when the consumption exceeds the production.
-Although simple, this idea causes numerous challenges.
-Depending on the size of the system considered, centralized or decentralized control is desirable.
-In[], the authors introduce a distributed energy management system with a high penetration of renewables such that power is scheduled in a distributed fashion.
-In [] the optimal storage capacity problem is adressed.
+Dynamic pricing is an interesting and useful tool, but it has also some limitations.
+It is likely that dynamic pricing will serve as a shaping mass tool while finer techniques will be needed localy.
+A popular approach in this direction consists in deploying storage devices in the network and using them as electricity buffers.
+Basically these storages would be charged when there is a surplus of production, and discharged when the consumption exceeds the production.
+Although quite simple, this idea causes numerous challenges when it comes to its real implementation.
+As long as the system considered remains small, a centralized control of these buffers could be envisioned.
+However, for real large systems it is likely that more sophisticated decentralized algorithms will be necessary.
+In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high penetration of renewables such that power is scheduled in a distributed fashion.
+In \cite{Shadmand2014} the optimal storage capacity problem is adressed.
 There is indeed an interesting tradeoff between the costs of the equipments and the expected availability of power.
@@
 
-Although there are great progresses in stabilizing the production by demand side management or storage architectures, the need for good prediction techniques is more important than ever.
-Understanding the statistical properties of wind or solar irradiance are already well-established areas of research.
-There exists indeed a very interesting litterature on techniques aiming at predicting wind or solar irradiance profiles for future periods with error margins.
-In this paper, we adopt a slightly different approach : we suppose that the agents have historical records of their production and consumption, and we seek aggregations of agents with high expected production and low variability.
-
-Optimization of expected returns to risk is a traditional goal in finance.
-It is indeed well-known, that the more risk one is willing to take, the higher his potential gains.
+These are only a few possibilities for balancing production and demand in the smart grid.
+Most of the time, these techniques will be coupled with predictions of the upcomming load curves and weather conditions.
+Combining all these technologies enables aggregators to quantify their expected production and the inherent risk that comes with it.
+The optimization of expected returns to risk is a traditional goal in finance, and a vast litterature exists on this topic.
+It is well-known for instance, that the more risk one is willing to take, the higher his potential gains.
 On the contrary, when investing exclusively on low rik assets, one should expect relatively small gains.
@@
 If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
+One of the most controversial assumptions in the portfolio theory is that returns are jointly normaly distributed (or, at least, that the return distribution is jointly elliptical).
+Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{Chicheportiche2010}.
 
-This problem exhibits similarities with our objective of forming stable coalitions since our expected return can be formulated as the expected production and the risk / volatility as the variance of the production.
-However, it is assumed in the portfolio theory \cite{Markowitz1959} that returns are normaly distributed.
-In this paper, we chose to use normal distributions only for explanation purposes, or for estimating some parameters.
-Therefore, all the results are computed based on observed distributions (sometines far from gaussians).
-
-One of the key point in the Markowitz theory is to consider explicitely the correlation structure between the assets since these correlation relationships impact directly the variances of the portfolios.
-Since the work of \cite{Mantegna1999}, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph where the weight of an edge between two series (nodes) is the metric value for these series.
-
+Nevertheless, one of the key point in the Markowitz theory is to consider explicitely the correlation between the assets since they impact directly the variances of the portfolios.
+Since the work of \cite{Mantegna1999}, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph.
+Nodes represent the series considered while the edges are weighted by the metric.
 Because the metric can be computed for all pairs, these graphs are complete and of little use as is.
-Historically, the approach used by \cite{Mantegna1999} was to compute a minimum spanning tree over the correlation graph as to extract a correlation structure of the form of a hierarchical clustering.
+Historically, the approach used by \cite{Mantegna1999} was to compute a minimum spanning tree as to obtain a hierarchical clustering of the series.
 Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph.
-Another filtering approach by mean of a threshold $ \epsilon $ on the edge weights solves this problem and allows one to exhibit clusters of correlated series.
-In the following we will refer to these filtered graphs as $ \epsilon $-graph.
+In this paper, we use another classical filtering technique called $ \epsilon $-graph \cite{Garas2008}.
+It consists in selecting a threshold $ \epsilon $, and filtering out edges with smaller weights.
+As we will see further in this paper, this approach has the advantage of preserving clusters of correlated series.
 

### 2015-07-28 Article.tex
correction monique

@@
 
-More precisely, in our model, any entity that whishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
+More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
 If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
 Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
-Using storage as buffers in order to reduce riks is a popular approach.
+Using storage as buffers in order to reduce risks is a popular approach.
 Actually, a whole branch of the smart grid litterature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
@@
 Dynamic pricing is an interesting and useful tool, but it has also some limitations.
-It is likely that dynamic pricing will serve as a shaping mass tool while finer techniques will be needed localy.
+It is likely that dynamic pricing will serve as a shaping mass tool while finer techniques will be needed locally.
 A popular approach in this direction consists in deploying storage devices in the network and using them as electricity buffers.
@@
 Combining all these technologies enables aggregators to quantify their expected production and the inherent risk that comes with it.
-The optimization of expected returns to risk is a traditional goal in finance, and a vast litterature exists on this topic.
+The optimization of expected returns to risk is a traditional goal in finance, and a wide litterature exists on this topic.
 It is well-known for instance, that the more risk one is willing to take, the higher his potential gains.
-On the contrary, when investing exclusively on low rik assets, one should expect relatively small gains.
+On the contrary, when investing exclusively on low risks assets, one should expect relatively small gains.
 This tradeoff is formalized in the Markowitz' portfolio theory \cite{Markowitz1959}.
@@
 If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
-One of the most controversial assumptions in the portfolio theory is that returns are jointly normaly distributed (or, at least, that the return distribution is jointly elliptical).
+One of the most controversial assumptions in the portfolio theory is that returns are jointly normally distributed (or, at least, that the return distribution is jointly elliptical).
 Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{Chicheportiche2010}.

### 2015-07-28 Article.tex
update

@@
 All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
-In this paper, we study the formation of coalitions of these futuristic agents that both produce and consume.
-These coalitions, which can be seen as aggregators of the electricity market, aim at selling the aggregated surplus of production of the agents they control.
-We rely on real weather data sampled accross stations of a given territory in order to simulate realistic production and consumption patterns for each agent.
+In this paper, we propose an formation of coalitions algorithm of agents that both produce and consume.
+Our algorithm is designed to be used by an aggregator in order to form coalitions that aim at selling the aggregated surplus of production of the prosumer it controls.
+We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each agent.
 This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
 As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
-We propose here an algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
+We propose here a graph based algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
 We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
+
+%which can be seen as of the electricity market, .
 \end{abstract}
@@
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
+
+%%%%%%s
+Rather than maximizing the profit of each agents in the grid (prosumer and grid operator) our algorithm is trying to find the right trade-off between the productivity and volatility of the power injected into the grid by coalitions of prosumers.
+However, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
+%%%%%%
 
@@
 \begin{figure}
-\includegraphics[scale=.45]{Fig2.pdf}
+\includegraphics[scale=.45]{Fig2}
 \caption{Process diagram}
@@
 \begin{figure}
-\includegraphics[scale=.6]{gaussian_P_star.pdf}
+\includegraphics[scale=.6]{gaussian_P_star}
 \caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for gaussian distributions (see equation \ref{eq:P_star}).
@@
 \begin{figure}
-\includegraphics[scale=.6]{mean_field.pdf}
+\includegraphics[scale=.6]{mean_field}
 \caption{Gaussian mean approximation.
@@
 \begin{figure}
-\includegraphics[scale=.6]{cliques_histo.pdf}
+\includegraphics[scale=.6]{cliques_histo}
 \caption{Histograms of utility values for coalitions of size 3.
@@
 \begin{figure}
-\includegraphics[scale=.6]{real_utility_2.pdf}
+\includegraphics[scale=.6]{real_utility_2}
 \caption{Utility of random coalitions depending on their size.
@@
 \begin{figure}
-\includegraphics[scale=.6]{search.pdf}
+\includegraphics[scale=.6]{search}
 \caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
@@
 \begin{figure}
-\includegraphics[scale=.6]{./figures/coalitions/coalitions6.pdf}
+\includegraphics[scale=.6]{./figures/coalitions/coalitions6}
 \caption{Coalitions formed in the (contract value, volatility) space.
@@
 \begin{figure}
-\includegraphics[scale=.6]{./figures/resilience_both.pdf}
+\includegraphics[scale=.6]{./figures/resilience_both}
 \caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }

### 2015-07-29 Article.tex
updates

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Correlation among smart grid actors}
+\title{Stability of virtual aggregations of prosumers through diversification}
 
@@
 All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
-In this paper, we propose an formation of coalitions algorithm of agents that both produce and consume.
-Our algorithm is designed to be used by an aggregator in order to form coalitions that aim at selling the aggregated surplus of production of the prosumer it controls.
-We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each agent.
+In this paper, we propose an algorithm that forms coalitions of agents, called prosumers, that both produce and consume.
+It is designed to be used by aggregators that aim at selling the aggregated surplus of production of the prosumers they control.
+We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each prosumer.
 This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
 As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
-We propose here a graph based algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
+The proposed graph based algorithm of this paper takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
 We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
@@
 If one is larger than the other, the system deviates from its synchronous state.
-If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures.
+If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures \cite{Brummitt2012} \cite{Wang2009}.
 With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption.
@@
 It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users \cite{Jiang2014}.
-The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market.
+The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market \cite{Europe}.
 Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
@@
 These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...).
-Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition.
+Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition \cite{Europe}.
 The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.
@@
 
-%%%%%%s
-Rather than maximizing the profit of each agents in the grid (prosumer and grid operator) our algorithm is trying to find the right trade-off between the productivity and volatility of the power injected into the grid by coalitions of prosumers.
-However, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
-%%%%%%
+Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of risk acceptable.
+In other words, it tries to maximize coalitions profits without considering individual retributions to single agents (a task often studied through game theory).
+Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
 
-Furthermore, because agents are susceptible to fail for diverse reasons, the propensy of the system to undertake these failures is critical \cite{Pahwa}.
+Because agents are susceptible to fail for diverse reasons, the propensy of the system to undertake these failures is critical \cite{Pahwa}.
 We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
@@
 
-In the following, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
+In this paper, we use weather quantities like wind speed or solar irradiance as alternative data for generating realistic production and consumption series.
+Fortunately, these kinds of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
+Since these quantities depend both on time and location, we discretize time into slots and space into zones in the following.
+A zone is simply a portion of the considered region of study for which we sampled data.
+Therefore, if prosumers i and j are positioned on the same zone, they are exposed to the same weather.
+Adding some intra-zone noise can easily be done though not considered in this paper.
+
+More formally, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
 
@@
 More details about power curves and their approximations can be found in the appendix and in \cite{Lydia2014}.
-
-Weather quantities like wind speed or solar irradiance appear thus as alternative data for generating the $ P_{i} $ series.
-Fortunately, these kinds of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
-A quantity like wind speed depends both on time and location (we neglect height since this information is rarely present in the data).
-In the following, time is discretized into slots and space into zones.
-A zone is simply a portion of the considered space for which we have weather data.
-Therefore, if prosumers i and j are positioned on the same zone, they are exposed to the same weather.
-Adding some intra-zone noise can easily be done though not considered in this paper.
 The process for generating the $ P_{i} $ series is pictured in the first block of the process diagram (see figure \ref{fig:process}).

### 2015-07-29 Article.tex
updates

@@
 This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
-Nodes that were simply pure loads yesterday could behave tomorow as generators or loads.
+Nodes that were simply pure loads yesterday could behave tomorow as generators or loads \cite{Ramchurn}.
 On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
@@
 Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side, since an aggregator can be considered as a single node by the level above it.
+Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
 On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.

### 2015-07-29 Article.tex
update

@@
 \usepackage{array}
-
+\usepackage{color}
 \usepackage{float}
@@
 As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
-The proposed graph based algorithm of this paper takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
+The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
 We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
@@
 Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities.
-Even if most individual consumers may not realize it, the electicity prices are not constant rates and evolve with the production/consumption conditions on the grid.
+Even if most individual consumers may not realize it, the electricity prices are not constant rates and evolve with the production/consumption conditions on the grid.
 It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users \cite{Jiang2014}.
-The pricing of electricity and the necessity for the grid operator to have different kinds of emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market \cite{Europe}.
+The pricing of electricity and the necessity for the grid operator to have different emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market \cite{Europe}.
 Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
@@
 
-The smart grid vision however goes further and revolutionalizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
+The smart grid vision however goes further and revolutionizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
 This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
-Nodes that were simply pure loads yesterday could behave tomorow as generators or loads \cite{Ramchurn}.
+Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads \cite{Ramchurn}.
 On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
@@
 A key point of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
-While the latters intend to maximize their benefits with higer production contracts, the operator primary concerns are related to the quality and stability of the injections.
+While the latter intend to maximize their benefits with higher production contracts, the operator primary concerns are related to the quality and stability of the injections.
 The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts.
@@
 More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
-If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalities if it fails.
+If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
 Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
 Using storage as buffers in order to reduce risks is a popular approach.
-Actually, a whole branch of the smart grid litterature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
+Actually, a whole branch of the smart grid literature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
 Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable.
@@
 
-Forming coalitions/aggregations is an envisaged solution to both the number of entities sollicitating the operator, and their high variabilities.
+Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variabilities.
 Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
@@
 We will see that the variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
-Understanding the correlation reliationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
+Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
@@
 
-Because agents are susceptible to fail for diverse reasons, the propensy of the system to undertake these failures is critical \cite{Pahwa}.
+Because agents are susceptible to fail for diverse reasons, the propensity of the system to undertake these failures is critical \cite{Pahwa}.
 We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
@@
 
-The paper is organized as follows, section \ref{sec:related} gives a brief overview of the related litterature, section \ref{sec:data} clarifies how we generated realistic prosumer production traces based on weather data.
+The paper is organized as follows, section \ref{sec:related} gives a brief overview of the related literature, section \ref{sec:data} clarifies how we generated realistic prosumer production traces based on weather data.
 In section \ref{sec:notations}, we define most of the notations and explain why correlation between prosumers is a quantity of interest for our objective.
@@
 Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production spilled as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
-Schematically, in periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
+
+{\color{red} Schematically, in periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
 On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
+}
 
@@
 In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high penetration of renewables such that power is scheduled in a distributed fashion.
-In \cite{Shadmand2014} the optimal storage capacity problem is adressed.
+In \cite{Shadmand2014} the optimal storage capacity problem is addressed.
 There is indeed an interesting tradeoff between the costs of the equipments and the expected availability of power.
@@
 These are only a few possibilities for balancing production and demand in the smart grid.
-Most of the time, these techniques will be coupled with predictions of the upcomming load curves and weather conditions.
+Most of the time, these techniques will be coupled with predictions of the upcoming load curves and weather conditions.
 Combining all these technologies enables aggregators to quantify their expected production and the inherent risk that comes with it.
-The optimization of expected returns to risk is a traditional goal in finance, and a wide litterature exists on this topic.
+The optimization of expected returns to risk is a traditional goal in finance, and a wide literature exists on this topic.
 It is well-known for instance, that the more risk one is willing to take, the higher his potential gains.
@@
 More precisely, given a set of assets for which we have some historic data of returns, the objective is to find a linear combination of these assets (the so-called portfolio) which maximizes the expected value while minimizing the variance of the portfolio's return.
-Markowitz's answer is a set of efficient portolios that all optimize in some sense this tradeoff.
+Markowitz's answer is a set of efficient portfolios that all optimize in some sense this tradeoff.
 If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
@@
 In this paper, we use weather quantities like wind speed or solar irradiance as alternative data for generating realistic production and consumption series.
-Fortunately, these kinds of data are easier to find, and since the developement of small personal weather stations, their geographical granularity keeps increasing.
+Fortunately, these kinds of data are easier to find, and since the development of small personal weather stations, their geographical granularity keeps increasing.
 Since these quantities depend both on time and location, we discretize time into slots and space into zones in the following.
@@
 
-For simplicity, in this paper we only consider windturbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
+For simplicity, in this paper we only consider wind-turbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
 
@@
 Where $ \mathcal{F}_{WT} $ (resp.
-$ \mathcal{F}_{PV} $) is the power curve for the windturbines (resp.
+$ \mathcal{F}_{PV} $) is the power curve for the wind-turbines (resp.
 photovoltaic panels).
-We made here the implicit assumption that all windturbines (resp.
+We made here the implicit assumption that all wind-turbines (resp.
 photovoltaic panels) have the same power curve.
-The model can be easily extended to muplitiple power curves accounting for different types of generators.
+The model can be easily extended to multiple power curves accounting for different types of generators.
 More details about power curves and their approximations can be found in the appendix and in \cite{Lydia2014}.
@@
 Note that a prosumer i is defined by his zone $ Z_{i} $ as well as the sets $ DER_{i} $ and $ load_{i} $.
-That is, a prosumer can be configured to represent anything from a single windturbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) through more complex combinations.
+That is, a prosumer can be configured to represent anything from a single wind-turbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) through more complex combinations.
 In practice, we use random configurations for the agents.
 
-In the rest of the paper, we use french weather data \cite{Infoclimat} starting in january 2006 and ending in december 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
+In the rest of the paper, we use french weather data \cite{Infoclimat} starting in January 2006 and ending in December 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
 
@@
 Suppose now that coalition S has to suggest a production value $ P_{S}^{CRCT} $ to enter the market.
-This means that, during the time S is on the market, it will have to inject in the grid exactly $ P_{S}^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalities if it deviates.
+This means that, during the time S is on the market, it will have to inject in the grid exactly $ P_{S}^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalties if it deviates.
 Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittencies in the production and consumption.
-If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalities, but it is losing some gains since it could have annonced a higher contract value.
+If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalties, but it is losing some gains since it could have announced a higher contract value.
 If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques (see section \ref{sec:related}), S could be able to maintain its production to the contract value at any time.
-Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalities.
-We can see that there is a return over risk tradeoff here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalities.
+Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalties.
+We can see that there is a return over risk tradeoff here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
 
@@
 
-The amount of risks a given coalition is willing to take depends on a lot of things, among wich its capacity to compensate for under-producing (using batteries, backup generators...), or the risk aversion of the aggregator.
+The amount of risks a given coalition is willing to take depends on a lot of things, among which its capacity to compensate for under-producing (using batteries, backup generators...), or the risk aversion of the aggregator.
 Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works.
-In order to keep the present paper in a reasonable length, we simplify the contract value selection problem a little bit by giving some responsabilities to a third party named the grid operator.
+In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the grid operator.
 The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
@@
 For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units.
-Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line).
+Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horizontal dashed line).
 }
@@
 
-In this section, we use the notions of contract values and valid coalitions developped in section \ref{sec:notations} in order to design a proper utility function.
+In this section, we use the notions of contract values and valid coalitions developed in section \ref{sec:notations} in order to design a proper utility function.
 The contract basically indicates the rate at which a coalition has to inject power in the grid.
-It seems then natural that coalitions are remunerated proportionnally to their contract values $ gain(S) \propto P_{S}^{CRCT \star} $.
+It seems then natural that coalitions are remunerated proportionally to their contract values $ gain(S) \propto P_{S}^{CRCT \star} $.
 More precisely, if $ \lambda $ is the unitary price rate for electricity, a coalition S injecting $ P_{S}^{CRCT \star} $ in the grid during a period $ [t_{0},t_{k}] $ earns :
@@
 Using $ gain(S) $ directly as a utility function suffers a major drawback.
-It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterparty.
+It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
 
@@
 If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
-Since tunning the utility function with $ \alpha $ is not necessarily convenient, we relate $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $.
-If we wish that $ \mathcal{U}(S) $ tends to form coalitions of size approximately of the order of $ \bar{N} $, then :
+The utility function with $ \alpha $ is not necessarily convenient, here we relate $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $.
+otherwise $ \mathcal{U}(S) $ tends to form coalitions of size approximately in the order of $ \bar{N} $, then :
 
@@
 Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}).
-Subplot \textbf{b} dislays the corresponding utility functions for different values of $ \alpha $.
+Subplot \textbf{b} displays the corresponding utility functions for different values of $ \alpha $.
 Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
@@
 
-As can be pointed out, the purpose of $ \mathcal{U} $ is not a study of coalitions stability against player defection, which game theory provides a lot of tools for.
-This is indeed a problem on its own.
-The redistribution of a coalition's utility in terms of individual payoffs will therefore not be considered directly in this paper.
-$ \mathcal{U} $ can be rather interpreted as a measure of how good a given coalition is according to our criteria since it favors coalitions with good production to risk ratios.
+As can be pointed out, the purpose of $ \mathcal{U} $ is not to study of coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
+But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
+In other terms does a given coalitions has a good production to risk ratio.
 
@@
 Section \ref{sec:notations} explained how contract values for the coalitions are computed, and in section \ref{sec:utility} we related this quantity to the utility and gains of a coalition.
-As the computation time of this quantity is not negligable, and is increasing in the sizes of the aggregations, we explain in this section the heuristic we used to compute the coalitions.
+Since the computation time of this quantity is not negligible, we derive in this section the heuristic we used to form the coalitions structure.
 
@@
 Selecting the right filter $ \epsilon $ is thus an important point since it affects the landscape search for the coalition formation.
-Unfortunately, there seems to be no clear consensus in the litterature on how to select such a threshold.
+Unfortunately, there seems to be no clear consensus in the literature on how to select such a threshold.
 We will see later in this section that cliques in $ G_{2} $ are potential seeds for the coalitions.
@@
 Where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
-In other words we select $ \epsilon^{\star} $ as the smallest threshold posible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
+In other words we select $ \epsilon^{\star} $ as the smallest threshold possible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
 The existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
 The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
-\\
 
 \subsection{Cliques as de-correlated starting seeds}
-Correlation can be seen as cosine of angles in $ L^{2} $, hence even if there is no strict transivity relation for correlation, there is, to a certain extent, some partial notion of it.
+{\color{red}
+Correlation can be seen as cosine of angles in $ L^{2} $, hence even if there is no strict a transitivity relation for correlation, there is, to a certain extent, some partial notion of it.
 More precisely, if a, b, and c are three items such that $ \rho_{ab} > \delta $ and $ \rho_{bc} > \delta $, then we know, by the cosine addition formula\footnote{ $ cos(a+b) = cos(a)cos(b) - sin(a)sin(b) $ }, that $ \rho_{ac} > 2 \delta^{2} - 1 $.
-That is, if a and b are strongly correlated ($\rho_{ab} > 0.9 $) and b and c are also strongly correlated ($\rho_{bc} > 0.9 $), then there is a high probability for a and c of being strongly correlated ($\rho_{ac} > 0.62 $).
+That is, if a and b are strongly correlated ($\rho_{ab} > 0.9 $) and b and c are also strongly correlated ($\rho_{bc} > 0.9 $), then there is a high probability for a and c of being strongly correlated ($\rho_{ac} > 0.62 $).}
 
@@
 It is indeed possible that adding agents to these cliques has the combined effect of increasing the expected production while decreasing its stability.
-The question revolves around measuring the benefits of this production surplus compared to the disadvantage of a higher unstability.
+The question revolves around measuring the benefits of this production surplus compared to the disadvantage having higher instability.
 This can be quantified by the marginal benefit of equation \ref{eq:marginal_benefit}.
@@
 Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
-In this paper, neither the communication network nor the electrical grid are explicitely considered.
-Thus, we do not have any technical constraints on coalition sizes even if we designed the utitility such that these could be taken into account.
+In this paper, neither the communication network nor the electrical grid are explicitly considered.
+Thus, we do not have any technical constraints on coalition sizes even if we designed the utility such that these could be taken into account.
 We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
@@
 Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
-The transition from unvalid to valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
+The transition from invalid to valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
 After this transition, coalition's utilities improve slowly up to a maximum point.
@@
 Each small dot stands for the mean position of all sampled coalitions of this given size.
-Variances are not indicated for readalibility, but are usually quite large since this sampling only take the size as a constraint.
+Variances are not indicated for readability, but are usually quite large since this sampling only take the size as a constraint.
 Moreover, note that only a subset of the points are labeled with the size they represents as not to overload the figure.
@@
 Power curves are functions that, for a given type of generator, map some input quantity to the output power produced.
-For windturbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Dans2007}.
+For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Dans2007}.
 For the wind turbines, the power curve can be specified by 4 values :
@@
 The most interesting part is the increase of output power when the wind speed is in the cut-in-speed rated-output-speed range.
-Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the litterature \cite{Lydia2014}.
+Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the literature \cite{Lydia2014}.
 
@@
 \subsubsection{Consumption}
-Modeling electric consumption has already been widely tackled in the litterature.
+Modeling electric consumption has already been widely tackled in the literature.
 Models can be basically divided in two main categories : Top-down and bottom-up approaches.
@@
 The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors.
-Business or residential areas for instance can be easily distinguised with this kind of model.
+Business or residential areas for instance can be easily distinguished with this kind of model.
 

### 2015-07-29 Article.tex
update

@@
 It is indeed possible that adding agents to these cliques has the combined effect of increasing the expected production while decreasing its stability.
-The question revolves around measuring the benefits of this production surplus compared to the disadvantage having higher instability.
+The question revolves around measuring the benefits of this production surplus compared to the disadvantage of having coalition with high volatility.
 This can be quantified by the marginal benefit of equation \ref{eq:marginal_benefit}.
@@
 Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
-The transition from invalid to valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
+The transition from an invalid to a valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
 After this transition, coalition's utilities improve slowly up to a maximum point.

### 2015-07-30 Article.tex
red paragraphs modif

@@
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
-One possible solution consists in shifting some of the delayable loads through dynamic pricing.
+One possible solution consists in shifting some of the loads towards high productivity periods.
 Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production spilled as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
-
-{\color{red} Schematically, in periods when the production is expected to be greater than the consumption, prices are scheduled low in order to encourage end users to delay their elastic loads to these periods.
-On the contrary, when the production is expected to be less than the consumption, high prices are scheduled to deter any delayable loads over these periods.
-}
+Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing is believed to be a good way to give incentives to them.
+By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extend.
 
@@
 
-\subsection{Cliques as de-correlated starting seeds}
-{\color{red}
-Correlation can be seen as cosine of angles in $ L^{2} $, hence even if there is no strict a transitivity relation for correlation, there is, to a certain extent, some partial notion of it.
-More precisely, if a, b, and c are three items such that $ \rho_{ab} > \delta $ and $ \rho_{bc} > \delta $, then we know, by the cosine addition formula\footnote{ $ cos(a+b) = cos(a)cos(b) - sin(a)sin(b) $ }, that $ \rho_{ac} > 2 \delta^{2} - 1 $.
-That is, if a and b are strongly correlated ($\rho_{ab} > 0.9 $) and b and c are also strongly correlated ($\rho_{bc} > 0.9 $), then there is a high probability for a and c of being strongly correlated ($\rho_{ac} > 0.62 $).}
+\subsection{Cliques}
+
+In \cite{Garas2008} the structural roles of weak and strong links on financial correlation graphs is investigated.
+The author shows that strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters.
+Indeed, if we consider three items, say a, b, and c such that a and b are strongly correlated and b and c are also strongly correlated, then it is likely that a and c are also strongly correlated.
+It can be easily shown using the cosine addition formula\footnote{ $ cos(a+b) = cos(a)cos(b) - sin(a)sin(b) $ }, that if $\rho_{ab} > x $ and $\rho_{bc} > x $ with $ x>0 $, then $\rho_{ac} > 2x^{2}-1$).
+Correlation graphs capture this weak transitivity notion through clusters of correlated series.
 

### 2015-07-30 Article.tex
spell check

@@
 On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
-These plants completely rely on the presence and the intensity of their respective ressources (wind, sun, tiles...).
+These plants completely rely on the presence and the intensity of their respective resources (wind, sun, tiles...).
 Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition \cite{Europe}.
@@
 In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
-More precisely, we consider a set of agents that own both renewable distributed energy ressources (DER) and electrical loads.
+More precisely, we consider a set of agents that own both renewable distributed energy resources (DER) and electrical loads.
 The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
@@
 
-Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variabilities.
+Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variability.
 Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
@@
 As explained in the section above, allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
-Indeed, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the ressource is present.
+Indeed, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the resource is present.
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
@@
 In \cite{Shadmand2014} the optimal storage capacity problem is addressed.
-There is indeed an interesting tradeoff between the costs of the equipments and the expected availability of power.
+There is indeed an interesting trade-off between the costs of the equipments and the expected availability of power.
 The authors develop a framework that enables them to exhibit a Pareto front of efficient solutions.
@@
 On the contrary, when investing exclusively on low risks assets, one should expect relatively small gains.
-This tradeoff is formalized in the Markowitz' portfolio theory \cite{Markowitz1959}.
+This trade-off is formalized in the Markowitz' portfolio theory \cite{Markowitz1959}.
 More precisely, given a set of assets for which we have some historic data of returns, the objective is to find a linear combination of these assets (the so-called portfolio) which maximizes the expected value while minimizing the variance of the portfolio's return.
-Markowitz's answer is a set of efficient portfolios that all optimize in some sense this tradeoff.
+Markowitz's answer is a set of efficient portfolios that all optimize in some sense this trade-off.
 If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
@@
 
-Nevertheless, one of the key point in the Markowitz theory is to consider explicitely the correlation between the assets since they impact directly the variances of the portfolios.
-Since the work of \cite{Mantegna1999}, an interresting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph.
+Nevertheless, one of the key point in the Markowitz theory is to consider explicitly the correlation between the assets since they impact directly the variances of the portfolios.
+Since the work of \cite{Mantegna1999}, an interesting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph.
 Nodes represent the series considered while the edges are weighted by the metric.
@@
 
-In this paper, we use weather quantities like wind speed or solar irradiance as alternative data for generating realistic production and consumption series.
+In this paper, we use weather quantities like wind speed or solar radiance as alternative data for generating realistic production and consumption series.
 Fortunately, these kinds of data are easier to find, and since the development of small personal weather stations, their geographical granularity keeps increasing.
@@
 
-For a prosumer i, it is possible to write both quantities as a sum over the distributed energy ressources ($ DER_{i} $) and loads ($ load_{i} $) of i :
+For a prosumer i, it is possible to write both quantities as a sum over the distributed energy resources ($ DER_{i} $) and loads ($ load_{i} $) of i :
 
@@
 
-We denote by $ \nu_{i}(t) $ and $ \Psi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar irradiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
+We denote by $ \nu_{i}(t) $ and $ \Psi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar radiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
 \begin{equation}
@@
 This means that, during the time S is on the market, it will have to inject in the grid exactly $ P_{S}^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalties if it deviates.
-Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittencies in the production and consumption.
+Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittences in the production and consumption.
 If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalties, but it is losing some gains since it could have announced a higher contract value.
@@
 Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalties.
-We can see that there is a return over risk tradeoff here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
+We can see that there is a return over risk trade-off here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
 
@@
 If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
-Indeed, in this kind of situations, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalities to the grid.
-The probability that $ \{i,j\}$ is underproducing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's quality.
+Indeed, in this kind of situations, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalties to the grid.
+The probability that $ \{i,j\}$ is under-producing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's quality.
 A well-known result for normal distributions is that the cumulative distribution function can be written as :
@@
 The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
-In the gaussian example, this translates by coalition $ \{i,j\}$ announcing :
+In the Gaussian example, this translates by coalition $ \{i,j\}$ announcing :
 
@@
 \includegraphics[scale=.6]{gaussian_P_star}
-\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for gaussian distributions (see equation \ref{eq:P_star}).
+\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
 Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5.
@@
 
-The gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
+The Gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
 Nevertheless, such assumption is rather unrealistic in practice.
-In the following, we keep the same framework but release this gaussian assumption unless the contrary is specified.
+In the following, we keep the same framework but release this Gaussian assumption unless the contrary is specified.
 This assumption will indeed be convenient for computing some parameter estimates.
@@
 
-In order to get an estimator for $ \alpha $, we solve equation \ref{eq:derivative} in a gaussian case (as in section \ref{sec:notations}).
+In order to get an estimator for $ \alpha $, we solve equation \ref{eq:derivative} in a Gaussian case (as in section \ref{sec:notations}).
 Furthermore, since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
@@
 Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
-Using equation \ref{eq:alpha_star} to estimate $ \alpha $ based on the mean quantities and gaussian approximations seems to give acceptable results for the utility function behavior.
+Using equation \ref{eq:alpha_star} to estimate $ \alpha $ based on the mean quantities and Gaussian approximations seems to give acceptable results for the utility function behavior.
 
@@
 
-And we extend this measure to the coalition stuctures :
+And we extend this measure to the coalition structures :
 
@@
 Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolve according to $ \psi $.
-On the top subplot, $ P^{MIN} $ was volontarily selected relatively low such that the resiliences of the three structures fit on the same figure.
+On the top subplot, $ P^{MIN} $ was voluntarily selected relatively low such that the resiliences of the three structures fit on the same figure.
 When the $ P^{MIN} $ requirement increases, the differences between the algorithms also increase as visible on the bottom subplot of figure \ref{fig:resilience}.
@@
 
-Because in real situations, agents are prone to failure, resilience is also an important criterium for the quality of the aggregations.
+Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations.
 We therefore studied how the coalitions are able to remain on the market when their agents fail randomly one by one.
@@
 \item Nebulosity (integer in $[0,8] $)
-\item Temperature (in degree celsius)
+\item Temperature (in degree Celsius)
 \end{itemize}
@@
 In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification.
-By using termodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :
+By using thermodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :
 

### 2015-07-31 Article.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Stability of virtual aggregations of prosumers through diversification}
+\title{Stability And Performance Of Virtual Aggregations Of Prosumers Through Diversification In The Smart Grid}
 
@@
 One of the most controversial assumptions in the portfolio theory is that returns are jointly normally distributed (or, at least, that the return distribution is jointly elliptical).
-Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{Chicheportiche2010}.
+Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{doi:10.1142/S0219024912500197}.
 
@@
 Power curves are functions that, for a given type of generator, map some input quantity to the output power produced.
-For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Dans2007}.
+For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007, Piedallu2008}.
 For the wind turbines, the power curve can be specified by 4 values :
@@
 The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data.
-As we mainly collected nebulosity series, we used the Helios model described in \cite{Dans2007}.
+As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007, Piedallu2008}.
 This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps.
-As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Dans2007} for more details) :
+As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007, Piedallu2008} for more details) :
 

### 2015-07-31 Article.tex
update

@@
 
-%Understanding power grids seems like a reasonable starting point for studying so-called \textit{smart grids} that gained a lot of attention recently.
-As there exists many good articles \cite{Ramchurn} presenting smart grid ideas, concepts, or architectures, we will not dwell too much on this point.
-
 In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
 More precisely, we consider a set of agents that own both renewable distributed energy resources (DER) and electrical loads.
-The production of an agent can be, of course, used to meet its own demand, but in cases where it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
+The production of an agent can be, of course, used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
 Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
-A key point of this work is to mix the interests of the grid operator with the ones of the prosumers in a single utility function.
+A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
 While the latter intend to maximize their benefits with higher production contracts, the operator primary concerns are related to the quality and stability of the injections.
 The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts.
-This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation come into play.
+This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation append.
 
@@
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
-We build a framework in which the system operator has the possibility of specifying both a maximum amount of risk and a minimum production acceptable to enter the market.
+We build a framework in which the system operator specify both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to take.
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).

### 2015-07-31 Article.tex
update

@@
 One possible solution consists in shifting some of the loads towards high productivity periods.
-Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production spilled as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
+Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production is wasted as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
 Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing is believed to be a good way to give incentives to them.

### 2015-07-31 Article.tex
update

@@
 Unfortunately, at the time this paper was written, production and consumption data for prosumers over a large region were not yet available to our knowledge.
-Some interesting experiments are notwithstanding being conducted and data are progressively made public \cite{ISSDA}.
+Some interesting experiments are nonetheless being conducted and data are progressively made public \cite{ISSDA}.
 
@@
 
-For simplicity, in this paper we only consider wind-turbines (WT) and photovoltaic panels (PV) as possible DER for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
+For simplicity, in this paper we only consider wind-turbines (WT) and photovoltaic panels (PV) as possible DERs for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
 
@@
 \begin{figure}
-\includegraphics[scale=.45]{Fig2}
+\includegraphics[scale=.49]{Fig2}
 \caption{Process diagram}
@@
 $ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
-If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants where $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
+If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants when $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
 Indeed, in this kind of situations, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalties to the grid.
@@
 
-The amount of risks a given coalition is willing to take depends on a lot of things, among which its capacity to compensate for under-producing (using batteries, backup generators...), or the risk aversion of the aggregator.
+The contract a given coalition is willing to take depends on its capacity to compensate for under-producing (using batteries, backup generators...), and its risk acceptance.
 Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works.
@@
 The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
-In the Gaussian example, this translates by coalition $ \{i,j\}$ announcing :
+In the Gaussian example, it implies that coalition $ \{i,j\}$ is announcing :
 

### 2015-07-31 Article.tex
update

@@
 
-In order not to overload the market with unrealistically small coalitions, the grid operator also specify a lower bound $ P^{MIN} $ on the contract values.
+In order not to overload the market with unrealistically small coalitions, the grid operator also specifies a lower bound $ P^{MIN} $ on the contract values.
 We thus characterized a valid coalition as one satisfying the two conditions :
@@
 Nevertheless, such assumption is rather unrealistic in practice.
-In the following, we keep the same framework but release this Gaussian assumption unless the contrary is specified.
+In the following, we keep the same framework but release this Gaussian assumption unless the contrary is specified (see eq.
+\ref{eq:alpha_star}).
 This assumption will indeed be convenient for computing some parameter estimates.

### 2015-07-31 Article.tex
update

@@
 \begin{equation}
-\delta_{S}(i) \geq 0 \Leftrightarrow P_{S+\{i\}}^{CRCT \star} \geq P_{S}^{CRCT \star} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
+\delta_{S}(i) \geq 0 \ \ \ \Leftrightarrow \ \ \ P_{S+\{i\}}^{CRCT \star} \geq P_{S}^{CRCT \star} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
 \label{eq:marginal_benefit}
@@
 Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $.
-These nice curves are of course only valid in the simplified example considered here, but they will provide some guidance when using real data.
+These curves are only valid in the simplified Gaussian example considered here.
+Nonetheless, they will provide some guidance when using real data.
 
@@
 \end{figure}
-
-%Clearly, if $ \alpha = 0 $, whatever the size of the coalition, additional agents are added as long as they increase the contract value.
-Values of $\alpha $ greater than zero implies that additional agents must improve the contract value by a certain coefficient that decreases and converges to one as the size of the coalition is growing.
-For instance, if $\alpha = 1$ and $ |S| = 1 $, an additional agent must first double the contract value to be included, the next agent will then have to increase it by a factor of $ 1.5 $, and so on...
-Obviously, $ \alpha $ impacts directly the sizes of the coalitions formed.
-In the following, we consider the case where $ \alpha = 1 $.
 
@@
 \end{equation} 
-
-Where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
+where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
 In other words we select $ \epsilon^{\star} $ as the smallest threshold possible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
@@
 The question revolves around measuring the benefits of this production surplus compared to the disadvantage of having coalition with high volatility.
-This can be quantified by the marginal benefit of equation \ref{eq:marginal_benefit}.
+This can be quantified by the marginal benefit in equation \ref{eq:marginal_benefit}.
 

### 2015-07-31 Article.tex
update

@@
 \caption{Utility of random coalitions depending on their size.
-Blue dots shows real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
+Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
 On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
@@
 The bottom right corner, with high contract values and low volatilities, is therefore the region where we wish to form our coalitions.
-A single coalition is represented by a marker, and the color and shape of a marker indicates by which algorithm the coalition has been formed.
+A single coalition is represented by a marker and the color and shape of a marker indicates by which algorithm the coalition has been formed.
 Besides, the sizes of the coalitions are indicated on the markers, and the marker size is also proportional to the coalition size.

### 2015-07-31 Article.tex
update

@@
 \begin{abstract}
-%Achieving a successful energetic transition became a major goal for the 21st century in most developped countries.
-The reconciliation of several research areas and their recent focus on the same smart grid objective has led to major technical improvements.
-For instance, allowing a large penetration of renewables into the electricity production grid seemed more like a dream yesterday, but appears more and more doable as research progresses.
-The inherent volatility of some of these energy sources poses indeed various problems in such a constrained environment.
-It is often supposed that a large number of these small generators will be decentralized down to the distribution network and even be owned by individuals.
-By enabling bi-directional flows, this vision revolutionises the nature of today simple consumers into futuristic agents that can both consume and produce.
-Such agents can obviously reduce their electricity bills by seeking energetic independance through demand side management programs coupled with storage.
-Nevertheless, as it could be the case that an agent's production exceeds its consumption and storage capacity, its ability to sell its extra-production appears as a way to save energy.
-
 Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
@@
 We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
-
-%which can be seen as of the electricity market, .
 \end{abstract}
@@
 The model can be easily extended to multiple power curves accounting for different types of generators.
-More details about power curves and their approximations can be found in the appendix and in \cite{Lydia2014}.
+More details about power curves and their approximations can be found in the appendix \ref{appendixB} and in \cite{Lydia2014}.
 The process for generating the $ P_{i} $ series is pictured in the first block of the process diagram (see figure \ref{fig:process}).
@@
 Each small dot stands for the mean position of all sampled coalitions of this given size.
-Variances are not indicated for readability, but are usually quite large since this sampling only take the size as a constraint.
-Moreover, note that only a subset of the points are labeled with the size they represents as not to overload the figure.
+Variances are not indicated for readability, but are usually quite large since this sampling only takes the size as a constraint.
 We can see that as coalitions get larger, they tend to increase on average their contract values, but at the price of a higher volatility.
 
-On figure \ref{fig:coalitions}, the results of the random coalition structure sampling is shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
+On figure \ref{fig:coalitions}, the results of the random coalition structure sampling are shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
 Since the center of the ellipses stands for the mean and each ellipse adds one standard deviation, more than $ 99\% $ of the sampled gravity centers are within the largest ellipse.
@@
 The closer to red the better (high contract values with small volatility).
-On the opposite, blue areas shows poor quality (small contract values with high volatility).
+On the opposite, blue areas show poor quality (small contract values with high volatility).
 Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions.
-The smaller yellow markers stands for the gravity centers of the coalition structures.
-The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increase (a subset of the points are labeled by the size of the coalition they represent).
+The smaller yellow markers stand for the gravity centers of the coalition structures.
+The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increases (a subset of the points are labeled by the size of the coalition they represent).
 Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition.
-As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse add 1 standard deviation).
-For comparison with our algorithm where the utility function favors balanced coalitions, we also constrained loosely the sizes of the coalitions in the random process.
+As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).
+For instance in our algorithm the utility function favors balanced coalitions.
+We also loosely constrained the sizes of the coalitions in the utility function.
 }
@@
 We consider that prosumers fail randomly, and we denote by $ \psi \in [0,1] $ the fraction of agents that failed.
-Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolve according to $ \psi $.
+Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolves according to $ \psi $.
 On the top subplot, $ P^{MIN} $ was voluntarily selected relatively low such that the resiliences of the three structures fit on the same figure.
 When the $ P^{MIN} $ requirement increases, the differences between the algorithms also increase as visible on the bottom subplot of figure \ref{fig:resilience}.
-The de-correlated coalitions seem to achieve a more resilient production on the market in the sense that they are able to sustain a higher fraction of node failures.
+The decorrelated coalitions seem to achieve a more resilient production on the market in the sense that they are able to sustain a higher fraction of node failures.
 
@@
 We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
-Because they tend to have globally better production over volatility ratios, these coalitions will tend to use less storage and spill less energy than more unstable ones.
+Because they tend to have globally a better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions.
 We plan to study these benefits for the control of the aggregations in future works.
@@
 
-\subsection{Net production series}
+\subsection{Net production series}\label{appendixB}
 

### 2015-08-03 Article.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Stability And Performance Of Virtual Aggregations Of Prosumers Through Diversification In The Smart Grid}
+\title{Stability and performance of coalition of prosumers through diversification in the Smart Grid}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicolas Gensollen, Monique Becker, Vincent Gauthier and Michel Marot  \\
+\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot  \\
 \IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
@@
 
-%Designing stable power systems is a classical engineering challenge since blackout events can have catastrophic consequences.
-A recent approach introduced by complex systems theorists consists in abstracting the power grid as a graph where nodes represent loads, generators, or transformers, while edges stand for electrical lines \cite{Pagani2012} \cite{Rosas-Casals2007} \cite{Pahwa}.
-As graphs, power grids display some characteristics like clustering coefficients, degree distributions, or density for instance.
-These metrics enable scientists to get a better understanding of the structure of such complex systems \cite{Pagani2012}.
-It appears that power grids are particular graphs compared to well-known large networks like the internet, the world wide web, or social networks.
-That is, they exhibit different characteritics, suggesting that they may behave differently \cite{Rosas-Casals2007}.
-A common hypothesis for such a difference is that power grids are completely designed by a very small number of entities (electricity operators mainly), whereas other pre-cited graphs results from the interactions of billions of independant entities.
-Therefore, properties like power law degree distributions, prerential attachment, or hub oriented attacks do not necessarily apply with power grids.
-
-%Actually, drawing conclusions based only on the study of the underlying graph structure may lead to unrealistic results \cite{Pahwa}.
-Power grids are ideed dynamical systems governed by physical laws.
-Concepts such as shortest path or centrality metrics have thus to be extended to take into account the particular nature of power grids.
-Besides, stability and resilience of such systems are much more complex than the simple connectivity of the underlying graph \cite{Pahwa}.
-It is well-known for instance that frequency synchronization in the grid is a necessary condition for stability.
-A common solution consists in modeling the grid as a network of coupled oscillators governed by a swing equation dynamic \cite{Filatrella2008}.
-It can then be shown that the topology of the network impacts directly the capacity of the oscillators to reach the synchronized state \cite{Dorfler2013}.
-
 Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences.
@@
 While the latter intend to maximize their benefits with higher production contracts, the operator primary concerns are related to the quality and stability of the injections.
-The coalitions that we wish to form should thus be both stable and productive as to satisfy both parts.
-This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation append.
+The coalitions that we wish to form should thus be both stable and productive.
+This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation happen.
 
@@
 Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variability.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
+Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
 On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
@@
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
-We build a framework in which the system operator specify both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to take.
+We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to take.
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
@@
 One possible solution consists in shifting some of the loads towards high productivity periods.
-Coupled with demand side management techniques implemented on the end users' smart meters, this could result in less production is wasted as well as smaller consumption peaks \cite{Milligan2010} \cite{Logenthiran2012}.
+Demand side management techniques implemented in the end users' smart meters, can mitigate the consumption peaks and wast less production \cite{Milligan2010} \cite{Logenthiran2012}.
 Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing is believed to be a good way to give incentives to them.
@@
 If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
-One of the most controversial assumptions in the portfolio theory is that returns are jointly normally distributed (or, at least, that the return distribution is jointly elliptical).
+One of the most controversial assumptions in the portfolio theory is that returns are jointly normally distributed (or, at least, that the returns distribution is jointly elliptical).
 Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{doi:10.1142/S0219024912500197}.
@@
 \end{equation}
-
 $ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
@@
 \end{equation}
-
-Where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
+where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
 
@@
 \end{equation} 
-
 (since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
@@
 \end{equation}
-
-Where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
+where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
 $ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
@@
 Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
-Solving equation \ref{eq:derivative} for $ \alpha $ in these conditions yields :
+Solving equation \ref{eq:derivative} for $ \alpha $ in these conditions leads to:
 
@@
 
-As can be pointed out, the purpose of $ \mathcal{U} $ is not to study of coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
+As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
 But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
@@
 
-%Let $ P^{MARKET} = \sum_{S \in CS} P_{S}^{CRCT} $ be the production on the market, where CS is the coalition structure formed $ CS = \{S_{1},...S_{N_{COAL}} \} $, and $ N_{COAL} $ is the number of coalitions in the structure.
-First of all, $ P^{MARKET} $ depends explicitely on $ N_{COAL} $ such that, forming more coalitions induces an increase of the production.
-However, for a finite pool of agents, we expect an optimal number of coalitions $ N_{COAL}^{\star} $ that maximizes the market production.
-The reason for this expectation is that, for a given number of agents, more coalitions causes necessarily less agents in the coalitions, and therefore less diversity.
-We expect thus a point where coalitions formed start to fail the market entrance, and deteriorate the market production.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-08-03 Article.tex
update

@@
 Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
-It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with a large penetration of small renewable generators and a proper efficient information system.
+It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system.
 All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
@@
 These kind of settings are already being used at the transport level of current power grid with large renewable plants in the production portfolio.
-Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the penetration of renewables is important.
+Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the use of renewables is important.
 
@@
 Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads \cite{Ramchurn}.
-On the other hand, the penetration of renewables in the production is continuously growing, and is expected to become majority in a near future.
+On the other hand, the user of renewables in the production is continuously growing, and is expected to become majority in a near future.
 These plants completely rely on the presence and the intensity of their respective resources (wind, sun, tiles...).
@@
 However, for real large systems it is likely that more sophisticated decentralized algorithms will be necessary.
-In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high penetration of renewables such that power is scheduled in a distributed fashion.
+In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion.
 In \cite{Shadmand2014} the optimal storage capacity problem is addressed.

### 2015-08-03 Article.tex
update

@@
-\documentclass[conference]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Stability and performance of coalition of prosumers through diversification in the Smart Grid}
+\title{Stability and Performance of Coalition of Prosumers Through Diversification in the Smart Grid}
 
@@
 \end{equation}
-
-Where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
+where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
 
@@
 \end{equation}
-
-Where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
+where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
 The very simple form of this power curve is due to some simplifications in order not to overload the model.
@@
 \end{equation}
-
-Where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
+where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
 In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification.
@@
 \end{equation}
-
-Where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
+where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
 
@@
 \end{equation}
-
-Where $ \epsilon $ is a noise term.
+where $ \epsilon $ is a noise term.
 The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors.

### 2015-08-03 Article.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Stability and Performance of Coalition of Prosumers Through Diversification in the Smart Grid}
+\title{Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid}
 
@@
 Modeling electric consumption has already been widely tackled in the literature.
-Models can be basically divided in two main categories : Top-down and bottom-up approaches.
+Models can be basically divided into two main categories : Top-down and bottom-up approaches.
 Top-down techniques take aggregated consumption data as inputs and try to estimate individual consumption patterns while bottom-up methods use a fine modeling of users consumptions as to obtain realistic aggregated consumption curves.

### 2015-08-03 Article.tex
cut the...
cut

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot  \\
-\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
-Telecom SudParis/Institut Mines Telecom\\
-Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
+\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot}
 
@@
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
-
 Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of risk acceptable.
@@
 Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{doi:10.1142/S0219024912500197}.
-
 Nevertheless, one of the key point in the Markowitz theory is to consider explicitly the correlation between the assets since they impact directly the variances of the portfolios.
@@
 Some interesting experiments are nonetheless being conducted and data are progressively made public \cite{ISSDA}.
-
 In this paper, we use weather quantities like wind speed or solar radiance as alternative data for generating realistic production and consumption series.
@@
 Adding some intra-zone noise can easily be done though not considered in this paper.
-
 More formally, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
@@
 As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ P_{i}^{P} $ and $ P_{i}^{D} $.
-
 For a prosumer i, it is possible to write both quantities as a sum over the distributed energy resources ($ DER_{i} $) and loads ($ load_{i} $) of i :
@@
 More details about power curves and their approximations can be found in the appendix \ref{appendixB} and in \cite{Lydia2014}.
-The process for generating the $ P_{i} $ series is pictured in the first block of the process diagram (see figure \ref{fig:process}).
-
-\begin{figure}
-\includegraphics[scale=.49]{Fig2}
-\caption{Process diagram}
-\label{fig:process}
-\end{figure}
 
@@
 In practice, we use random configurations for the agents.
-
 In the rest of the paper, we use french weather data \cite{Infoclimat} starting in January 2006 and ending in December 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
@@
 As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts.
-
-In order not to overload the market with unrealistically small coalitions, the grid operator also specifies a lower bound $ P^{MIN} $ on the contract values.
+The grid operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions.
 We thus characterized a valid coalition as one satisfying the two conditions :
@@
 It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
-
 Such a model, that virtually allows infinitely large coalitions and contract values, is in practice not realistic.
@@
 $ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
-
 Based on $ \mathcal{U} $, the marginal contribution of an agent i can be expressed as $ \delta_{S}(i) = \mathcal{U}(S+\{i\}) - \mathcal{U}(S) $.
@@
 If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
-The utility function with $ \alpha $ is not necessarily convenient, here we relate $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $.
-otherwise $ \mathcal{U}(S) $ tends to form coalitions of size approximately in the order of $ \bar{N} $, then :
+We relate now $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $ by solving the following equation :
 
@@
 Nonetheless, they will provide some guidance when using real data.
+As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
+But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
 
@@
 \end{figure}
-
-As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
-But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
-In other terms does a given coalitions has a good production to risk ratio.
 
@@
 By using a metric to map this matrix in a weighted adjacency matrix (see section \ref{sec:related}), it is possible to obtain a graph representation of the correlation relationships between the agents.
-
 In the following, we use two opposite distance metrics for this mapping :
@@
 For any i and j, the weight of the edge $ e_{ij} $ is $ d_{ij}^{1} $ in $ G_{1} $ and $ d_{ij}^{2} $ in $ G_{2} $.
-
 In both cases, we want to keep only the edges which weights are located in the lower tail of the distance distributions.
 In other words, we want to compute the $ \epsilon$-graphs of $ G_{1} $ and $ G_{2} $ such that only meaningful edges remain.
-Selecting the right filter $ \epsilon $ is thus an important point since it affects the landscape search for the coalition formation.
+Selecting the filter $ \epsilon $ is an important point affecting the landscape search for the coalition formation.
 Unfortunately, there seems to be no clear consensus in the literature on how to select such a threshold.
@@
 Besides, since we consider coalitions as disjoint, the starting cliques should be non overlapping.
-We thus select our optimal threshold for $ G_{2} $ as :
+We select our optimal threshold for $ G_{2} $ as :
 
@@
 Searching for clusters in $ G_{2}^{\epsilon} $ and hoping that this strategy will provide a nice coalition structure of internally uncorrelated coalitions seems thus pointless.
-
 Consider now a clique in  $ G_{2}^{\epsilon} $, which is a complete subgraph of $ G_{2}^{\epsilon} $.
-This is indeed a structure of interest for our purpose.
 Since there is a link for every pairs of nodes, we know, by construction, that a clique has a mean correlation and a maximum correlation less than $ \epsilon $.
@@
 \begin{figure}
-\includegraphics[scale=.6]{cliques_histo}
+\includegraphics[scale=.5]{cliques_histo}
 \caption{Histograms of utility values for coalitions of size 3.
@@
 There is a point where cliques results more from noisy edges than true de-correlation, which decreases the quality of the results.
-
 Directly mapping cliques to coalitions by this de-correlation oriented approach is thus not sufficient.
@@
 Algorithm \ref{alg:algo2} uses such a sampling and returns the highest utility coalition structure sampled.
-This algorithm will be refered to as "\textit{random}" in the following.
-\item \textbf{Correlated :} This is the complete opposite of our algorithm.
-It uses the correlation graph $ G_{1} $ and performs a community detection.
+We refer to his algorithm as "\textit{random}" in the following.
+\item \textbf{Correlated :} Opposite version of our algorithm.
+It performs a community detection on the correlation graph $ G_{1} $.
 The resulting coalitions have thus very high internal correlations.
@@
 
-Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
-The transition from an invalid to a valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
-After this transition, coalition's utilities improve slowly up to a maximum point.
-
 \begin{figure}
-\includegraphics[scale=.6]{real_utility_2}
+\includegraphics[scale=.5]{real_utility_2}
 \caption{Utility of random coalitions depending on their size.
@@
 \label{fig:real_utility2}
-\end{figure}
-
-\begin{figure}
-\includegraphics[scale=.6]{search}
-\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
-\label{fig:search}
 \end{figure}
@@
 The coalitions of correlated agents (green squares) are clearly of poor quality according to our criteria since they can only afford small production contracts, and with a very high volatility.
-
-On figure \ref{fig:coalitions}, the decorrelated coalitions (blue dots) are closer to the bottom right corner indicating a much better quality in term of productivity over volatility ratio.
+The decorrelated coalitions (blue dots) are closer to the bottom right corner indicating a much better quality in term of productivity over volatility ratio.
 The black dotted line indicates the mean values for the random coalitions sampling technique.
@@
 We can see that as coalitions get larger, they tend to increase on average their contract values, but at the price of a higher volatility.
-
-On figure \ref{fig:coalitions}, the results of the random coalition structure sampling are shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
+The results of the random coalition structure sampling are shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
 Since the center of the ellipses stands for the mean and each ellipse adds one standard deviation, more than $ 99\% $ of the sampled gravity centers are within the largest ellipse.
@@
 \begin{figure}
-\includegraphics[scale=.6]{./figures/coalitions/coalitions6}
+\includegraphics[scale=.5]{./figures/coalitions/coalitions6}
 \caption{Coalitions formed in the (contract value, volatility) space.
@@
 Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition.
-As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).
-For instance in our algorithm the utility function favors balanced coalitions.
-We also loosely constrained the sizes of the coalitions in the utility function.
-}
+As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}
 \label{fig:coalitions}
@@
 A key point for the coalitions, besides stability and productivity, is their resilience.
-The resilience of a system can be roughly described as its ability to perform its tasks when subject to failures of its components.
+The resilience of a system refers to its ability to perform its tasks when subject to failures of its components.
 Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when node failures occur.
@@
 When the $ P^{MIN} $ requirement increases, the differences between the algorithms also increase as visible on the bottom subplot of figure \ref{fig:resilience}.
-The decorrelated coalitions seem to achieve a more resilient production on the market in the sense that they are able to sustain a higher fraction of node failures.
+The decorrelated coalitions achieve a more resilient production on the market in the sense that they sustain a higher fraction of node failures.
 
 \begin{figure}
-\includegraphics[scale=.6]{./figures/resilience_both}
+\includegraphics[scale=.5]{./figures/resilience_both}
 \caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
@@
 The power that a coalition is able to propose on the market is therefore related to production and stability.
-As the correlations between the prosumers that form these coalitions impact directly their volatilities, we seek uncorrelated aggregations of agents.
+The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents.
 We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation.
@@
 We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
-Because they tend to have globally a better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions.
-We plan to study these benefits for the control of the aggregations in future works.
-
+Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions.
 Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations.
-We therefore studied how the coalitions are able to remain on the market when their agents fail randomly one by one.
+We therefore studied how the coalitions remain on the market when their agents fail randomly.
 We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.
@@
 \end{algorithm}
+
+\begin{figure}
+\includegraphics[scale=.5]{search}
+\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
+\label{fig:search}
+\end{figure}
+
+Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
+The transition from an invalid to a valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
+After this transition, coalition's utilities improve slowly up to a maximum point.
 
@@
 
-Data were collected from \cite{Infoclimat} (similar data can be found at \cite{NCDC} for the United States). The variables used in the simulation are :
+Data were collected from \cite{Infoclimat} (data for the United States at \cite{NCDC}). The variables used in the simulation are :
 
@@
 where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
-
 Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :
@@
 where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
-
 We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers.

### 2015-08-27 Appendix.tex
drastic cut

--- 2015-08-03 Article.tex
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Title
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\title{Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Authors
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot}
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Abstract
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{abstract}
-Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
-It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system.
-All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
-In this paper, we propose an algorithm that forms coalitions of agents, called prosumers, that both produce and consume.
-It is designed to be used by aggregators that aim at selling the aggregated surplus of production of the prosumers they control.
-We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each prosumer.
-This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
-As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
-The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
-We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
-\end{abstract}
-
-\IEEEpeerreviewmaketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section I: Introduction
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-\label{sec:introduction}
-
-Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences.
-One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand.
-If one is larger than the other, the system deviates from its synchronous state.
-If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures \cite{Brummitt2012} \cite{Wang2009}.
-With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption.
-Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities.
-Even if most individual consumers may not realize it, the electricity prices are not constant rates and evolve with the production/consumption conditions on the grid.
-It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users \cite{Jiang2014}.
-The pricing of electricity and the necessity for the grid operator to have different emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market \cite{Europe}.
-Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
-These kind of settings are already being used at the transport level of current power grid with large renewable plants in the production portfolio.
-Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the use of renewables is important.
-
-The smart grid vision however goes further and revolutionizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
-This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
-Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads \cite{Ramchurn}.
-On the other hand, the user of renewables in the production is continuously growing, and is expected to become majority in a near future.
-These plants completely rely on the presence and the intensity of their respective resources (wind, sun, tiles...).
-Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition \cite{Europe}.
-The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.
-
-In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
-More precisely, we consider a set of agents that own both renewable distributed energy resources (DER) and electrical loads.
-The production of an agent can be, of course, used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
-Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
-A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
-While the latter intend to maximize their benefits with higher production contracts, the operator primary concerns are related to the quality and stability of the injections.
-The coalitions that we wish to form should thus be both stable and productive.
-This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation happen.
-
-More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
-If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
-Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
-Using storage as buffers in order to reduce risks is a popular approach.
-Actually, a whole branch of the smart grid literature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
-Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable.
-This is clearly unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.
-
-Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variability.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
-On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
-One thus expect a more stable and predictable energy production for a coalition than for a single agent.
-Nevertheless, all coalitions are obviously not equally stable or productive, which means that special attention should be paid to the aggregation step.
-Recall that prosumers have both a consumption and a production component, each depending on location and time, meaning that there are complex underlying correlations between the agents.
-This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
-
-We will see that the variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
-Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
-More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
-We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to take.
-We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
-Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of risk acceptable.
-In other words, it tries to maximize coalitions profits without considering individual retributions to single agents (a task often studied through game theory).
-Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
-
-Because agents are susceptible to fail for diverse reasons, the propensity of the system to undertake these failures is critical \cite{Pahwa}.
-We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
-Despite the fact that losing agents is usually detrimental to the coalitions, we will see that the coalitions formed with our algorithm tend to be less impacted by random failures.
-
-The paper is organized as follows, section \ref{sec:related} gives a brief overview of the related literature, section \ref{sec:data} clarifies how we generated realistic prosumer production traces based on weather data.
-In section \ref{sec:notations}, we define most of the notations and explain why correlation between prosumers is a quantity of interest for our objective.
-Based on the conclusions of section \ref{sec:notations}, we develop in section \ref{sec:utility} a utility function quantifying how much power a coalition can announce on the market given an accepted risk level.
-We then develop, in section \ref{sec:forming} a greedy optimization algorithm that uses decorrelated cliques as inputs and improves their utility over a correlation-constrained environment.
-Finally, section \ref{sec:results} provides some results both on performance of the method and resilience of the coalitions formed.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section II: Related Work
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Related Work}
-\label{sec:related}
-
-As explained in the section above, allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
-Indeed, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the resource is present.
-Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
-One possible solution consists in shifting some of the loads towards high productivity periods.
-Demand side management techniques implemented in the end users' smart meters, can mitigate the consumption peaks and wast less production \cite{Milligan2010} \cite{Logenthiran2012}.
-Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing is believed to be a good way to give incentives to them.
-By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extend.
-
-Dynamic pricing is an interesting and useful tool, but it has also some limitations.
-It is likely that dynamic pricing will serve as a shaping mass tool while finer techniques will be needed locally.
-A popular approach in this direction consists in deploying storage devices in the network and using them as electricity buffers.
-Basically these storages would be charged when there is a surplus of production, and discharged when the consumption exceeds the production.
-Although quite simple, this idea causes numerous challenges when it comes to its real implementation.
-As long as the system considered remains small, a centralized control of these buffers could be envisioned.
-However, for real large systems it is likely that more sophisticated decentralized algorithms will be necessary.
-In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion.
-In \cite{Shadmand2014} the optimal storage capacity problem is addressed.
-There is indeed an interesting trade-off between the costs of the equipments and the expected availability of power.
-The authors develop a framework that enables them to exhibit a Pareto front of efficient solutions.
-
-These are only a few possibilities for balancing production and demand in the smart grid.
-Most of the time, these techniques will be coupled with predictions of the upcoming load curves and weather conditions.
-Combining all these technologies enables aggregators to quantify their expected production and the inherent risk that comes with it.
-The optimization of expected returns to risk is a traditional goal in finance, and a wide literature exists on this topic.
-It is well-known for instance, that the more risk one is willing to take, the higher his potential gains.
-On the contrary, when investing exclusively on low risks assets, one should expect relatively small gains.
-This trade-off is formalized in the Markowitz' portfolio theory \cite{Markowitz1959}.
-More precisely, given a set of assets for which we have some historic data of returns, the objective is to find a linear combination of these assets (the so-called portfolio) which maximizes the expected value while minimizing the variance of the portfolio's return.
-Markowitz's answer is a set of efficient portfolios that all optimize in some sense this trade-off.
-If one is able to put a number on his risk acceptance or on the target expected return, the corresponding efficient portfolio is a priori the best option.
-One of the most controversial assumptions in the portfolio theory is that returns are jointly normally distributed (or, at least, that the returns distribution is jointly elliptical).
-Some economist have pointed out the fact that this assumption might not capture well the reality of financial markets \cite{doi:10.1142/S0219024912500197}.
-Nevertheless, one of the key point in the Markowitz theory is to consider explicitly the correlation between the assets since they impact directly the variances of the portfolios.
-Since the work of \cite{Mantegna1999}, an interesting approach consists in computing a distance metric based on the correlation coefficients in order to organize the series in a correlation graph.
-Nodes represent the series considered while the edges are weighted by the metric.
-Because the metric can be computed for all pairs, these graphs are complete and of little use as is.
-Historically, the approach used by \cite{Mantegna1999} was to compute a minimum spanning tree as to obtain a hierarchical clustering of the series.
-Later on, it was pointed out that, by definition, a spanning tree could not capture the underlying clustering structure hidden in the correlation graph.
-In this paper, we use another classical filtering technique called $ \epsilon $-graph \cite{Garas2008}.
-It consists in selecting a threshold $ \epsilon $, and filtering out edges with smaller weights.
-As we will see further in this paper, this approach has the advantage of preserving clusters of correlated series.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section III: Generating realistic prosumer patterns
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Generating realistic prosumer patterns}
-\label{sec:data}
-
-An essential component of the smart grid is the smart meter which makes the interface between the end user and the rest of the system.
-Smart meters coupled with sensors measure quantities of interest (like instantaneous consumption), receive informations from the grid (electricity prices for instance), and take actions accordingly (demand side management programs).
-Smart meters are currently and gradually deployed, and will probably provide interesting datasets to work on.
-Unfortunately, at the time this paper was written, production and consumption data for prosumers over a large region were not yet available to our knowledge.
-Some interesting experiments are nonetheless being conducted and data are progressively made public \cite{ISSDA}.
-In this paper, we use weather quantities like wind speed or solar radiance as alternative data for generating realistic production and consumption series.
-Fortunately, these kinds of data are easier to find, and since the development of small personal weather stations, their geographical granularity keeps increasing.
-Since these quantities depend both on time and location, we discretize time into slots and space into zones in the following.
-A zone is simply a portion of the considered region of study for which we sampled data.
-Therefore, if prosumers i and j are positioned on the same zone, they are exposed to the same weather.
-Adding some intra-zone noise can easily be done though not considered in this paper.
-More formally, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
-
-\begin{equation}
-P_{i}(t) = P_{i}^{P}(t) - P_{i}^{D}(t)
-\end{equation}
-
-Where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
-In other words, $ P_{i}(t) $ represents the instantaneous surplus of power that agent i is willing to sell at time t.
-As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ P_{i}^{P} $ and $ P_{i}^{D} $.
-For a prosumer i, it is possible to write both quantities as a sum over the distributed energy resources ($ DER_{i} $) and loads ($ load_{i} $) of i :
-
-\begin{equation}
-P_{i}^{P}(t) = \sum_{k \in DER_{i}} P_{k}(t)
-\end{equation}
-\begin{equation}
-P_{i}^{D}(t) = \sum_{k \in load_{i}} P_{k}(t)
-\end{equation}
-
-For simplicity, in this paper we only consider wind-turbines (WT) and photovoltaic panels (PV) as possible DERs for the agents ($ DER_{i} = WT_{i} \cup PV_{i} $):
-
-\begin{equation}
-P_{i}^{D}(t) = \sum_{k \in WT_{i}} P_{k}(t) + \sum_{k \in PV_{i}} P_{k}(t)
-\end{equation}
-
-We denote by $ \nu_{i}(t) $ and $ \Psi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar radiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
-\begin{equation}
- P_{i}^{P}(t) = \sum_{k \in WT_{i}} \mathcal{F}_{WT}( \nu_{i}(t) ) + \sum_{k \in PV_{i} } \mathcal{F}_{PV}(\Psi_{i}(t) ) 
-\end{equation}
-Where $ \mathcal{F}_{WT} $ (resp.
-$ \mathcal{F}_{PV} $) is the power curve for the wind-turbines (resp.
-photovoltaic panels).
-We made here the implicit assumption that all wind-turbines (resp.
-photovoltaic panels) have the same power curve.
-The model can be easily extended to multiple power curves accounting for different types of generators.
-More details about power curves and their approximations can be found in the appendix \ref{appendixB} and in \cite{Lydia2014}.
-
-Note that a prosumer i is defined by his zone $ Z_{i} $ as well as the sets $ DER_{i} $ and $ load_{i} $.
-That is, a prosumer can be configured to represent anything from a single wind-turbine for instance ($ DER_{i} = \{ WT_{0} \} $ and $ load_{i} = \emptyset $) to a pure load ($ DER_{i} = \emptyset $ and $ load_{i} = \{ L_{0} \} $) through more complex combinations.
-In practice, we use random configurations for the agents.
-In the rest of the paper, we use french weather data \cite{Infoclimat} starting in January 2006 and ending in December 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section III: Model
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Notations}
-\label{sec:notations}
-
-This section provides most of the notations and introduces important concepts for the rest of the paper.
-As explained in section \ref{sec:data}, we consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
-Based on these historical values, our objective is now to form groups of prosumers (the so-called coalitions) so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
-Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the extra-production of coalition S at time t.
-
-Suppose now that coalition S has to suggest a production value $ P_{S}^{CRCT} $ to enter the market.
-This means that, during the time S is on the market, it will have to inject in the grid exactly $ P_{S}^{CRCT} $ at any time t and will be rewarded proportionally to this amount, with penalties if it deviates.
-Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittences in the production and consumption.
-If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalties, but it is losing some gains since it could have announced a higher contract value.
-If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques (see section \ref{sec:related}), S could be able to maintain its production to the contract value at any time.
-Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalties.
-We can see that there is a return over risk trade-off here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
-
-Let us illustrate the rest of the notations and concepts with a simple example.
-We consider only two agents i and j such that the distribution of their extra-production can be approximated by normal distributions : $ P_{i} \sim \mathcal{N}(\mu_{i}, \sigma_{i} ) $ and $ P_{j} \sim \mathcal{N}(\mu_{j}, \sigma_{j} ) $.
-This is only for explanation purposes as it is of course rather unrealistic in real situations where the distributions are skewed.
-Using simple statistics, we can write the distribution of the coalition $ S = \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-		\mu_{ij} = \mu_{i} + \mu_{j} \\
-		\sigma_{ij} = \sqrt{\sigma_{i}^{2} + \sigma_{j}^{2} + \rho_{ij} \sigma_{i} \sigma_{j} }
-\end{array} \right.
-\end{equation}
-$ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
-If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants when $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
-Indeed, in this kind of situations, $ \{i,j\}$ will either have to discharge batteries to keep up with its contract, or pay penalties to the grid.
-The probability that $ \{i,j\}$ is under-producing compared to the contract : $ Pr[P_{i,j} \leq P^{CRCT}] $ is thus an important indicator of the coalition's quality.
-A well-known result for normal distributions is that the cumulative distribution function can be written as :
-\begin{equation}
-Pr[P_{ij} \leq P_{S}^{CRCT}] = \dfrac{1}{2} \left[ 1+ erf \left( \dfrac{P_{S}^{CRCT} - \mu_{ij}}{\sigma_{ij}\sqrt{2}} \right) \right] 
-\end{equation}
-where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
-
-The contract a given coalition is willing to take depends on its capacity to compensate for under-producing (using batteries, backup generators...), and its risk acceptance.
-Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works.
-In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the grid operator.
-The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
-More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the grid operator as a maximum value for the probability of under-producing.
-The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
-In the Gaussian example, it implies that coalition $ \{i,j\}$ is announcing :
-
-\begin{equation}
-\label{eq:P_star}
-P_{S}^{CRCT \star} = \mu_{ij} - \sqrt{2} \sigma_{ij} erf^{-1}( 1 - 2 \phi )
-\end{equation}
-
-\begin{figure}
-\includegraphics[scale=.6]{gaussian_P_star}
-\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
-Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5.
-Under a grid policy of $ \phi = 0.1 $, it is able to announce a contract value of $ P_{S}^{CRCT \star} = 4.36 $.
-The same coalition in term of expected production ($\mu = 5 $), but with a higher variance ($ \sigma = 1.5 $, cyan curve with circles) can only afford a smaller contract value of $ P_{S}^{CRCT \star} = 3.07 $.
-The red curve with diamonds stands for a coalition with a higher expected production $(\mu = 7)$, but with a very high unpredictability ($ \sigma = 5 $).
-For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units.
-Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horizontal dashed line).
-}
-\label{fig:Gaussian}
-\end{figure}
-
-This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator.
-Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $.
-For illustration, the range of $ \phi $ values is shown from 0 to 1, but in practice, only small values of $ \phi $ really make sense : $ \phi = 1 $ for instance means that coalitions can announce absolutely anything since the probability of producing less than any contract value is necessarily less than one by trivial definition of a probability.
-As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts.
-The grid operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions.
-We thus characterized a valid coalition as one satisfying the two conditions :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-			Pr[P_{ij} \leq P_{S}^{CRCT}] \leq \phi \\
-			P_{S}^{CRCT} \geq P^{MIN}
-\end{array} \right.
-\end{equation}
-
-On figure \ref{fig:Gaussian}, $ P^{MIN} $ is fixed to 2 units for illustration purpose.
-For $ \phi = 0.1 $, only blue triangles and cyan circles coalitions are valid while red diamonds coalition is not.
-
-The Gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
-Nevertheless, such assumption is rather unrealistic in practice.
-In the following, we keep the same framework but release this Gaussian assumption unless the contrary is specified (see eq.
-\ref{eq:alpha_star}).
-This assumption will indeed be convenient for computing some parameter estimates.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section V: Utility Function
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Utility Function}
-\label{sec:utility}
-
-In this section, we use the notions of contract values and valid coalitions developed in section \ref{sec:notations} in order to design a proper utility function.
-The contract basically indicates the rate at which a coalition has to inject power in the grid.
-It seems then natural that coalitions are remunerated proportionally to their contract values $ gain(S) \propto P_{S}^{CRCT \star} $.
-More precisely, if $ \lambda $ is the unitary price rate for electricity, a coalition S injecting $ P_{S}^{CRCT \star} $ in the grid during a period $ [t_{0},t_{k}] $ earns :
-
-\begin{equation}
-gain(S) = \int_{t_{0}}^{t_{k}} \lambda P_{S}^{CRCT \star} dt = P_{S}^{CRCT \star} \int_{t_{0}}^{t_{k}} \lambda dt
-\end{equation} 
-(since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
-Using $ gain(S) $ directly as a utility function suffers a major drawback.
-It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
-Such a model, that virtually allows infinitely large coalitions and contract values, is in practice not realistic.
-There are indeed costs (communication costs for instance) that increase with the coalitions sizes.
-We take this observation into account by rescaling the utility of a coalition S by its size in term of number of agents ($|S|$):
-
-\begin{equation}
-\mathcal{U}(S) = \left\{ \begin{array}{lll}
-							\dfrac{1}{|S|^{\alpha}} \dfrac{ P_{S}^{CRCT \star} }{P^{MAX}},\ if\ S\ is\ valid, \\
-							0,\ if\ S\ is\ not\ valid
-						 \end{array}
-				  \right.
-\end{equation}
-where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
-$ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
-Based on $ \mathcal{U} $, the marginal contribution of an agent i can be expressed as $ \delta_{S}(i) = \mathcal{U}(S+\{i\}) - \mathcal{U}(S) $.
-A coalition S has thus an interest in adding an additional agent i if this marginal contribution is positive :
-
-\begin{equation}
-\delta_{S}(i) \geq 0 \ \ \ \Leftrightarrow \ \ \ P_{S+\{i\}}^{CRCT \star} \geq P_{S}^{CRCT \star} \left( \dfrac{|S|+1}{|S|} \right)^{\alpha}
-\label{eq:marginal_benefit}
-\end{equation}
-
-If $ \alpha $ is set to zero, agents are added as long as they increase the contract value of the coalition.
-If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
-We relate now $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $ by solving the following equation :
-
-\begin{equation}
-\left[ \dfrac{\partial{ U}}{ \partial{|S|}} \right]_{|S| = \bar{N}} = 0
-\label{eq:derivative}
-\end{equation}
-
-In order to get an estimator for $ \alpha $, we solve equation \ref{eq:derivative} in a Gaussian case (as in section \ref{sec:notations}).
-Furthermore, since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
-Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
-Solving equation \ref{eq:derivative} for $ \alpha $ in these conditions leads to:
-
-\footnotesize
-\begin{equation}
-\alpha^{\star}_{\bar{N}} = \dfrac{0.7 \bar{\sigma}(\bar{\rho}-1)erf^{-1}(2 \phi - 1)}{\bar{\mu}\sqrt{\bar{N}(\bar{\rho}\bar{N}-\bar{\rho}+1)}+1.4 \bar{\sigma} erf^{-1} (2 \phi -1 ) (\bar{\rho}\bar{N}-\bar{\rho} + 1)}
-\label{eq:alpha_star}
-\end{equation}
-\normalsize
-
-Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $.
-These curves are only valid in the simplified Gaussian example considered here.
-Nonetheless, they will provide some guidance when using real data.
-As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
-But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
-
-\begin{figure}
-\includegraphics[scale=.6]{mean_field}
-\caption{Gaussian mean approximation.
-Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}).
-Subplot \textbf{b} displays the corresponding utility functions for different values of $ \alpha $.
-Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
-Finally, the red curve with squares has an optimal size of 15 agents.
-}
-\label{fig:mean_approx}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section VI: Coalition Formation
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Coalition Formation}
-\label{sec:forming}
-Section \ref{sec:notations} explained how contract values for the coalitions are computed, and in section \ref{sec:utility} we related this quantity to the utility and gains of a coalition.
-Since the computation time of this quantity is not negligible, we derive in this section the heuristic we used to form the coalitions structure.
-
-\subsection{Representing the correlation structure}
-As seen in section \ref{sec:notations}, the variance of the aggregated production impacts directly the contract values, and depends on the covariances between the agents productions.
-We argue here that, by having some representation of the correlation structure between the agents, the search landscape for high utility coalitions could be reduced, such that good coalitions are more likely to be found quickly.
-Usually, this correlation structure is formalized with a covariance matrix or a correlation matrix that contains all the correlation coefficients between the agents : $ M = (\rho_{ij})_{\forall i,j \in \mathcal{A}^{2}}$.
-By using a metric to map this matrix in a weighted adjacency matrix (see section \ref{sec:related}), it is possible to obtain a graph representation of the correlation relationships between the agents.
-In the following, we use two opposite distance metrics for this mapping :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-			d_{ij}^{1} = 1 - \rho_{ij}^{2}, \\
-			d_{ij}^{2} = \rho_{ij}^{2} = 1 - d_{ij}^{1}
-\end{array} \right.
-\end{equation}
-
-Clearly, $ d^{1} $ (resp.
-$ d^{2} $) maps two correlated series as close points (resp.
-distant) while two uncorrelated series are distant (resp.
-close).
-These metrics enable us to compute a correlation graph $ G_{1} = (\mathcal{A}, E_{1}) $ and a "de-correlation" graph $ G_{2} = (\mathcal{A}, E_{2} ) $.
-For any i and j, the weight of the edge $ e_{ij} $ is $ d_{ij}^{1} $ in $ G_{1} $ and $ d_{ij}^{2} $ in $ G_{2} $.
-In both cases, we want to keep only the edges which weights are located in the lower tail of the distance distributions.
-In other words, we want to compute the $ \epsilon$-graphs of $ G_{1} $ and $ G_{2} $ such that only meaningful edges remain.
-Selecting the filter $ \epsilon $ is an important point affecting the landscape search for the coalition formation.
-Unfortunately, there seems to be no clear consensus in the literature on how to select such a threshold.
-We will see later in this section that cliques in $ G_{2} $ are potential seeds for the coalitions.
-Since we want to generate $ N_{COAL} $ coalitions, we need at least $ N_{COAL} $ cliques of a given size to start.
-Besides, since we consider coalitions as disjoint, the starting cliques should be non overlapping.
-We select our optimal threshold for $ G_{2} $ as :
-
-\begin{equation}
-\label{epsilon_star}
-\epsilon^{\star} = min_{ \epsilon \in [0,1]} \left\{ \epsilon\ s.t.\ |\Theta_{k}(G_{2}^{\epsilon})| \geq N_{COAL} \right\}
-\end{equation} 
-where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
-In other words we select $ \epsilon^{\star} $ as the smallest threshold possible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
-The existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
-The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
-
-\subsection{Cliques}
-
-In \cite{Garas2008} the structural roles of weak and strong links on financial correlation graphs is investigated.
-The author shows that strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters.
-Indeed, if we consider three items, say a, b, and c such that a and b are strongly correlated and b and c are also strongly correlated, then it is likely that a and c are also strongly correlated.
-It can be easily shown using the cosine addition formula\footnote{ $ cos(a+b) = cos(a)cos(b) - sin(a)sin(b) $ }, that if $\rho_{ab} > x $ and $\rho_{bc} > x $ with $ x>0 $, then $\rho_{ac} > 2x^{2}-1$).
-Correlation graphs capture this weak transitivity notion through clusters of correlated series.
-
-Nevertheless de-correlation seems like a more complex concept than correlation in the sense that there is not even a partial notion of transitivity when it comes to it.
-Therefore, the clustering coefficients of $ G_{1}^{\epsilon} $ is much higher than the one of $ G_{2}^{\epsilon} $.
-This can be seen as another formulation of \cite{Garas2008} on the structural roles of weak and strong links on financial correlation graphs.
-Strong links, accounting for strong correlation relationships, are responsible for the clustering, while weak links provide the connectivity between clusters.
-Searching for clusters in $ G_{2}^{\epsilon} $ and hoping that this strategy will provide a nice coalition structure of internally uncorrelated coalitions seems thus pointless.
-Consider now a clique in  $ G_{2}^{\epsilon} $, which is a complete subgraph of $ G_{2}^{\epsilon} $.
-Since there is a link for every pairs of nodes, we know, by construction, that a clique has a mean correlation and a maximum correlation less than $ \epsilon $.
-
-\begin{figure}
-\includegraphics[scale=.5]{cliques_histo}
-\caption{Histograms of utility values for coalitions of size 3.
-Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets.
-As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
-\label{fig:histo_cliques}
-\end{figure}
-
-Figure \ref{fig:histo_cliques} shows the distributions of the utility values for cliques of size 3 (triangles) in $ G_{2}^{\epsilon^{\star}} $ and for all the other possible triplets of agents.
-It is clearly visible that cliques tend to exhibit higher utilities because of their de-correlation property.
-Choosing cliques in $ G_{2}^{\epsilon^{\star}} $ as coalitions seems therefore appealing.
-Nevertheless, the quality of the results seems to decrease as the sizes of the cliques increase.
-Indeed, the larger the desired cliques, the more dense $ G_{2}^{\epsilon^{\star}} $ becomes (see equation \ref{epsilon_star}).
-There is a point where cliques results more from noisy edges than true de-correlation, which decreases the quality of the results.
-Directly mapping cliques to coalitions by this de-correlation oriented approach is thus not sufficient.
-It is indeed possible that adding agents to these cliques has the combined effect of increasing the expected production while decreasing its stability.
-The question revolves around measuring the benefits of this production surplus compared to the disadvantage of having coalition with high volatility.
-This can be quantified by the marginal benefit in equation \ref{eq:marginal_benefit}.
-
-\subsection{Algorithm}
-The algorithm takes inputs from :
-
-\begin{itemize}
-	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
-	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
-	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
-\end{itemize}
-
-The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
-Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
-The next step is a local greedy improvement over the landscape represented by  $ G_{2}^{\epsilon^{\star}} $.
-Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
-This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
-The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
-See the details in algorithm \ref{alg:algo1} in the appendix.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section VII: Results
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-
-The algorithm presented in the previous section is supposed to generate a given number of coalitions that have good utilities.
-As it comprises mainly of a greedy optimization based on local improvements, there is no guarantee that the algorithm finds the global optimum.
-Since there is, to our knowledge, no state of the art algorithm that aggregates uncorrelated agents in an optimum way (see section \ref{sec:related} for related problems), we compare the results with :
-
-\begin{itemize}
-\item \textbf{Random sampling of coalitions :} Coalitions are formed randomly without any other constraint that the desired size.
-This enables us to have an idea about the distributions of utility values for coalitions of a given size.
-\item \textbf{Random sampling of coalition structures :} Coalition structures are sampled randomly by shuffling and random divisions of the agents.
-Algorithm \ref{alg:algo2} uses such a sampling and returns the highest utility coalition structure sampled.
-We refer to his algorithm as "\textit{random}" in the following.
-\item \textbf{Correlated :} Opposite version of our algorithm.
-It performs a community detection on the correlation graph $ G_{1} $.
-The resulting coalitions have thus very high internal correlations.
-We thus expect this algorithm to perform very bad compared to the others.
-See algorithm \ref{alg:algo3}.
-\end{itemize}
-
-Before running the algorithms, we need to calibrate the utility function by choosing the value of the $ \alpha $ parameter.
-Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
-In this paper, neither the communication network nor the electrical grid are explicitly considered.
-Thus, we do not have any technical constraints on coalition sizes even if we designed the utility such that these could be taken into account.
-We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
-Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
-Using equation \ref{eq:alpha_star} to estimate $ \alpha $ based on the mean quantities and Gaussian approximations seems to give acceptable results for the utility function behavior.
-
-\begin{figure}
-\includegraphics[scale=.5]{real_utility_2}
-\caption{Utility of random coalitions depending on their size.
-Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
-On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
-\label{fig:real_utility2}
-\end{figure}
-
-Figure \ref{fig:coalitions} shows the coalitions formed with the considered algorithms in the contract value / volatility space.
-The color map in the background indicates regions where we expect high utilities (red) and the ones where we expect very poor utility values (blue).
-The bottom right corner, with high contract values and low volatilities, is therefore the region where we wish to form our coalitions.
-A single coalition is represented by a marker and the color and shape of a marker indicates by which algorithm the coalition has been formed.
-Besides, the sizes of the coalitions are indicated on the markers, and the marker size is also proportional to the coalition size.
-We can see that the utility function results in approximatively balanced coalitions.
-Small yellow markers indicates the gravity centers of their respective coalition structures.
-The coalitions of correlated agents (green squares) are clearly of poor quality according to our criteria since they can only afford small production contracts, and with a very high volatility.
-The decorrelated coalitions (blue dots) are closer to the bottom right corner indicating a much better quality in term of productivity over volatility ratio.
-The black dotted line indicates the mean values for the random coalitions sampling technique.
-Each small dot stands for the mean position of all sampled coalitions of this given size.
-Variances are not indicated for readability, but are usually quite large since this sampling only takes the size as a constraint.
-We can see that as coalitions get larger, they tend to increase on average their contract values, but at the price of a higher volatility.
-The results of the random coalition structure sampling are shown with the red ellipses that represent the distribution of the gravity centers of the sampled structures.
-Since the center of the ellipses stands for the mean and each ellipse adds one standard deviation, more than $ 99\% $ of the sampled gravity centers are within the largest ellipse.
-The small yellow dot below the ellipses indicates the gravity center of our solution.
-It is thus visible that our greedy graph based algorithm is able to find a quite good coalition structure in terms of volatility and contract values.
-
-\begin{figure}
-\includegraphics[scale=.5]{./figures/coalitions/coalitions6}
-\caption{Coalitions formed in the (contract value, volatility) space.
-The color map indicates qualities of portions of the plane.
-The closer to red the better (high contract values with small volatility).
-On the opposite, blue areas show poor quality (small contract values with high volatility).
-Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions.
-The smaller yellow markers stand for the gravity centers of the coalition structures.
-The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increases (a subset of the points are labeled by the size of the coalition they represent).
-Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition.
-As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}
-\label{fig:coalitions}
-\end{figure}
-
-A key point for the coalitions, besides stability and productivity, is their resilience.
-The resilience of a system refers to its ability to perform its tasks when subject to failures of its components.
-Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when node failures occur.
-According to our model, the grid operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $.
-As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market.
-We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold :
-
-\begin{equation}
-\mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}]
-\end{equation}
-
-And we extend this measure to the coalition structures :
-
-\begin{equation}
-\mathcal{R}_{CS} = \prod_{S \in CS} \left( 1 - Pr[ P_{S} < P^{MIN} ] \right)
-\label{eq:resilience}
-\end{equation}
-
-We consider that prosumers fail randomly, and we denote by $ \psi \in [0,1] $ the fraction of agents that failed.
-Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolves according to $ \psi $.
-On the top subplot, $ P^{MIN} $ was voluntarily selected relatively low such that the resiliences of the three structures fit on the same figure.
-When the $ P^{MIN} $ requirement increases, the differences between the algorithms also increase as visible on the bottom subplot of figure \ref{fig:resilience}.
-The decorrelated coalitions achieve a more resilient production on the market in the sense that they sustain a higher fraction of node failures.
-
-\begin{figure}
-\includegraphics[scale=.5]{./figures/resilience_both}
-\caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
-\label{fig:resilience}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Section VI: Conclusion
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-\label{sec:conclusion}
-
-In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator.
-By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions.
-The power that a coalition is able to propose on the market is therefore related to production and stability.
-The correlations between the prosumers impact directly the volatility of the coalitions, which led us to seek uncorrelated aggregations of agents.
-We used a graph representation of the correlation relationships between the agents as a reduced landscape for the coalition formation.
-A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results.
-We compare these results with random samplings, and an opposite strategy that clusters correlated agents together.
-We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.
-Because of their better production over volatility ratios, these coalitions will tend to use less storage and waste less energy than more unstable coalitions.
-Because in real situations, agents are prone to failure, resilience is also an important criterion for the quality of the aggregations.
-We therefore studied how the coalitions remain on the market when their agents fail randomly.
-We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Bibliography
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{IEEEtran}  
-\bibliography{Article}
-
 \appendix
@@
 
-\begin{figure}
+\begin{figure}[b]
 \includegraphics[scale=.5]{search}
@@
 
+\bibliographystyle{IEEEtran}  
+\bibliography{Article}
 \end{document}

### 2015-08-27 Article.tex
drastic cut

@@
 One possible solution consists in shifting some of the loads towards high productivity periods.
-Demand side management techniques implemented in the end users' smart meters, can mitigate the consumption peaks and wast less production \cite{Milligan2010} \cite{Logenthiran2012}.
-Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing is believed to be a good way to give incentives to them.
+Demand side management techniques can indeed mitigate the consumption peaks and wast less production \cite{Milligan2010} \cite{Logenthiran2012}.
+Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing appears as a good way to give incentives to them.
 By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extend.
@@
 In other words, $ P_{i}(t) $ represents the instantaneous surplus of power that agent i is willing to sell at time t.
-As explained above, since large datasets containing this quantity over time are not yet available, we simulated these traces by considering separately $ P_{i}^{P} $ and $ P_{i}^{D} $.
+We simulated these traces by considering separately $ P_{i}^{P} $ and $ P_{i}^{D} $.
 For a prosumer i, it is possible to write both quantities as a sum over the distributed energy resources ($ DER_{i} $) and loads ($ load_{i} $) of i :
@@
 This section provides most of the notations and introduces important concepts for the rest of the paper.
-As explained in section \ref{sec:data}, we consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
-Based on these historical values, our objective is now to form groups of prosumers (the so-called coalitions) so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
+We consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
+Based on these historical values, our objective is now to form groups of prosumers so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
 Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the extra-production of coalition S at time t.
@@
 If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalties, but it is losing some gains since it could have announced a higher contract value.
-If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques (see section \ref{sec:related}), S could be able to maintain its production to the contract value at any time.
+If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques, S could be able to maintain its production to the contract value at any time.
 Nevertheless, if the oscillations are too important compared to the available storage capacity, S will probably break the contract and pay penalties.
-We can see that there is a return over risk trade-off here, meaning that coalitions should find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
+We can see that there is a return over risk trade-off here.
+Coalitions should thus find the right balance between announcing too low and losing some potential gains, and claiming too high and paying penalties.
 
@@
 This is only for explanation purposes as it is of course rather unrealistic in real situations where the distributions are skewed.
-Using simple statistics, we can write the distribution of the coalition $ S = \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :
+We can write the distribution of the coalition $ S = \{i,j\} $ as $ P_{\{i,j\}} \sim \mathcal{N}(\mu_{ij}, \sigma_{ij}) $, where :
 
@@
 \begin{figure}
-\includegraphics[scale=.6]{gaussian_P_star}
-\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
+\includegraphics[scale=.48]{gaussian_P_star}
+\caption{{\footnotesize $ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
 Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5.
@@
 For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units.
-Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horizontal dashed line).
-}
+Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horizontal dashed line).} }
 \label{fig:Gaussian}
@@
 (since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
-Using $ gain(S) $ directly as a utility function suffers a major drawback.
-It is indeed not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
+Unfortunately $ gain(S) $ is not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
 Such a model, that virtually allows infinitely large coalitions and contract values, is in practice not realistic.
@@
 where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
-$ P^{MAX} $ can be seen as the maximum production which can be injected in the grid.
 Based on $ \mathcal{U} $, the marginal contribution of an agent i can be expressed as $ \delta_{S}(i) = \mathcal{U}(S+\{i\}) - \mathcal{U}(S) $.
-A coalition S has thus an interest in adding an additional agent i if this marginal contribution is positive :
+A coalition S has thus an interest in adding an additional agent i if :
 
@@
 If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
-We relate now $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $ by solving the following equation :
+We relate now $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $ by solving the following equation in the gaussian case (as in section \ref{sec:notations}) :
 
@@
 
-In order to get an estimator for $ \alpha $, we solve equation \ref{eq:derivative} in a Gaussian case (as in section \ref{sec:notations}).
 Furthermore, since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
@@
 As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
-But we aim to design $ \mathcal{U} $ as a measure of how good a given coalition is according to our criteria.
+$ \mathcal{U} $ is rather designed as a quality measure for the coalitions.
 
 \begin{figure}
-\includegraphics[scale=.6]{mean_field}
-\caption{Gaussian mean approximation.
-Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}).
+\includegraphics[scale=.48]{mean_field}
+\caption{{\footnotesize Gaussian mean approximation.
+Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean size of the coalitions (see equation \ref{eq:alpha_star}).
 Subplot \textbf{b} displays the corresponding utility functions for different values of $ \alpha $.
-Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
-Finally, the red curve with squares has an optimal size of 15 agents.
-}
+Blue curve with diamonds favors small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions.
+Finally, the red curve with squares has an optimal size of 15 agents.} }
 \label{fig:mean_approx}
@@
 By using a metric to map this matrix in a weighted adjacency matrix (see section \ref{sec:related}), it is possible to obtain a graph representation of the correlation relationships between the agents.
-In the following, we use two opposite distance metrics for this mapping :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-			d_{ij}^{1} = 1 - \rho_{ij}^{2}, \\
-			d_{ij}^{2} = \rho_{ij}^{2} = 1 - d_{ij}^{1}
-\end{array} \right.
-\end{equation}
+In the following, we use two opposite distance metrics $  d_{ij}^{1} = 1 - \rho_{ij}^{2} $ and $ d_{ij}^{2} = \rho_{ij}^{2} = 1 - d_{ij}^{1} $.
 
@@
 \begin{figure}
-\includegraphics[scale=.5]{cliques_histo}
-\caption{Histograms of utility values for coalitions of size 3.
-Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets.
-As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
+\includegraphics[scale=.48]{cliques_histo}
+\caption{{\footnotesize Histograms of utility values for coalitions of size 3.
+Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triplets.
+Cliques tend to exhibit higher utilities than random coalitions.}}
 \label{fig:histo_cliques}
@@
 Cliques add alternatively the node $ i^{\star} $ in their neighborhood that yields the best marginal benefit $ MAX_{ i \in N(clique) } \delta_{clique}(i) $ where $ N(clique) $ is the neighborhood of a given clique.
-This addition occurs only if $ i^{\star} $ is not already involved in another coalition, and if $ \delta_{clique}(i^{\star}) \geq 0 $, meaning that utilities are increasing.
-The algorithm stops when all nodes are distributed in a coalition or when the global utility stops increasing.
-See the details in algorithm \ref{alg:algo1} in the appendix.
+See algorithms section in the appendix.
 
@@
 \item \textbf{Random sampling of coalition structures :} Coalition structures are sampled randomly by shuffling and random divisions of the agents.
-Algorithm \ref{alg:algo2} uses such a sampling and returns the highest utility coalition structure sampled.
-We refer to his algorithm as "\textit{random}" in the following.
+This algorithm (see appendices) uses such a sampling and returns the highest utility coalition structure sampled.
+We refer to this algorithm as "\textit{random}".
 \item \textbf{Correlated :} Opposite version of our algorithm.
@@
 We thus expect this algorithm to perform very bad compared to the others.
-See algorithm \ref{alg:algo3}.
+See appendices.
 \end{itemize}
@@
 Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
-In this paper, neither the communication network nor the electrical grid are explicitly considered.
-Thus, we do not have any technical constraints on coalition sizes even if we designed the utility such that these could be taken into account.
+In this paper, we do not have any technical constraints on coalition sizes even if we designed the utility such that these could be taken into account.
 We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
 Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
-Using equation \ref{eq:alpha_star} to estimate $ \alpha $ based on the mean quantities and Gaussian approximations seems to give acceptable results for the utility function behavior.
-
 \begin{figure}
-\includegraphics[scale=.5]{real_utility_2}
-\caption{Utility of random coalitions depending on their size.
+\includegraphics[scale=.48]{real_utility_2}
+\caption{{\footnotesize Utility of random coalitions depending on their size.
 Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
-On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
+On this plot the $ \alpha $ parameter of the utility function was selected according to eq.
+\ref{eq:alpha_star} in order to favor 40 agents coalitions.}}
 \label{fig:real_utility2}
@@
 \begin{figure}
-\includegraphics[scale=.5]{./figures/coalitions/coalitions6}
-\caption{Coalitions formed in the (contract value, volatility) space.
-The color map indicates qualities of portions of the plane.
+\includegraphics[scale=.48]{./figures/coalitions/coalitions6}
+\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space.
+The color map indicates the underlying quality.
 The closer to red the better (high contract values with small volatility).
@@
 Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition.
-As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}
+We also show the distribution of the gravity centers of random coalition structures with red ellipses (center is the mean, and each ellipse corresponds to one  standard deviation).}}
 \label{fig:coalitions}
@@
 As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market.
-We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold :
-
-\begin{equation}
-\mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}]
-\end{equation}
-
+We define the resilience of a coalition S as the probability that S produces more than the $ P^{MIN} $ threshold : $ \mathcal{R}_{S} = Pr[P_{S} >= P^{MIN}] = 1 - Pr[P_{S} < P^{MIN}] $.
 And we extend this measure to the coalition structures :
@@
 \begin{figure}
-\includegraphics[scale=.5]{./figures/resilience_both}
-\caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
+\includegraphics[scale=.48]{./figures/resilience_both}
+\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }
 \label{fig:resilience}
@@
 
-\appendix
-\subsection{Algorithms}
-
-\begin{algorithm}
- \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- Compute $ G_{2}^{\epsilon^{\star}} $ \;
- Find the $ N_{COAL} $ cliques in $ G_{2}^{\epsilon^{\star}} $\;
- \While{$ \mathcal{U}(CS) $ is improving}{
- 	\For{each clique}{
- 		Find $ i^{\star} $ \;
- 		\If{ $ \delta_{clique}(i^{\star}) \geq 0 $ }{
- 			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
- 			}
- 		\If{ $ \exists j \in clique,\ s.t\ \delta_{clique}(j) < 0 $}{
- 			$ clique \leftarrow clique - \{j \} $ \;
- 			}
-   		}
-  	}
- \caption{Local greedy optimization algorithm}
- \label{alg:algo1}
-\end{algorithm}
-
-\begin{figure}
-\includegraphics[scale=.5]{search}
-\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
-\label{fig:search}
-\end{figure}
-
-Figure \ref{fig:search} displays the evolution of the global utility and the number of involved agents during the course of the greedy algorithm \ref{alg:algo1}.
-The transition from an invalid to a valid coalitions is clearly visible on the blue diamond curve and occurs between iteration 10 and 15.
-After this transition, coalition's utilities improve slowly up to a maximum point.
-
-\begin{algorithm}
- \KwData{Agent set $\mathcal{A}$,\\ Desired number of coalitions $ N_{COAL} $, \\ Maximum number of iterations $ Loop^{max} $}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- $ N^{loop} \leftarrow 0 $ \;
- $ CS^{\star} \leftarrow \emptyset $\;
- \While{$ N^{loop} < Loop^{max} $}{
- 	$ CS \leftarrow SelectRandomCS( ) $\;
- 	\If{ $ \mathcal{U}(CS) > \mathcal{U}(CS^{\star}) $ }{
- 		$ CS^{\star } \leftarrow CS $\;
- 		}
- 	$ N^{loop} \leftarrow  N^{loop} + 1 $\;
-  	}
-  return $ CS^{\star} $
- \caption{Random algorithm}
- \label{alg:algo2}
-\end{algorithm}
-
-\begin{algorithm}
- \KwData{$P_{i}$ series,\\ Desired number of coalitions $ N_{COAL} $,\\ search step size $ \beta << 1 $}
- \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
- $ \epsilon \leftarrow 1 $ \;
- $ CS \leftarrow \emptyset $\;
- \While{$ |CS| < N_{COAL} $}{
- 	Compute $ G_{1}^{\epsilon} $ \;
- 	$ CS \leftarrow computeClusters( G_{1}^{\epsilon} ) $\;
- 	\If{ $ |CS| = N_{COAL} $ }{
- 		return CS\;
- 		}
- 	\Else{ 
- 	$ \epsilon \leftarrow \epsilon - \beta $\;
- 	}
-  	}
- \caption{Correlated algorithm}
-\label{alg:algo3}
-\end{algorithm}
-
-\subsection{Net production series}\label{appendixB}
-
-Data were collected from \cite{Infoclimat} (data for the United States at \cite{NCDC}). The variables used in the simulation are :
-
-\begin{itemize}
-\item Average wind speed (in $m.s^{-1}$)
-\item Nebulosity (integer in $[0,8] $)
-\item Temperature (in degree Celsius)
-\end{itemize}
-
-\subsubsection{Wind power curve}
-Power curves are functions that, for a given type of generator, map some input quantity to the output power produced.
-For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007, Piedallu2008}.
-For the wind turbines, the power curve can be specified by 4 values :
-
-\begin{itemize}
-\item Cut-in-speed : The wind speed at which the turbine first starts to rotate and generates power.
-\item Rated-output-power : The maximum power that the turbine can generate.
-\item Rated-output-speed : The wind speed at which the turbine attains its rated output power.
-\item Cut-out-speed : The speed at which the turbine is turned off as not to damage the rotor.
-\end{itemize}
-
-The most interesting part is the increase of output power when the wind speed is in the cut-in-speed rated-output-speed range.
-Even if sometimes a simple linear model is used, the increase has been shown to be non linear and some more complex exponential fit can be found in the literature \cite{Lydia2014}.
-
-\subsubsection{Solar power curve}
-The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data.
-As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007, Piedallu2008}.
-This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps.
-As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007, Piedallu2008} for more details) :
-
-\begin{equation}
-\left\{ \begin{array}{lll}
-			\Psi{real}(t) = \Psi_{perfect}(t) \eta(t) \\
-			\eta(t) = 1-0.75 \left( \dfrac{N(t)}{8} \right)^{3.4}
-\end{array} \right.
-\end{equation}
-where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
-Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :
-
-\begin{equation}
-\mathcal{F}_{PV}(\Psi{real}(t)) =  S_{PV} \Psi{real}(t)  e_{PV}
-\end{equation}
-where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
-The very simple form of this power curve is due to some simplifications in order not to overload the model.
-For instance, it does not take into account angles and orientations degradations.
-These could be incorporated if needed by changing the power curve in the simulations.
-
-\subsubsection{Consumption}
-Modeling electric consumption has already been widely tackled in the literature.
-Models can be basically divided into two main categories : Top-down and bottom-up approaches.
-Top-down techniques take aggregated consumption data as inputs and try to estimate individual consumption patterns while bottom-up methods use a fine modeling of users consumptions as to obtain realistic aggregated consumption curves.
-In this paper, we used a bottom-up model since the end user, or relatively small aggregations of end users, are in our interest.
-The main objective was to capture both daily patterns and seasonal variations of the consumptions.
-We assumed an additive model where the consumption of an agent is the sum of a seasonal heating term that depends on the outside temperature and an electronic consumption term that only depends on the hour of the day.
-By denoting $ \tau(t) $ the outside temperature at timestamp t, we can express the consumption $ P_{i}^{D}(t) $ of agent i at time t :
-
-\begin{equation}
-P_{i}^{D}(t) = \mathcal{F}_{i}^{heat}(\tau(t), t) + \mathcal{F}_{i}^{elec}(t)
-\end{equation}
-where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
-In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification.
-By using thermodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :
-
-\begin{equation}
-\mathcal{F}_{i}^{heat}(\tau(t), t) = \dfrac{B_{i}}{R_{i}} \left[ T_{i} - \tau(t) \right]
-\end{equation}
-where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
-We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers.
-We also denote by $ \omega_{i}(t) = \{ \omega_{i}(t_{0}),...,\omega_{i}(t_{24}) \} $ the vector of the average fraction of $ \Omega_{i} $ used for each hour.
-We can therefore write :
-
-\begin{equation}
-\mathcal{F}_{i}^{elec}(t) = \Omega_{i} ( \omega_{i}(t) + \epsilon )
-\end{equation}
-where $ \epsilon $ is a noise term.
-The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors.
-Business or residential areas for instance can be easily distinguished with this kind of model.
-
 \end{document}

### 2015-11-02 Article.tex
figures sizes
update

@@
 
+{\bf Keywords:} prosumers, smart-grid, coalitions, correlation, stability 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Designing stable power systems is a classical engineering challenge since blackouts can have catastrophic consequences.
-One obvious condition for stability revolves around sustaining at any time a power generation that meets the demand.
-If one is larger than the other, the system deviates from its synchronous state.
-If nothing is done to return to the synchronized equilibrium, this can lead to catastrophic cascades of failures \cite{Brummitt2012} \cite{Wang2009}.
-With traditional power plants based on fossil energies, the production can be scheduled in order to sustain the predicted consumption.
-Deviations from this schedule can then be supported almost on real time by using fast response power plants, interconnections with border countries, as well as regulator entities.
-Even if most individual consumers may not realize it, the electricity prices are not constant rates and evolve with the production/consumption conditions on the grid.
-It is often assumed that the granularity of these prices is meant to increase in the future as a way to pass on the production conditions on the end users \cite{Jiang2014}.
-The pricing of electricity and the necessity for the grid operator to have different emergency reserves lead to an economy setting for electricity, where operations and various kind of contracts are decided on a market \cite{Europe}.
-Such a market environment clearly necessitates communication in order to monitor and obtain information from the grid as well as exchange between participants.
-These kind of settings are already being used at the transport level of current power grid with large renewable plants in the production portfolio.
-Using electricity markets appears thus as a way to manage the reliability of the whole system, especially when the use of renewables is important.
-
-The smart grid vision however goes further and revolutionizes this top-down centralized architecture by assuming that bidirectional power flows are allowed.
-This would change completely the nature of the traditional grid since production could be located down to the very end of the distribution networks.
-Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads \cite{Ramchurn}.
-On the other hand, the user of renewables in the production is continuously growing, and is expected to become majority in a near future.
-These plants completely rely on the presence and the intensity of their respective resources (wind, sun, tiles...).
-Balancing production and demand in such a scenario appears much more challenging and it seems inevitable that the grid should modernize its infrastructures to sustain this transition \cite{Europe}.
-The key to drive safely such a system is assumed to be information, and more precisely, the capacity to measure, communicate, and analyze data on real time.
-
-In this paper, we focus on these "\textit{nodes}" in the distribution network that can produce and consume electricity.
+The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures and algorithms to name a few, sketch a new power system.
+The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems.
+Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition.
+Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages [5].
+In this paper, we focus on these "nodes" in the distribution network that can produce and consume electricity.
 More precisely, we consider a set of agents that own both renewable distributed energy resources (DER) and electrical loads.
 The production of an agent can be, of course, used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
-Such an agent model is known as a "\textit{prosumer}" \cite{Rathnayaka2012} (and will be called accordingly in this paper).
+Such an agent model is known as a "prosumer" [6].
 A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
@@
 
+\begin{figure}
+\label{tab:main_notations}
+	\begin{tabular}{ | l | r | }
+   						N & Number of Prosumers \\
+  						$ P_i(t) $ & Extra-production of agent i at time t \\
+   						$ P_S^{CRCT} $ & Contract value of coalition S \\
+   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
+   						$ P^{MIN} $ & Minimum Contract value \\
+   						$ \alpha $ & Parameter that controls the coalitions sizes \\
+   						$ \epsilon $ & Correlation graph filtering threshold \\
+   						$ N_{COAL} $ & Number of coalitions \\
+   						$ \mathcal{R}_S $ & Resilience of coalition S
+ 	\end{tabular}
+ 	\caption{Main notations.}
+\end{figure}
+
 \begin{equation}
@@
 
+\begin{figure}
+\label{fig:process}
+	\includegraphics[scale=.5]{Fig2}
+	\caption{Process.}
+\end{figure}
+
 \begin{equation}
@@
 The model can be easily extended to multiple power curves accounting for different types of generators.
-More details about power curves and their approximations can be found in the appendix \ref{appendixB} and in \cite{Lydia2014}.
+More details about power curves and their approximations can be found in the appendix and in \cite{Lydia2014}.
 
@@
 \begin{figure}
-\includegraphics[scale=.48]{gaussian_P_star}
+\includegraphics[scale=.48]{./figs/figure_1}
 \caption{{\footnotesize $ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
@@
 \begin{figure}
-\includegraphics[scale=.48]{mean_field}
+\includegraphics[scale=.48]{./figs/figure_2}
 \caption{{\footnotesize Gaussian mean approximation.
@@
 \begin{figure}
-\includegraphics[scale=.48]{cliques_histo}
+\includegraphics[scale=.48]{./figs/figure_3}
 \caption{{\footnotesize Histograms of utility values for coalitions of size 3.
@@
 \begin{figure}
-\includegraphics[scale=.48]{real_utility_2}
+\includegraphics[scale=.48]{./figs/figure_4}
 \caption{{\footnotesize Utility of random coalitions depending on their size.
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figures/coalitions/coalitions6}
+\includegraphics[scale=.48]{./figs/figure_5}
 \caption{{\footnotesize Coalitions formed in the (contract value, volatility) space.
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figures/resilience_both}
+\includegraphics[scale=.48]{./figs/figure_6}
 \caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }

### 2015-11-03 Article.tex
update

@@
 
-{\bf Keywords:} prosumers, smart-grid, coalitions, correlation, stability 
+{\bf Keywords:} prosumers, coalition formation, correlation, stability, clique expansion, electricity market
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 In other words we select $ \epsilon^{\star} $ as the smallest threshold possible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
-The existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
+In practice, since finding cliques requires exponential time, we use triangles \cite{Schank2001} ($k=3$ in eq.
+\ref{epsilon_star}) rather than cliques as soon as the number of nodes is not small.
+Note also that the existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
 The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
@@
 The resilience of a system refers to its ability to perform its tasks when subject to failures of its components.
-Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when node failures occur.
+We consider here the case of random failures of the power electronics of some agents that has the consequence of preventing them from participating in the market.
+Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when some of its internal agents are removed.
 According to our model, the grid operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $.

### 2015-11-04 Appendix.tex
figures sizes
update

@@
 \begin{figure}[b]
-\includegraphics[scale=.5]{search}
+\includegraphics[scale=.5]{./figs/figure_7}
 \caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
@@
 
-\subsection{Net production series}\label{appendixB}
+\subsection{Net production series}
+\label{appendixB}
 
@@
 
+\subsection{Resilience algorithm}
+\label{appendix_resilience}
+The following algorithm (see algorithm \ref{alg:algo4}) explains the process through which the resilience of a coalition structure is explored.
+For a given coalition structure, we consider random failures of physical components of the agents.
+This results in the agent being unable to support its coalition.
+Therefore, the agent is simply removed from the coalition, which impacts the coalition's production.
+Eventually, after some number of agent failures, some coalitions might not be sufficiently stable and productive to fulfill the market's constraints.
+In this case, the coalition fails, and is removed from the market.
+\begin{algorithm}
+	\KwData{$ CS = \left\{ S_1,S_2,...,S_k \right\} $ Coalition structure \;}
+
+	$ pool = \cup_{k \in N_{COAL}} S_k $ \; 
+	\While{ $ pool \neq \emptyset $}{
+		Select randomly agent i from pool\;
+		i fails $ \Rightarrow $ Remove i from its coalition : $ S_k \leftarrow S_k - \{i\} $ \;
+		$ pool = pool - \{i\} $\;
+		\If{ $ Pr \left[ P_{S_k} < P^{MIN} \right] > \phi $}{
+			$S_k$ fails $ \Rightarrow $ Remove $S_k$ from the market
+			} 
+		}
+	\caption{Random failures algorithm}
+\label{alg:algo4}
+\end{algorithm}
+
 \bibliographystyle{IEEEtran}  

### 2015-11-04 Article.tex
update

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot}
+\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot \\
+\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
+Telecom SudParis/Institut Mines Telecom\\
+Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
 
@@
 More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
-If a contract is bought, the entity commits to injecting exactly, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
+If a contract is bought, the entity commits to injecting, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
 Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
-Using storage as buffers in order to reduce risks is a popular approach.
-Actually, a whole branch of the smart grid literature is even studying the possibility to use electric vehicles as moving storage capacities for stabilizing the grid.
+Using storage as buffers as to reduce risks is a popular approach.
 Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable.
-This is clearly unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.
+This is unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.
 
 Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variability.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}, since an aggregator can be considered as a single node by the level above it.
-On the other hand, it is well-known that diversification of the assets is a way of minimizing risks when constructing a portfolio.
+Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}.
+On the other hand, it is known that diversification of the assets is a way of minimizing risks when building a portfolio.
 One thus expect a more stable and predictable energy production for a coalition than for a single agent.
-Nevertheless, all coalitions are obviously not equally stable or productive, which means that special attention should be paid to the aggregation step.
+Nevertheless, all coalitions are not equally stable or productive.
+Special attention should thus be paid to the aggregation step.
 Recall that prosumers have both a consumption and a production component, each depending on location and time, meaning that there are complex underlying correlations between the agents.
@@
 Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of risk acceptable.
-In other words, it tries to maximize coalitions profits without considering individual retributions to single agents (a task often studied through game theory).
+In other words, it tries to maximize coalitions profits without considering individual retributions to single agents.
 Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
@@
 Finally, section \ref{sec:results} provides some results both on performance of the method and resilience of the coalitions formed.
+The overall process is illustrated on figure \ref{fig:process}.
 
@@
 
-As explained in the section above, allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
-Indeed, on the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables by definition only produce when the resource is present.
+Allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
+On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables only produce when the resource is present.
 Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
 One possible solution consists in shifting some of the loads towards high productivity periods.
-Demand side management techniques can indeed mitigate the consumption peaks and wast less production \cite{Milligan2010} \cite{Logenthiran2012}.
+Demand side management techniques can indeed mitigate the consumption peaks and waste less production \cite{Milligan2010} \cite{Logenthiran2012}.
 Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing appears as a good way to give incentives to them.
@@
 
+\begin{figure}
+\label{fig:process}
+	\includegraphics[scale=.5]{Fig2}
+	\caption{Process}
+\end{figure}
+
 An essential component of the smart grid is the smart meter which makes the interface between the end user and the rest of the system.
@@
 Fortunately, these kinds of data are easier to find, and since the development of small personal weather stations, their geographical granularity keeps increasing.
-Since these quantities depend both on time and location, we discretize time into slots and space into zones in the following.
+Since these quantities depend both on time and location, we discretize time into slots and space into zones in the following (see block 1 of fig.
+\ref{fig:process}).
 A zone is simply a portion of the considered region of study for which we sampled data.
@@
 
-\begin{figure}
-\label{fig:process}
-	\includegraphics[scale=.5]{Fig2}
-	\caption{Process.}
-\end{figure}
-
 \begin{equation}
@@
 Obviously, the actual extra-production will not be constant at this value and will oscillate due to intermittences in the production and consumption.
-If S always produces more than $ P_{S}^{CRCT} $, it will never have to pay penalties, but it is losing some gains since it could have announced a higher contract value.
+If S has an available production always greater than $ P_{S}^{CRCT} $, it is losing some gains since it could have announced a higher contract value.
 If the production oscillates around $ P_{S}^{CRCT} $, by using batteries or demand side management techniques, S could be able to maintain its production to the contract value at any time.
@@
 We consider that prosumers fail randomly, and we denote by $ \psi \in [0,1] $ the fraction of agents that failed.
-Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolves according to $ \psi $.
+Figure \ref{fig:resilience} exhibits how the resilience of the coalition structures evolves according to $ \psi $ (see appendix section \textit{Resilience algorithm} for more details).
 On the top subplot, $ P^{MIN} $ was voluntarily selected relatively low such that the resiliences of the three structures fit on the same figure.

### 2015-11-05 Article.tex
update

@@
 This section provides most of the notations and introduces important concepts for the rest of the paper.
-We consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers configured randomly, and for each agent, we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
+We consider a set $ \mathcal{A} = \{a_{1},a_{2},...,a_{N} \} $ of N prosumers of the distribution network.
+Each agent is configured randomly and we simulate its extra-production $ P_{i}(t),\ \forall i \in \mathcal{A} $ from 2006 to 2012.
 Based on these historical values, our objective is now to form groups of prosumers so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.

### 2015-11-13 letter.tex
respond letter latex version

\documentclass{letter}
\makeatletter
\newenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\usepackage[]{algorithm2e}
\newcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother
\signature{Vincent Gauthier}
\address{Vincent Gauthier \\ Institut Mines Telecom \\ Nano Innov, Saclay \\ France}
\begin{document}

\begin{letter}{Dr. Jianhui Wang \\ Editor-in-Chief \\ Journal Transactions on Smart Grid } % Addressee name and address

\opening{ Dear Dr. Jianhui Wang,}

The authors would like to thank you for giving us the opportunity to revise and resubmit this manuscript. We resubmit this revision before the agreed upon deadline, December 21, 2015. We appreciate the time and details provided by each reviewer and by you and have incorporated the suggested changes into the manuscript to the best of our ability. The manuscript has certainly benefited from these insightful revision suggestions.

See bellow our replies to the reviewers’ comments:

\textbf{Reviewer: 1}

\textbf{Comment 1:} However I can see a case study is presented and the results are summarized in the last section, but I recommend you to give more detail about the case study conditions. I can see so many details are presented in the main body of the manuscript but it is difficult to track till end of the paper.

\textbf{Comment 2:} I do recommend you also to have a general schematic of the system to make it clear the elements of your system.

\textbf{[Answer to 1 and 2]} We agree with the reviewer that a general schematic would help the reader. In order to illustrate the overall process we added a figure (Figure 1) in the article. This figure explains the main steps from the simulation of the agents’ production time series to the formation of the coalitions. We also added a table of the main notations that might help the reader.

\textbf{Comment 3:} In fig. 5, how does the sampling frequency affect the volatility? How does the sampling frequency of your data affect the variance? There are some paper that demonstrates the effectiveness of the available resources data resolution on the expected variability of the renewable energy systems, I think this could be interesting if you can provide some information in this regard.

\textbf{[Answer]} The results presented in this paper come from hourly sampled weather data. Therefore, all the correlations between the agents are derived from hourly sampled time series. This low frequency sampling has the effect of hiding short time intermittencies that are present in both wind speed and solar irradiance real dynamics. According to \cite{Anvari2015}, “extreme events in the solar irradiance and wind power fluctuations are much more frequent in short time scales” than expected with a Gaussian process. This is especially true and understandable for solar irradiance that can change drastically in case of clouds passing in a sunny sky for instance. Although we did not, in this paper, study how the correlation structure evolves with the sampling frequency of the data, we believe that the variance of the power distribution of the coalitions will increase with the sampling frequency. Nevertheless, as our algorithm tries to minimize intra-coalition correlations, we believe that short time fluctuations will tend to impact less “our” coalitions than the ones obtained with the other presented strategies.

\textbf{Comment 4:} What is the affect of using stochastic modeling of the resources on the coalitions resulting from the proposed method? Using a real data set is a good case study, but in the conclusion you mentioned "the coalitions resulting from our algorithm better withstand losses of agents", did you do any comparison? It is not clear how you came up to this conclusion. I think you should have an apple-to-apple comparison, stochastic modeling can help you.

\textbf{Comment 5:} There are several parameters that makes the systems fail, in page 7 what do you mean by node failure? Do you mean lack of resources OR failure of power electronics interfaces of renewable energy systems? Please clarify that in the paper.

\textbf{[Answer to 4 and 5]} Lack of resources is a key issue of this paper. However in figure 8, we specifically concentrate on the power electronics or line failures using a simplified algorithm. More precisely, we want to study how coalitions resulting from different formation algorithms respond to failures. Failures are considered random and when some agent fails, it is deleted from its coalition. This has consequences on the production of the coalition that may cause the violation of one of the market rules (stability and minimum production), which in turn will lead to the deletion of the coalition from the market. We used this method in order to compare how the coalition structures obtained with the three different algorithms (“correlated”, “de-correlated”, and “random”) respond to this type of failure. We concluded that the coalitions resulting from our “de-correlation” algorithm were able to better sustain agents loses than the two others. We agree with the reviewer and added, for clarity, a section in the appendix (see section Resilience algorithm) that gives the pseudo code of the process discussed above (reproduced here as algorithm \ref{alg:algo4}).

\begin{algorithm}
	\KwData{$ CS = \left\{ S_1,S_2,...,S_k \right\} $ Coalition structure \;}

	$ pool = \cup_{k \in N_{COAL}} S_k $ \; 
	\While{ $ pool \neq \emptyset $}{
		Select randomly agent i from pool\;
		i fails $ \Rightarrow $ Remove i from its coalition : $ S_k \leftarrow S_k - \{i\} $ \;
		$ pool = pool - \{i\} $\;
		\If{ $ Pr \left[ P_{S_k} < P^{MIN} \right] > \phi $}{
			$S_k$ fails $ \Rightarrow $ Remove $S_k$ from the market
			} 
		}
	\caption{Random failures algorithm}
\label{alg:algo4}
\end{algorithm}

 We also added the following lines in the article:

\textit{“We consider here the case of random failures of the power electronics of some agents that has the consequence of preventing them from participating in the market. Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when some of its internal agents are removed.”}

\textbf{Comment 6:} Fig. 5 is a great visualization of the problem, but how are you going to come-up to a final decision? Can you provide formulation to make decision using decision theory? I think this will help and support your great work.

\textbf{[Answer]} We thank the reviewer for this comment. In figure 5 we represented the distribution of the coalition structures using Monte Carlo sampling as comparison between the results of the three algorithms. All algorithms need a desired number of coalitions that they should output. In our work, the decision on the number of coalitions is therefore assumed to come from the user, and is not investigated. Given the desired number of coalitions, the algorithms use the utility function (eq. 11) as the decision criterion between the different coalition structures (all containing the same number of coalitions). This can be seen in more details in the pseudo-code of the algorithms in the appendix (see section Algorithms of the appendix). We also reproduce the pseudo code of the greedy optimization below (alg. \ref{alg:algo1}). Our algorithm for instance starts with seeds in the correlation graph and expands them as long as the utility is improved. It stops when no agent is able to increase the coalitions’ utilities.

\begin{algorithm}
 \KwData{$P_{i}$ series,\\ Grid policy $ (P^{MIN},\phi) $,\\ Desired number of coalitions $ N_{COAL} $,\\ size of starting cliques k}
 \KwResult{ $ CS = \{ S_{1},...,S_{N_{COAL}}\} $ }
 Compute $ G_{2}^{\epsilon^{\star}} $ \;
 Find the $ N_{COAL} $ cliques in $ G_{2}^{\epsilon^{\star}} $\;
 \While{$ \mathcal{U}(CS) $ is improving}{
 	\For{each clique}{
 		Find $ i^{\star} $ \;
 		\If{ $ \delta_{clique}(i^{\star}) \geq 0 $ }{
 			$ clique \leftarrow clique \cup \{i^{\star} \} $ \;
 			}
 		\If{ $ \exists j \in clique,\ s.t\ \delta_{clique}(j) < 0 $}{
 			$ clique \leftarrow clique - \{j \} $ \;
 			}
   		}
  	}
 \caption{Local greedy optimization algorithm}
 \label{alg:algo1}
\end{algorithm}

\textbf{Comment 7:} In general the paper is in well format, however it needs some grammar revision as well as proof reading, I saw couple of errors such as the question marks in page 3, column 2.

\textbf{[Answer]} We tried to correct all the remaining errors in the paper.

\textbf{Reviewer: 2}

\textbf{Comment 1:} Using a Table with ALL notations at the very beginning of the manuscript.

\textbf{[Answer]} We respond to the reviewer’s comment and added a table of the main notations in the article (figure 2). Because of space limitation, we were not able to put all the notations in this figure. Therefore, we also added a table of all the notations in the appendix.

\textbf{Comment 2:} Making clear the fact which grid/system operator is referred to (distribution vs transmission)

\textbf{[Answer]} In this paper, we consider aggregations of prosumers located in the distribution network as stated in the introduction:
“Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages. In this paper, we focus on these "nodes" in the distribution network that can produce and consume electricity.”

For clarity, we added the following line in the beginning of section 4:

\textit{“We consider a set A of N prosumers of the distribution network.”}

\textbf{Reviewer: 3}

\textbf{Comment 1:} In Section III, is load uncertainty included in the generation of prosumer patterns? If yes, what load patterns were used? What happens if the load is so high that prosumers cannot generate energy?

\textbf{[Answer]} In this paper, we generate consumption historical series using real temperature traces (see section Consumption of the appendix). More precisely, we decompose the consumption of a prosumer into a heating and an electronic consumption term.

\begin{equation}
P_{i}^{D}(t) = \mathcal{F}_{i}^{heat}(\tau(t), t) + \mathcal{F}_{i}^{elec}(t)
\end{equation}
where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature $ \tau(t) $ to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day. The purpose of this decomposition is to reproduce both seasonal patterns as well as daily patterns of consumption in a simple model.

In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification. By using thermodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :

\begin{equation}
\mathcal{F}_{i}^{heat}(\tau(t), t) = \frac{B_{i}}{R_{i}} \left[ T_{i} - \tau(t) \right]
\end{equation}
where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.

We further denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers. We also denote by $ \omega_{i}(t) = \{ \omega_{i}(t_{0}),...,\omega_{i}(t_{24}) \} $ the vector of the average fraction of $ \Omega_{i} $ used for each hour ($\forall i,\ \forall t,\ \omega_i(t) \in [0,1] $). We can therefore write :

\begin{equation}
\mathcal{F}_{i}^{elec}(t) = \Omega_{i} ( \omega_{i}(t) + \epsilon )
\end{equation}
where $ \epsilon $ is a noise term drawn from a normal distribution. The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors. Business or residential areas for instance can be easily distinguished with this kind of model.

At time t, if the consumption of a prosumer is larger than its production, then this prosumer will consume energy from its coalition rather than contributes to the production. If all prosumers of a same coalition are strongly correlated and tend to behave in the same way this results in a very unstable output production. However, in the case of uncorrelated prosumers, over-production of some of them might tend to compensate the under-production of the others, yielding a more stable output. In the case of a prosumer that almost never produce, its marginal utility would be very small or negative, such that it will not be selected in the expansion phase.

\textbf{Comment 2:} The selection of the parameter alpha of the coalition utility function [eq. (14)] relies heavily on the Gaussianity assumption for the generated net power, which may not hold.

\textbf{[Answer]} We agree with the reviewer’s comment. The purpose of the alpha parameter is to control to some extent the expected sizes of the coalitions. In order to obtain an analytical formulation for this parameter, we used the assumption that the distributions of the coalitions’ production were Gaussians. When the number of agents in the coalitions increases, the distribution of the coalition will tend to a normal distribution for independent agents. Although independence is generally not true and is not required for our method, we found that using this assumption for tuning alpha gives good results in practice. Other parameter selection methods such as grid search could also be used although not presented in the article.

\textbf{Comment 3:} The selection of the threshold epsilon for the decorrelation graph [eq. (15)] appears to incur exponential complexity, as it requires availability of cliques of the graph.

\textbf{[Answer]} We agree with the reviewer’s point that finding cliques in graphs is usually a hard problem. In order to overcome this difficulty when the number of nodes is not small, we restrict the algorithm to triangles rather than cliques. Finding $ \epsilon{\star} $ through (eq.(15)) with $ k = 3 $ (looking only at triangles) requires to count the number of triangles in a graph. There exist several techniques to achieve this (see \cite{Schank2001}). One of the most straightforward is:

\begin{equation}
 N_{triangles} = \frac{1}{6} Tr \left[ A^3 \right] 
\end{equation}

Where A is the graph adjacency matrix. This leads to a counting algorithm which runs in $ O(n^q) $ where q is the matrix multiplication exponent (considered to be less than 2.4).

For finding the starting seeds for the clique expansion, we need not only to count the number of triangles, but we also have to enumerate them. Again, several techniques exist such as node-iterator or edge-iterator (more information in \cite{Schank2001}).

We clarify this point by adding the following text in section IV-A:

\textit{“In practice, since finding cliques requires exponential time, we use triangles [14] (k = 3 in eq. 15) rather than cliques as soon as the number of nodes is not small.” }

\textbf{Comment 4:} Implementation of the algorithm also requires the correlation coefficient between any two prosumers, which may as well not be available.

\textbf{[Answer]} We agree with the reviewer’s comment. In order for our algorithm to work, we need the correlation coefficients between any two prosumers that can be obtained from historical time series. Having these series is an assumption of our work.

\textbf{Reviewer: 4}

\textbf{Comment 1:} The reviewer is wondering if the research question is application for the general context or only for specific environment. In general, it would be expected that the prosumers have their own freedom to choose their aggregator, similar way to have the electricity contract with the energy supplier. Forcing the end users to be in dedicated coalition would be realistic.

\textbf{[Answer]} In this paper, we indeed do not consider agent freedom to choose what coalition to join. We only consider the maximization of a global utility function that seeks de-correlation inside the coalitions. Perhaps, a more realistic framework based on Game theory will consider a utility function per agent with different parameters. Each (rational) agent would then focus on the maximization of its own profits, and coalition structures might emerge from series of split and merge actions. Although not considered in the present work, we believe that building a game which would take into account the stability of power injections through the correlation structure might be possible.

\textbf{Comment 2:} The reviewer was not convincing about the statement in page 4, line 21-24, “If S always procedures… a higher contract value”. The reviewer believes that if S deviates from the scheduled value, even if higher extra-production it should still have the penalty cost.

\textbf{[Answer]} We agree with the reviewer that over-producing is also problematic toward the grid’s stability. And coalitions should pay penalties if deviations (positive or negative) from the contract occur. However, in our framework, we make the distinction between the available power of a coalition, and the power it really injects in the grid. Although the latter has to be equal to the contract, we allow the former to be larger since the difference can be used for storage or other usages. We changed the misleading statement to:

\textit{“If S has an available production always greater than $ P_{S}^{CRCT} $, it is losing some gains since it could have announced a higher contract value.”}

\textbf{Comment 3:} The reviewer is doubting about the interaction model developed in the paper, specifically mentioned in the second column of page 4. Why should the grid operator concern about the coalition of prosumers with its probability of under-producing? Normally saying, under-producing should not cause any grid issue while significant over-producing from the end users would might cause current congestions or voltage violation for the distribution network.

\textbf{[Answer]} In this paper, we consider the case of a large penetration of renewables in the distribution network. A significant portion of the production is assumed to come from these coalitions of prosumers, in such a way that the grid operator is scheduling the plants according to the contract values. If coalitions do not satisfy their contracts by under-producing, it might be the case that the grid operator has to re-balance consumption and production by turning on fast ramping generators. This might generates high operational costs for the grid operator.

\textbf{Comment 4:} Keywords are missing in this paper!

\textbf{[Answer]} We have considered the reviewers request and added keywords at the beginning of the article.

\textbf{Comment 5:} The introduction part is too long with well-known context that can be found in many previous works. It is recommended to refer more to the literature and shorten the introduction part to go direct to the research problem/questions

\textbf{[Answer]} We agree with the reviewer’s comment and we shortened the introduction.

\textbf{Comment 6:} Terminologies of reliability and stability have been used similarly here. Actually, power system researcher would consider stability with different meaning.

\textbf{[Answer]} In this paper, we used the word “stability” to characterize the “smootheness” of the power output of the coalitions. In this paper, this “smootheness” appears as a consequence of the variance of the coalitions’ power distributions. Therefore the “stability” and “reliability” concepts of our paper are strongly connected but we wanted to make the distinction. Naturally, we agree with the reviewer that these terms are used here in a special way, and might be considered differently by power system researchers.

\bibliographystyle{unsrt}
\bibliography{letter}

\end{letter}
\end{document}

### 2015-11-19 Appendix.tex
update

@@
 \usepackage{fancyhdr}
+\usepackage{multicol}
+
 %
@@
 
+\onecolumn
+\centering
 \appendix
+   \subsection{Notations}
+
+	\begin{tabular}{ | l | r | }
+   						N & Number of Prosumers \\
+   						$ P_i^P(t) $ & Available production of agent i at time t \\
+   						$ P_i^D(t) $ & Consumption of agent i at time t \\ 
+  						$ P_i(t) = P_i^P(t) - P_i^D(t) $ & Available extra-production of agent i at time t \\
+  						$  WT_i $ & Wind turbine set of agent i \\
+  						$ PV_i $ & Solar array set of agent i \\
+  						$ DER_i = WT_i \cup PV_i $ & DER of agent i \\
+  						$ \nu_i(t) $ & wind speed (in $ m s^{-2} $) at agent i's location and time t \\
+  						$ \Psi_i(t) $ & solar irradiance (in $ W m^{-2} $) at agent i's location and time t \\
+  						$ \mathcal{F}_{WT} $ & power curve for the wind turbines \\
+  						$ \mathcal{F}_{PV} $ & power curve for the solar arrays \\
+   						$ P_S^{CRCT} $ & Contract value of coalition S \\
+   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
+   						$ P^{MIN} $ & Minimum Contract value \\
+   						$ \alpha $ & Parameter that controls the coalitions sizes \\
+   						$ \epsilon $ & Correlation graph filtering threshold \\
+   						$ N_{COAL} $ & Number of coalitions \\
+   						$ \mathcal{R}_S $ & Resilience of coalition S \\
+   						$ \lambda $ & unitary price rate for electricity \\
+   						$ \delta_S(i) $ & marginal utility of agent i in coalition S \\
+   						$ \rho_{ij} $ & Pearson's correlation coefficient between $ P_i$ and $P_j$ \\
+   						$ \Theta_k(G) $ & set of non overlapping cliques of size k in graph G \\
+   						$ \mathcal{R}_S $ & resilience of coalition S
+ 	\end{tabular}
+
+\twocolumn
 \subsection{Algorithms}
@@
 \bibliography{Article}
+
 \end{document}

### 2015-11-19 Article.tex
update

@@
 
-The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures and algorithms to name a few, sketch a new power system.
+The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system.
 The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems.
@@
 In this paper, we focus on these "nodes" in the distribution network that can produce and consume electricity.
-More precisely, we consider a set of agents that own both renewable distributed energy resources (DER) and electrical loads.
-The production of an agent can be, of course, used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
+More precisely, we consider agents owning both renewable distributed energy resources (DER) and electrical loads.
+The production of an agent can be used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
 Such an agent model is known as a "prosumer" [6].
 A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
-While the latter intend to maximize their benefits with higher production contracts, the operator primary concerns are related to the quality and stability of the injections.
-The coalitions that we wish to form should thus be both stable and productive.
+While the latter intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the injections.
+
+%The coalitions that we wish to form should thus be both stable and productive.
 This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation happen.
 
-More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
+%More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
 If a contract is bought, the entity commits to injecting, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
@@
 
-Forming coalitions/aggregations is an envisaged solution to both the number of entities soliciting the operator, and their high variability.
+We rely on a market representation of the interactions.
+Coalitions of prosumers estimate a production capacity for an upcoming period and announce it on the market.
+If some other entity buys the contract, the coalition has to inject exactly the contracted amount of power at any time.
+Obviously, deviations will cause financial penalties to the coalitions, such that respecting the contracts is paramount.
+Since prosumers use exclusively renewables, we are particularly concerned with the probability that the available capacity of a coalition falls below its contract value.
+In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.
+
+We focus on coalitions because it appears as a solution to both the number of prosumers communicating with the operator as well as their high variabilities.
 Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}.
 On the other hand, it is known that diversification of the assets is a way of minimizing risks when building a portfolio.
-One thus expect a more stable and predictable energy production for a coalition than for a single agent.
+We thus expect more stable and predictable energy productions for coalitions than for single agents.
 Nevertheless, all coalitions are not equally stable or productive.
 Special attention should thus be paid to the aggregation step.
-Recall that prosumers have both a consumption and a production component, each depending on location and time, meaning that there are complex underlying correlations between the agents.
+Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents.
 This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
@@
 Adding some intra-zone noise can easily be done though not considered in this paper.
-More formally, we denote by $ P_{i}(t) $ the instantaneous extra-production of agent i at time t :
+More formally, we denote by $ P_{i}(t) $ the instantaneous available extra-production of agent i at time t :
 
@@
 In the rest of the paper, we use french weather data \cite{Infoclimat} starting in January 2006 and ending in December 2012, with a sampling frequency of three hours, and generate N timeseries of extra-production over this date range.
+Note that this low sampling frequency might hide short time intermittencies (see \cite{Anvari2015} for more information).
+Studying the evolution of the correlation structure with the sampling frequency is out of the scope of this paper, but might lead to interresting results.
 
@@
 Based on these historical values, our objective is now to form groups of prosumers so that the global power production resulting from the superposition of individual's extra-productions be both sufficiently high and predictable.
-Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the extra-production of coalition S at time t.
+Let $ P_{S}(t) = \sum_{i \in S} P_{i}(t) $ be the available extra-production of coalition S at time t.
 

### 2015-11-19 letter.tex
update

@@
 Nevertheless, as our algorithm tries to minimize intra-coalition correlations, we believe that short time fluctuations will tend to impact less “our” coalitions than the ones obtained with the other presented strategies.
+
+We added the following text at the end of section 3 :
+
+\textit{Note that this low sampling frequency might hide short time intermittencies (see \cite{Anvari2015} for more information).
+Studying the evolution of the correlation structure with the sampling frequency is out of the scope of this paper, but might lead to interresting results.}
 
@@
 
-\textbf{[Answer]} In this paper, we consider the case of a large penetration of renewables in the distribution network.
-A significant portion of the production is assumed to come from these coalitions of prosumers, in such a way that the grid operator is scheduling the plants according to the contract values.
-If coalitions do not satisfy their contracts by under-producing, it might be the case that the grid operator has to re-balance consumption and production by turning on fast ramping generators.
-This might generates high operational costs for the grid operator.
+\textbf{[Answer]} We agree with the reviewer that, in the case of a large penetration of renewables in the distribution network, the production capacity should be larger than for traditional power plants.
+This is the consequence of the uncertainty that comes with most renewables.
+It results from this point that, at certain periods, the production might be drastically larger than the consumption, which might indeed cause various problems in the network.
+Storage, demand-side managment, inter-connection with neighboring countries, as well as energy markets are often assumed to provide ways of stabilizing the grid in such situations.
+
+Nevertheless, we want to point out that, in this paper, we focus on the market side of the system rather than the grid pure stability.
+More precisely, we consider the formation of coalitions of prosumers that seek to sell energy to some entity (grid operator, aggregators).
+If this entity buys a contract, the coalition commits to injecting exactly $ P_S^{CRCT} $ :
+
+\begin{equation}
+\forall t,\ P_S^{injected}(t) = P_S^{CRCT} 
+\end{equation}
+
+Obviously, if the actual injected power deviates from the contract, the coalition should pay penalties proportional to $ \sum_t \left| P_S^{injected}(t) - P_S^{CRCT} \right| $.
+However, within this framework, we build the coalitions by looking at their available power $P_S(t)$, i.e, the maximum production that a coalition can inject at time t : $ P_S^{injected}(t) \leq P_S(t) $.
+In this situation, we believe that the risk $ Pr[P_S(t) < P^{CRCT} ] $ is a key quantity since, in this case, the contract cannot be respected.
+In cases where the available production is higher than the contract $ P_S(t) > P^{CRCT} $, it might be possible to shut some renewables down, or to use storage, to maintain the power injection at the contract value.
+
+We added the following text in the introduction as to clarify this point :
+
+\textit{We rely on a market representation of the interactions.
+Coalitions of prosumers estimate a production capacity for an upcoming period and announce it on the market.
+If some other entity buys the contract, the coalition has to inject exactly the contracted amount of power at any time.
+Obviously, deviations will cause financial penalties to the coalitions, such that respecting the contracts is paramount.
+Since prosumers use exclusively renewables, we are particularly concerned with the probability that the available capacity of a coalition falls below its contract value.
+In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.}
 

### 2015-11-20 Article.tex
corrections Monique

@@
 A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
-While the latter intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the injections.
+While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the injections.
 
@@
 We focus on coalitions because it appears as a solution to both the number of prosumers communicating with the operator as well as their high variabilities.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performance on the communication side \cite{Negeri2012}.
+Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}.
 On the other hand, it is known that diversification of the assets is a way of minimizing risks when building a portfolio.
@@
 Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents.
-This is a central topic of the present paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
+This is a central topic of this paper : given N prosumers, what coalitions should be formed so that the compromise between expected production and variability is optimized ?
 
@@
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
-We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to take.
+We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to accept.
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
-Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of risk acceptable.
+Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of acceptable risk.
 In other words, it tries to maximize coalitions profits without considering individual retributions to single agents.
@@
 
-Allowing entities using mainly renewables to inject power in the grid is a difficult challenge.
-On the contrary to fossil plants whose production can be scheduled in advance to meet the expected consumption, renewables only produce when the resource is present.
-Unfortunately, these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}.
+Since renewables only produce when the resource is present and 
+because these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}, allowing large penetration of renewables in the grid appears as a difficult challenge.
 One possible solution consists in shifting some of the loads towards high productivity periods.
@@
 Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing appears as a good way to give incentives to them.
-By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extend.
+By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extent.
 
@@
 The optimization of expected returns to risk is a traditional goal in finance, and a wide literature exists on this topic.
-It is well-known for instance, that the more risk one is willing to take, the higher his potential gains.
+It is well-known for instance, that the more risk one is willing to accept, the higher his potential gains.
 On the contrary, when investing exclusively on low risks assets, one should expect relatively small gains.
@@
 \label{fig:process}
-	\includegraphics[scale=.5]{Fig2}
+	\includegraphics[scale=.45]{Fig2}
 	\caption{Process}
@@
 Adding some intra-zone noise can easily be done though not considered in this paper.
-More formally, we denote by $ P_{i}(t) $ the instantaneous available extra-production of agent i at time t :
+We denote by $ P_{i}(t) $ the instantaneous available extra-production of agent i at time t :
 
@@
 
-This is the best contract value that the coalition S can afford giving the stability policy $\phi$ of the grid operator.
+This is the best contract value that the coalition S can afford for a given stability policy $\phi$ of the grid operator.
 Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $.
@@
 The Gaussian assumption of this small example is convenient as it allows us to write $ P_{S}^{CRCT \star} $ analytically.
-Nevertheless, such assumption is rather unrealistic in practice.
+Nevertheless, such an assumption is rather unrealistic in practice.
 In the following, we keep the same framework but release this Gaussian assumption unless the contrary is specified (see eq.
@@
 \caption{{\footnotesize Histograms of utility values for coalitions of size 3.
-Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triplets.
+Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triples.
 Cliques tend to exhibit higher utilities than random coalitions.}}
@@
 Indeed, the larger the desired cliques, the more dense $ G_{2}^{\epsilon^{\star}} $ becomes (see equation \ref{epsilon_star}).
-There is a point where cliques results more from noisy edges than true de-correlation, which decreases the quality of the results.
+There is a point where cliques result more from noisy edges than true de-correlation, which decreases the quality of the results.
 Directly mapping cliques to coalitions by this de-correlation oriented approach is thus not sufficient.
@@
 
-The first steps consists in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
+The first steps consist in computing the de-correlation graph $ G_{2} $ as well as the optimal threshold $ \epsilon^{\star} $.
 Cliques of size k in $ G_{2}^{\epsilon^{\star}} $ are considered as coalition seeds.
@@
 Recall that the purpose of this parameter is to take into account some constraints on the coalition's sizes if needed.
-In this paper, we do not have any technical constraints on coalition sizes even if we designed the utility such that these could be taken into account.
+In this paper, we do not have any technical constraints on coalition sizes though we designed the utility such that these could be taken into account.
 We select the desired size as being $ \lfloor N/N_{COAL} \rfloor $ (where $ \lfloor.\rfloor $ means floor).
@@
 A greedy algorithm that starts with cliques of the "de-correlation" graph of the agents and makes local improvements offers a good compromise between speed and quality of the results.
-We compare these results with random samplings, and an opposite strategy that clusters correlated agents together.
+We compared these results with random samplings, and an opposite strategy that clusters correlated agents together.
 We showed that the coalitions resulting from our algorithm are able to provide more power to the grid with a lower volatility.

### 2015-11-23 Article.tex
authors update

@@
 \begin{document}
-\pagestyle{plain}
-\pagenumbering{arabic}
+%\pagestyle{plain}
+%\pagenumbering{arabic}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot \\
-\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
-Telecom SudParis/Institut Mines Telecom\\
-Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
+\author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot
+%\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
+%Telecom SudParis/Institut Mines Telecom\\
+%Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
+\thanks{Authors are with CNRS UMR 5157 SAMOVAR, Telecom SudParis/Institut Mines Telecom, e-mail: (\{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu)}}%
 
@@
 \begin{figure}
-\label{fig:process}
 	\includegraphics[scale=.45]{Fig2}
 	\caption{Process}
+\label{fig:process}
 \end{figure}

### 2016-02-02 Article.tex
numeric setup figure 4
numerical setup figure 5

@@
 \includegraphics[scale=.48]{./figs/figure_2}
-\caption{{\footnotesize Gaussian mean approximation.
+\caption{{\footnotesize Gaussian mean approximation ($N=100,\ \phi=0.1,\ \bar{\mu}=5,\ \bar{\sigma}=.5,\ \bar{\rho}=0.2 $).
 Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean size of the coalitions (see equation \ref{eq:alpha_star}).
@@
 Note also that the existence of $ \epsilon^{\star} $  as defined in equation  \ref{epsilon_star} is not guaranteed.
-The users has indeed to provide consistent values of $ N_{COAL} $ or k compared to the size of the agent population $ \mathcal{A} $.
 
@@
 \includegraphics[scale=.48]{./figs/figure_3}
-\caption{{\footnotesize Histograms of utility values for coalitions of size 3.
+\caption{{\footnotesize Histograms of utility values for coalitions of size 3 in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
 Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triples.

### 2016-02-02 Article.tex
numeric setup figure 6

@@
 \includegraphics[scale=.48]{./figs/figure_3}
-\caption{{\footnotesize Histograms of utility values for coalitions of size 3 in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
+\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
 Red bars stand for cliques in the decorrelation graph, and blue bars for all the other possible triples.
@@
 \includegraphics[scale=.48]{./figs/figure_4}
-\caption{{\footnotesize Utility of random coalitions depending on their size.
+\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
 Blue dots show real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter.
-On this plot the $ \alpha $ parameter of the utility function was selected according to eq.
-\ref{eq:alpha_star} in order to favor 40 agents coalitions.}}
+On this plot, $ \alpha^{\star} = 0.006 $, and was selected according to eq.
+\ref{eq:alpha_star} in order to favor $\bar{N} = 40$ agents coalitions.}}
 \label{fig:real_utility2}

### 2016-02-02 Article.tex
modifs section V

@@
 
-If $ \alpha $ is set to zero, agents are added as long as they increase the contract value of the coalition.
-If $ \alpha $ is greater than zero, additional agents have to increase the contract value by some factor.
-We relate now $ \alpha $ to the mean sizes of the coalitions $ \bar{N} $ by solving the following equation in the gaussian case (as in section \ref{sec:notations}) :
+If $ \alpha = 0 $, agents are added as long as they increase the contract value of the coalition. If $ \alpha > 0 $, additional agents have to increase the contract value by some factor.
+
+In real situations, the shape of such utility function might be complex with numerous local maximums.
+We choose in this paper to favor the exploration of the space around balanced regions, i.e regions where coalitions sizes are relatively balanced.
+Therefore, we use $ \alpha $ as a proxy for having both a concave utility function and preferences in the search space.
+Given a situation with N prosumers and $ N_{COAL} $ desired coalitions, we wish to select $ \alpha $ such that the utility function tend to favor coalitions of approximately $ \bar{N}=N/N_{COAL} $ agents.
+
+In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated.
+To overcome this problem, we express $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $.
+In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details).
+We select $ \alpha^{\star} $ such that :
 
 \begin{equation}
-\left[ \dfrac{\partial{ U}}{ \partial{|S|}} \right]_{|S| = \bar{N}} = 0
+\left\{ \begin{array}{lll}
+	          \left[ \dfrac{\partial{ U}}{ \partial{|S|}} \right]_{|S| = \bar{N}} = 0 \\
+	          \left[ \dfrac{\partial^2 U}{\partial |S|^2} \right]_{ \alpha = \alpha^{\star} } \leq 0
+	    \end{array} 
+\right.
 \label{eq:derivative}
@@
 
-Furthermore, since considering all the possible interactions between agents is analytically intractable, we use here a mean approximation.
-Any quantity x that varies over the agent set is thus simplified in its mean value $ \bar{x} $.
-Solving equation \ref{eq:derivative} for $ \alpha $ in these conditions leads to:
+Which yields to the following expression for $ \alpha^{\star} $ (see Appendix) :
 
@@
 Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $.
-These curves are only valid in the simplified Gaussian example considered here.
-Nonetheless, they will provide some guidance when using real data.
 As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.

### 2016-02-04 Article.tex
related work update

@@
 
-Since renewables only produce when the resource is present and 
-because these moments do not necessarily coincide with the consumption peak hours \cite{Milligan2010}, allowing large penetration of renewables in the grid appears as a difficult challenge.
-One possible solution consists in shifting some of the loads towards high productivity periods.
-Demand side management techniques can indeed mitigate the consumption peaks and waste less production \cite{Milligan2010} \cite{Logenthiran2012}.
-Since end users tend to seek a maximum utility at a minimum cost, dynamic pricing appears as a good way to give incentives to them.
-By carefully scheduling the prices, it is assumed that the load curve can be shaped to some extent.
+Assuming a large penetration of DER, it is expected that the managment of the system would have to evolve in order to copy with the large amount of unpredictable small generators.
+In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion.
+Allowing the formation of coalitions whose participants commit to achieving some objective against a reward appears also as a promising approach.
+Nevertheless, for these structures to be stable toward defection, agents should have an incentive to join, and cooperate.
+Coalition formation is therefore usually studied through the framework of game theory \cite{Saad2011}.
+In this direction, \cite{Nguyen2013} introduces an agent-based strategy for managing DER aiming at optimal performance in the electricity supply system.
 
-Dynamic pricing is an interesting and useful tool, but it has also some limitations.
-It is likely that dynamic pricing will serve as a shaping mass tool while finer techniques will be needed locally.
-A popular approach in this direction consists in deploying storage devices in the network and using them as electricity buffers.
-Basically these storages would be charged when there is a surplus of production, and discharged when the consumption exceeds the production.
-Although quite simple, this idea causes numerous challenges when it comes to its real implementation.
-As long as the system considered remains small, a centralized control of these buffers could be envisioned.
-However, for real large systems it is likely that more sophisticated decentralized algorithms will be necessary.
-In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion.
+Since most of the DER have a stochastic production, maintaining the system equilibrium in these conditions is a complicated task.
+It is often assumed that electricity storage will provide some possibilities in order to copy with the uncertainties.
 In \cite{Shadmand2014} the optimal storage capacity problem is addressed.
@@
 
-These are only a few possibilities for balancing production and demand in the smart grid.
-Most of the time, these techniques will be coupled with predictions of the upcoming load curves and weather conditions.
-Combining all these technologies enables aggregators to quantify their expected production and the inherent risk that comes with it.
+In addition to storage and distibuted algorithms, prediction techniques of the upcoming load curves and weather conditions are of paramount importance.
+Combining all these technologies could enable aggregators to quantify their expected production and the inherent risk that comes with it.
 The optimization of expected returns to risk is a traditional goal in finance, and a wide literature exists on this topic.
@@
 \begin{figure}
-	\includegraphics[scale=.45]{Fig2}
+	\begin{center}
+		\includegraphics[scale=.4]{Fig2}
 		\caption{Process}
 		\label{fig:process}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
+	\begin{center}
+		\begin{scriptsize}
 		\label{tab:main_notations}
@@
  		\caption{Main notations.}
+ 	\end{scriptsize}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_1}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_1}
 		\caption{{\footnotesize $ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
@@
 		\label{fig:Gaussian}
+	\end{center}
 \end{figure}
@@
 \ref{eq:alpha_star}).
-This assumption will indeed be convenient for computing some parameter estimates.
 
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_2}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_2}
 		\caption{{\footnotesize Gaussian mean approximation ($N=100,\ \phi=0.1,\ \bar{\mu}=5,\ \bar{\sigma}=.5,\ \bar{\rho}=0.2 $).
@@
 		\label{fig:mean_approx}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_3}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_3}
 		\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 		\label{fig:histo_cliques}
+	\end{center}
 \end{figure}
@@
 Figure \ref{fig:real_utility2} shows how the mean utility of a coalition evolves with its size when the optimum size is set to 40 agents.
+
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_4}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_4}
 		\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 		\label{fig:real_utility2}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_5}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_5}
 		\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space.
@@
 		\label{fig:coalitions}
+	\end{center}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[scale=.48]{./figs/figure_6}
+	\begin{center}
+		\includegraphics[scale=.46]{./figs/figure_6}
 		\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }
 		\label{fig:resilience}
+	\end{center}
 \end{figure}

### 2016-02-04 Appendix.tex
last minute changes
appendix update

@@
 
-\onecolumn
-\centering
+%\onecolumn
 \appendix
-   \subsection{Notations}
 
-	\begin{tabular}{ | l | r | }
-   						N & Number of Prosumers \\
-   						$ P_i^P(t) $ & Available production of agent i at time t \\
-   						$ P_i^D(t) $ & Consumption of agent i at time t \\ 
-  						$ P_i(t) = P_i^P(t) - P_i^D(t) $ & Available extra-production of agent i at time t \\
-  						$  WT_i $ & Wind turbine set of agent i \\
-  						$ PV_i $ & Solar array set of agent i \\
-  						$ DER_i = WT_i \cup PV_i $ & DER of agent i \\
-  						$ \nu_i(t) $ & wind speed (in $ m s^{-2} $) at agent i's location and time t \\
-  						$ \Psi_i(t) $ & solar irradiance (in $ W m^{-2} $) at agent i's location and time t \\
-  						$ \mathcal{F}_{WT} $ & power curve for the wind turbines \\
-  						$ \mathcal{F}_{PV} $ & power curve for the solar arrays \\
-   						$ P_S^{CRCT} $ & Contract value of coalition S \\
-   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
-   						$ P^{MIN} $ & Minimum Contract value \\
-   						$ \alpha $ & Parameter that controls the coalitions sizes \\
-   						$ \epsilon $ & Correlation graph filtering threshold \\
-   						$ N_{COAL} $ & Number of coalitions \\
-   						$ \mathcal{R}_S $ & Resilience of coalition S \\
-   						$ \lambda $ & unitary price rate for electricity \\
-   						$ \delta_S(i) $ & marginal utility of agent i in coalition S \\
-   						$ \rho_{ij} $ & Pearson's correlation coefficient between $ P_i$ and $P_j$ \\
-   						$ \Theta_k(G) $ & set of non overlapping cliques of size k in graph G \\
-   						$ \mathcal{R}_S $ & resilience of coalition S
- 	\end{tabular}
-
-\twocolumn
+%\twocolumn
 \subsection{Algorithms}
@@
 \end{algorithm}
+
+ \subsection{Notations}
+
+Figure \ref{fig:notations} presents all the notations used in the article.
+
+\begin{figure*}
+\begin{minipage}{\textwidth}
+\centering
+	\begin{tabular}{ | l | r | }
+   						N & Number of Prosumers \\
+   						$ P_i^P(t) $ & Available production of agent i at time t \\
+   						$ P_i^D(t) $ & Consumption of agent i at time t \\ 
+  						$ P_i(t) = P_i^P(t) - P_i^D(t) $ & Available extra-production of agent i at time t \\
+  						$  WT_i $ & Wind turbine set of agent i \\
+  						$ PV_i $ & Solar array set of agent i \\
+  						$ DER_i = WT_i \cup PV_i $ & DER of agent i \\
+  						$ \nu_i(t) $ & wind speed (in $ m s^{-2} $) at agent i's location and time t \\
+  						$ \Psi_i(t) $ & solar irradiance (in $ W m^{-2} $) at agent i's location and time t \\
+  						$ \mathcal{F}_{WT} $ & power curve for the wind turbines \\
+  						$ \mathcal{F}_{PV} $ & power curve for the solar arrays \\
+   						$ P_S^{CRCT} $ & Contract value of coalition S \\
+   						$ \phi \in \left[ 0,1 \right] $ & Reliability Threshold \\
+   						$ P^{MIN} $ & Minimum Contract value \\
+   						$ \alpha $ & Parameter that controls the coalitions sizes \\
+   						$ \epsilon $ & Correlation graph filtering threshold \\
+   						$ N_{COAL} $ & Number of coalitions \\
+   						$ \mathcal{R}_S $ & Resilience of coalition S \\
+   						$ \lambda $ & unitary price rate for electricity \\
+   						$ \delta_S(i) $ & marginal utility of agent i in coalition S \\
+   						$ \rho_{ij} $ & Pearson's correlation coefficient between $ P_i$ and $P_j$ \\
+   						$ \Theta_k(G) $ & set of non overlapping cliques of size k in graph G \\
+   						$ \mathcal{R}_S $ & resilience of coalition S
+ 	\end{tabular}
+
+\end{minipage}
+\caption{Notations}
+\label{fig:notations}
+\end{figure*}
 
@@
 
+\subsection{Utility function}
+\label{appendix_utility}
+
+We have the following utility function (see main article) that attributes a real value to any coalition S depending on the best contract value that S can propose given the grid policy $ (P^{MIN},\phi)$ :
+
+\[ U(S) = \frac{1}{|S|^{\alpha}} \frac{ P_S^{CRCT \star} }{P^{MAX}} \]
+
+$ U $ depends on the specific agents within the coalition S because $ P_S^{CRCT \star} $ is directly impacted by the correlations between the agents.
+$ U $ is also sensitive to the size of the coalition S through the term $ \frac{1}{|S|^{ \alpha } } $.
+The purpose of this term is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value tends to grow as well but the cost of coordinating all the agents also becomes larger.
+The parameter $ \alpha $ is used in order to control to what extend the size of the coalition impacts its utility.
+
+Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function.
+For some range of values, the utility might not even be a concave function of the coalition size.
+Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, we want to obtain some estimation of $ \alpha $ such that the utility function is concave, and such that the maximum lies around a desired size $ \bar{N} = \frac{N}{N_{COAL}} $.
+We thus target a local maximum in this balanced region of the space by shaping the utility function through $ \alpha$.
+\\
+In order to obtain an analytical expression for $ \alpha $, we use two simplifying assumptions.
+First, we consider the distributions to be normal distributions, such that the optimal contract value can be written in the following way :
+
+\[ P_S^{CRCT \star } = \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \]
+
+Therefore, we can writte the utility function as :
+
+\[ U(S) = \frac{1}{|S|^{\alpha} P^{MAX} } \left[ \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \right] \]
+
+With :
+
+\[ \left\{ \begin{array}{lll} \mu_{S} = \sum_{i \in S} \mu_i \\ \sigma_{S} = \sqrt{ \sum_{i \in S} \sigma_i^2 + 2 \sum_{i} \sum_{j} \rho_{ij} \sigma_i \sigma_j } \end{array} \right. \]
+
+We then make a second assumption for simplifying this expression : we consider that :
+
+\[ \left\{ \begin{array}{lll} \forall i \in \mathcal{A}\ \mu_i = \bar{ \mu } = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i  \\ \forall i \in \mathcal{A}\ \sigma_i = \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i \\ \forall i,j \in \mathcal{A}^2\  \rho_{ij} = \bar{ \rho } = \frac{2}{N(N-1)} \sum_{i \in \mathcal{A}} \sum_{j \in \mathcal{A}} \rho_{ij} \end{array} \right.\]
+
+That is, any agent is replaced by a mean approximation. In these conditions, we have :
+
+\[ U(S) = \frac{ \bar{ \mu } }{ |S|^{\alpha -1} P^{MAX} } - \frac{ \sqrt(2) \bar{ \sigma } }{ |S|^{\alpha} P^{MAX} } \sqrt{|S| + \bar{ \rho } |S| ( |S| - 1 ) } \]
+
+$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $.
+Depending on $ \alpha $ (and $ \bar{ mu },\ \bar{ \sigma },\ \bar{ rho }, and\ P^{MAX} $), the utility function can have very different shapes.
+\\
+We find $\alpha^{\star} $ by solving :
+
+\[ \left\{ \begin{array}{lll} \left| \frac{ \partial U }{ \partial |S| } \right|_{|S|=\bar{N}} = 0 \\ \left| \frac{ \partial^2 U }{ \partial |S|^2 } \right|_{\alpha = \alpha^{\star}} \leq 0 \end{array} \right.
+\]
+
+Solving this equation with a formal calculus software yields the equation of the paper.
+
 \bibliographystyle{IEEEtran}  
 \bibliography{Article}
-
 \end{document}

### 2016-02-05 Appendix.tex
appendix+V-VI

--- 2015-11-19 Appendix.tex
@@
 
+\subsection{Utility function}
+\label{appendix_utility}
+
+We have the following utility function (see main article) that attributes a real value to any coalition S depending on the best contract value that S can propose given the grid policy $ (P^{MIN},\phi)$ :
+
+\begin{equation}
+	U(S) = \frac{1}{|S|^{\alpha}} \frac{ P_S^{CRCT \star} }{P^{MAX}}
+\end{equation}
+
+$ U $ depends on the specific agents within the coalition S because $ P_S^{CRCT \star} $ is directly impacted by the correlations between the agents.
+$ U $ is also sensitive to the size of the coalition S through the term $ \frac{1}{|S|^{ \alpha } } $.
+The purpose of this term is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value $ P_S^{CRCT \star} $ tends to grow as well but the cost of coordinating all the agents also becomes larger.
+The parameter $ \alpha $ is used in order to control to what extend the size of the coalition impacts its utility.
+
+Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function.
+For some range of values, the utility might not even be a concave function of the coalition size.
+Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, the algorithm \ref{alg:algo1} will start with $ N_{COAL} $ seeds and increase their sizes as long as the utilities are improved.
+The order through which coalitions incorporate new prosumers is randomized but leads generally to relatively balanced coalitions.
+We denote by $ \bar{N} $ the desired approximate final size of the coalitions, i.e we target a utility function that is concave in the coalition size with a maximum at $ \bar{N} $ (in the article, $ \bar{N} $ is fixed at $\frac{N}{N_{COAL}}$).
+We propose to obtain such utility function through tunning $\alpha $.
+Ideally, we seek to express $ \alpha $ as a function of $ \bar{N} $ and the agents power distributions and correlations.
+\\
+In the general case, neither the power distributions nor the correlation structure have special forms which makes obtaining an analytical expression for $ \alpha $ difficult.
+Therefore we use two simplifying assumptions in order to have an estimate for $ \alpha $.
+First, we approximate the power distributions by normal distributions, such that the optimal contract value can be written in the following way :
+
+\begin{equation}
+	P_S^{CRCT \star } = \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi)
+\end{equation}
+
+With :
+
+\begin{equation}
+	\left\{ \begin{array}{lll} \mu_{S} = \sum_{i \in S} \mu_i \\ \sigma_{S} = \sqrt{ \sum_{i \in S} \sigma_i^2 + 2 \sum_{i} \sum_{j} \rho_{ij} \sigma_i \sigma_j } \end{array} \right.
+\end{equation}
+
+Therefore, we can write the utility function as :
+
+\begin{equation}
+	U(S) = \frac{1}{|S|^{\alpha} P^{MAX} } \left[ \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \right]
+\end{equation}
+
+As $ \mu_S $ and $ \sigma_S $ depends on the agents in S (and consequently on the size of S), we make a second simplifying assumption :
+
+\begin{equation}
+	\left\{ \begin{array}{lll} \forall i \in \mathcal{A}\ \mu_i = \bar{ \mu } = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i  \\ \forall i \in \mathcal{A}\ \sigma_i = \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i \\ \forall i,j \in \mathcal{A}^2\  \rho_{ij} = \bar{ \rho } = \frac{2}{N(N-1)} \sum_{i \in \mathcal{A}} \sum_{j \in \mathcal{A}} \rho_{ij} \end{array} \right.
+\end{equation}
+
+That is, any agent is replaced by a mean approximation. In these conditions, we have :
+
+\begin{equation}
+	\left\{ \begin{array}{lll} 
+				\mu_S = |S| \bar{\mu} \\
+				\sigma_S = \bar{\sigma} \sqrt{|S| \left[ 1 + \bar{\rho} (|S|-1) \right] }	
+	\end{array} \right.
+\end{equation}
+
+And the utility function becomes :
+
+\begin{equation}
+	U(S) = \frac{ \bar{ \mu } }{ |S|^{\alpha -1} P^{MAX} } - \frac{ \sqrt(2) \bar{ \sigma } }{ |S|^{\alpha} P^{MAX} } \sqrt{|S| + \bar{ \rho } |S| ( |S| - 1 ) }
+\end{equation}
+
+$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $.
+Depending on $ \alpha $, the utility function can have very different shapes.
+\\
+We select $\alpha^{\star} $ such that :
+
+\begin{equation}
+	\left\{ \begin{array}{lll} \left| \frac{ \partial U }{ \partial |S| } \right|_{|S|=\bar{N}} = 0 \\ \left| \frac{ \partial^2 U }{ \partial |S|^2 } \right|_{\alpha = \alpha^{\star}} \leq 0 \end{array} \right.
+\end{equation}
+
+Solving these equations yields the following expression :
+
+\begin{equation}
+	\begin{tiny}
+		\alpha^{\star} ( \bar{N} ) = \dfrac{0.7 \bar{\sigma}(\bar{\rho}-1)erf^{-1}(2 \phi - 1)}{\bar{\mu}\sqrt{\bar{N}(\bar{\rho}\bar{N}-\bar{\rho}+1)}+1.4 \bar{\sigma} erf^{-1} (2 \phi -1 ) (\bar{\rho}\bar{N}-\bar{\rho} + 1)}
+	\end{tiny}
+\end{equation}
+
+Since $ \alpha^{\star} ( \bar{N} ) $ is obtained for a simplified case derived from the real situation, it should be considered as a soft constraint on the coalitions sizes.
+
+Using such an approach on 100 prosumers examples with $N_{COAL} = 4 $ and $\bar{N}=25$ results in coalitions which sizes are comprised between 17 and 34 agents.
+
 \bibliographystyle{IEEEtran}  
 \bibliography{Article}
-
 \end{document}

### 2016-02-05 Article.tex
appendix+V-VI

@@
 In the general case where distributions and correlation structures have no special form, finding an analytical expression for $ \alpha $ appears complicated.
-To overcome this problem, we express $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $.
+To overcome this problem, we seek an approximation for $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $.
 In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details).
-We select $ \alpha^{\star} $ such that :
+We then select $ \alpha^{\star} $ such that :
 
@@
 
-Which yields to the following expression for $ \alpha^{\star} $ (see Appendix) :
+Which leads to the following expression for $ \alpha^{\star} $ (see Appendix) :
 
@@
 Figure \ref{fig:mean_approx} shows how $ \alpha^{\star} $ and the utility function evolves according to the mean size of the coalitions $ \bar{N} $.
-As can be pointed out, the purpose of $ \mathcal{U} $ is not to study coalitions stability against player defection which could be done through game theory, nor to redistribute the coalition's utility in terms of individual payoffs.
-$ \mathcal{U} $ is rather designed as a quality measure for the coalitions.
+Given a policy $ (\phi, P^{MIN} ) $ and a pool of N prosumers, we are now able to quantify the quality of any coalition S.
 
@@
 \label{sec:forming}
-Section \ref{sec:notations} explained how contract values for the coalitions are computed, and in section \ref{sec:utility} we related this quantity to the utility and gains of a coalition.
-Since the computation time of this quantity is not negligible, we derive in this section the heuristic we used to form the coalitions structure.
+We aim at forming $ N_{COAL} $ coalitions $ \{ S_1;S_2;...;S_{N_{COAL}} \} $ such that the global utility $ U = \sum_{i = 1}^{N_{COAL}} \mathcal{U}(S_i) $ is maximized.
+Since achieving such a goal is a well-known hard problem due to the combinatorial explosion in the number of possible aggregations, we propose in this section a greedy heuristic using decorrelation graphs to solve the problem.
 

### 2016-02-05 Article.tex
respond letter

@@
 To overcome this problem, we seek an approximation for $ \alpha $ in a simplified situation where all power distributions are approximated by normal distributions and all prosumers are considered equivalent to a mean agent : $ \forall i \in \mathcal{A},\ P_i \sim \mathcal{N}(\bar{\mu}, \bar{\sigma}) $, where $ \bar{\mu} = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i $, $ \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i $, and $\forall i,j,\ \rho_{ij} = \bar{\rho} $.
-In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix for more details).
+In this simplified case, we can express the utility as a function of $ \bar{\mu} $, $ \bar{\sigma} $, $\bar{\rho} $, $ \bar{N} $, and $ \alpha $ (see Appendix).
 We then select $ \alpha^{\star} $ such that :
@@
 
-Which leads to the following expression for $ \alpha^{\star} $ (see Appendix) :
+Which leads to (see Appendix) :
 
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_3}
+		\includegraphics[scale=.45]{./figs/figure_3}
 		\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_4}
+		\includegraphics[scale=.45]{./figs/figure_4}
 		\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_5}
+		\includegraphics[scale=.45]{./figs/figure_5}
 		\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space.
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_6}
+		\includegraphics[scale=.45]{./figs/figure_6}
 		\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }
@@
 We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.
+A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.
 

### 2016-02-05 review.tex
respond letter

\documentclass{letter}
\makeatletter
\newenvironment{thebibliography}[1]
     {\list{\@biblabel{\@arabic\c@enumiv}}%
           {\settowidth\labelwidth{\@biblabel{#1}}%
            \leftmargin\labelwidth
            \advance\leftmargin\labelsep
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
      \sloppy
      \clubpenalty4000
      \@clubpenalty \clubpenalty
      \widowpenalty4000%
      \sfcode`\.\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
\usepackage[]{algorithm2e}
\newcommand\newblock{\hskip .11em\@plus.33em\@minus.07em}
\makeatother
\signature{Vincent Gauthier}
\address{Vincent Gauthier \\ Institut Mines Telecom \\ Nano Innov, Saclay \\ France}
\begin{document}

\begin{letter}{Dr. Jianhui Wang \\ Editor-in-Chief \\ Journal Transactions on Smart Grid } % Addressee name and address

\opening{ Dear Dr. Jianhui Wang,}

The authors would like to thank you for giving us the opportunity to revise and resubmit this manuscript. We resubmit this revision before the agreed upon deadline, March 29, 2016. We appreciate the time and details provided by each reviewer and by you and have incorporated the suggested changes into the manuscript to the best of our ability. The manuscript has certainly benefited from these insightful revision suggestions.

See bellow our replies to the reviewers’ comments:

\begin{itemize}

	\item \textbf{Reviewer: 1}

Thank you for addressing my comments and suggestions, the revised manuscript is improved and I have no more questions/comments. I recommend this paper for publication in IEEE Transactions on Smart Grid.

	\item \textbf{Reviewer: 2}

The paper has been further improved, but several unclear points remain. These are listed next, but a general comment is that the contents of Sections V and VI are not developed in a clear manner, and the link between the two sections is hardly described.
	\begin{itemize}

		\item The development following eq. (12) until the end of Section V is very confusing. Here are some of the confusing points, but overall, the previously mentioned part needs lots of explanations. Do all coalitions have the same size? (It seems yes, based on $ \bar{N} $, but this seems to be an unjustified arbitrary choice.) What is the logic behind the derivative in (13)? The utility depends on the members of the coalition. In other words, two coalitions of the same size may have different utilities, depending on the correlations between the participating prosumers. Thus taking the derivative to be zero does not seem to be the correct approach to maximize the utility. Which quantities are replaced by their mean values? In a system with N prosumers, there is a correlation matrix of dimension N x N. Is it implied that all pairwise correlations in this matrix are substituted by $ \bar{rho} $ ? The derivation and all symbols of eq. (14) must be explained. Is it always the case that the utility is a concave function of the number of agents, as depicted in Fig. 4? (It seems very unlikely.)
\\

\textbf{Answer :} The end of section V is devoted to propose a method to select the parameter $ \alpha $ properly. The purpose of this parameter is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value tends to grow as well but the cost of coordinating all the agents also becomes larger. We wish to obtain a concave function in the coalition size such that adding agents increases the utility in the begining of the process while showing diminishing returns after a given point. Therefore, we select $ \alpha $ such that the utility is concave and exhibits a maximum "around" (because $ \alpha $ is selected based on approximations) $\bar{N} $. The resulting coalitions do not have strictly the same size but tend to partition agents in relatively balanced groups. As pointed by the reviewer, agents do not increase the utility of a coalition by the same amount (see $P_S^{CRCT \star}$), but the size cost incurred do not depend on the agents.

We agree with the reviewer that this part of the paper may lack some precisions. We therefore reformulated this part of the paper as well as the transition from part V to part VI. Because of space limitation, we also added more precisions about the utility function to the appendix (see the "\textit{utility function}" section of the appendix). For convenience we summurize this section here :
\\
We have the following utility function that attributes a real value to any coalition S depending on the best contract value that S can propose given the grid policy $ (P^{MIN},\phi)$ :

\[ U(S) = \frac{1}{|S|^{\alpha}} \frac{ P_S^{CRCT \star} }{P^{MAX}} \].

We completely agree with the reviewer that this utility function depends on the specific agents within the coalition S. Nevertheless, the utility is also sensitive to the size of the coalition it evaluates, and this effect is tunned through the $ \alpha $ parameter. The purpose of this parameter is to introduce in the utility function the following tradeoff : as a coalition grows, its contract value tends to grow as well but the cost of coordinating all the agents also becomes larger.

However, knowing what value to use for $ \alpha $ is not trivial and it is important since it influences the shape of the utility function. As mentionned by the reviewer, for some range of values, the utility might not even be a concave function of the coalition size. Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, we want to obtain some estimation of this $ \alpha $ parameter such that the utility function is concave, and such that the maximum lies around a desired size $ \bar{N} = \frac{N}{N_{COAL}} $. We thus target a local maximum in this balanced region of the space by shaping the utility function through $ \alpha$. In other words, we want to favor structures of balanced coalitions such that, if we have, for instance, 100 agents and we desire 4 coalitions, the algorithm will tend to form coalitions around 25 agents each (but not necessarilly all equals to 25). We agree that this choice is debatable since a completely unbalanced structure with 97 agents in one coalition and 3 other coalitions of 1 agent each could perhaps achieve a better global utility if we were not using $\alpha$. Nevertheless, we believe that such cases could lead to potentially risky aggregations of agents.
\\
As stated above, we want an approximation for $ \alpha $ such that we use 2 simplifying assumptions as to obtain an analytical expression for it.
\\
First, we consider the distributions to be normal distributions (as in the case introduced in the previous section), such that the optimal contract value can be written in the following way :

\[ P_S^{CRCT \star } = \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \]

Therefore, we can writte the utility function as :

\[ U(S) = \frac{1}{|S|^{\alpha} P^{MAX} } \left[ \mu_S - \sqrt{2} \sigma_S erf^{-1}(1 - 2 \phi) \right] \]

With :

\[ \left\{ \begin{array}{lll} \mu_{S} = \sum_{i \in S} \mu_i \\ \sigma_{S} = \sqrt{ \sum_{i \in S} \sigma_i^2 + 2 \sum_{i} \sum_{j} \rho_{ij} \sigma_i \sigma_j } \end{array} \right. \]

We then make a second strong assumption for simplifying this expression : we consider that :

\[ \left\{ \begin{array}{lll} \forall i \in \mathcal{A}\ \mu_i = \bar{ \mu } = \frac{1}{N} \sum_{i \in \mathcal{A}} \mu_i  \\ \forall i \in \mathcal{A}\ \sigma_i = \bar{ \sigma } = \frac{1}{N} \sum_{i \in \mathcal{A}} \sigma_i \\ \forall i,j \in \mathcal{A}^2\  \rho_{ij} = \bar{ \rho } = \frac{2}{N(N-1)} \sum_{i \in \mathcal{A}} \sum_{j \in \mathcal{A}} \rho_{ij} \end{array} \right.\]

That is, any agent is replaced by a mean approximation. In these conditions, we have :

\[ U(S) = \frac{ \bar{ \mu } }{ |S|^{\alpha -1} P^{MAX} } - \frac{ \sqrt(2) \bar{ \sigma } }{ |S|^{\alpha} P^{MAX} } \sqrt{|S| + \bar{ \rho } |S| ( |S| - 1 ) } \]

$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $. Depending on $ \alpha $ (and $ \bar{ mu },\ \bar{ \sigma },\ \bar{ rho }, and\ P^{MAX} $), the utility function can have very different shapes.
\\
We find $\alpha^{\star} $ by solving :

\[ \left\{ \begin{array}{lll} \left| \frac{ \partial U }{ \partial |S| } \right|_{|S|=\bar{N}} = 0 \\ \left| \frac{ \partial^2 U }{ \partial |S|^2 } \right|_{\alpha = \alpha^{\star}} \leq 0 \end{array} \right. \]

Solving this equation with a formal calculus software yields the equation of the paper.
\\
Please note, that we do not solve the coalition problem by taking the derivative of the utility function to be zero, we only use this in order to obtain a reasonable guess for $ \alpha $.
\\

		\item It would be very helpful if the numerical example that Fig. 4 is based on is fully described. How many prosumers are considered? What is their correlation?

		\textbf{Answer :} We added information in the caption of figure 4. The curves are plotted for a situation with $N=100$ prosumers and $ \phi = 0.1 $, such that $ \bar{\mu} = 5 $, $ \bar{\sigma} = 0.5 $, and $ \bar{\rho} = 0.2 $.

		\item What is the connection between Sections V and VI?

		\textbf{Answer :} We present in section VI the method used in order to optimize the utility function presented in section V.

		\item Why is the correlation graph introduced in Section VI? It appears that all developments are actually utilizing the decorrelation graph.
		\textbf{Answer :} We introduced the concept of decorrelation graphs in section VI by first citing existing works on correlation graphs. Furthermore, we use correlation graphs as a comparison with our method (see "correlated coalitions" in  the results section).

		\item The last sentence of Section VI-A is confusing.

		\textbf{Answer :} We removed this sentence from the paper.

		\item The numerical setup used to generate Fig. 5 needs to be explained.

		\textbf{Answer :} We added information in the caption of figure 5. The histograms show utility values for triplets (triangles of decorrelation graph in red, other triplets in blue). The total number of prosumer is $N=200$. $ \phi = 0.1 $, $ \bar{\mu} = 3.9 $MW, $ \bar{\sigma} = 1.9 $MW, $ \bar{\rho} = 0.69 $, $\alpha^{\star} = 0.08 $.

		\item Are the clusters produced by the algorithms in Section VI sensitive to the selected parameters $ N_{COAL} $ and k?

		\textbf{Answer :} The clusters are sensible to $ N_{COAL} $ (number of coalitions).

		\item The numerical setup for Fig. 6 must be explained. In particular, how are real data mapped to the quantities of eq. (14)?

		\textbf{Answer :} We added information in the caption of figure 6. The total number of prosumer is $N=200$, $ \phi = 0.1 $, $ \bar{\mu} = 3.9 $MW, $ \bar{\sigma} = 1.9 $MW, $ \bar{\rho} = 0.69 $. On this figure, we show an example where $ \bar{N} = 40 $, such that $ \alpha^{\star} = 0.006$ from eq. 14.

	\end{itemize}

	\item \textbf{Reviewer: 3}

The revised version of the paper has been improved significantly. However, some questions are from the reviewers are still not addressed completely and need further clarification as follows:

	\begin{itemize}

		\item Scientifically, you can neglect the agent freedom to choose what coalition to join. However, an active end-user would not join this demand response program, unless a suitable collective incentive will be created. As mentioned, some game-theory frameworks, such as [1], can be an option that should be referred in the related work section for the sake of completeness.

		\item In the related work section, two first paragraphs and citied works there are not much relevant to this research, while some relevant works are missing (as mentioned with game-theory, but can be other methods).

		\item Making a good order of references, i.e. you start with Introduction part and ref [5].

		\textbf{Answer :}  We followed the advices of the reviewer by incorporating the suggested reference, and making a better order of references.

		\item It is a bit confuse to mention that grid operator plays a role to in the market system. Are you considering grid operator as a balance responsible party who contributes to solve the imbalance of the whole system? And what a coalition can offer here is an optimal portfolio for such party to minimize the risk of having over/under-production/consumption?!

		\textbf{Answer :} In this work, the grid operator is only responsible for selecting the minimum production threshold $P^{MIN}$ and the reliability threshold $ \phi $ whom coalitions should comply with. Furthermore, the consumption production balance over the grid is not addressed here. We indeed focus on the formation of productive coalitions that propose some production capacity over a period of time. In a more complex and maybe more realistic framework, we could imagine that these parameters ($P^{MIN}$ and $ \phi $) are selected by some market process, or that the electricity price depends on the uncertainty of the production, such that power from unreliable coalitions is cheaper than power from reliable ones.

		\item At this stage, the research shows a nice and promising mathematical method. However, it misses in the validation part a more realistic use case/scenario to test. At least, a better link between between the proposed model to the possible considered test system should be there.

		\textbf{Answer :} In this work, we used real weather data as to be confronted with the problems that come with it (noise, non-gaussianity, and so on...). However, prosumers are not configured according to real data (their positions, number of DER owned and so on are drawn from random distributions). A more realistic work would perhaps incorporate to this model data from real aggregators and market conditions. Some information concerning the European market and wind power can be found in \cite{Europe}. However, we believe that such improvements would require extensive work and is out of the scope of this paper. We added the following text to the end of the conclusion :

		"\textit{A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions}".

	\end{itemize}

\end{itemize}
\bibliographystyle{unsrt}
\bibliography{review}

\end{letter}
\end{document}

### 2016-02-11 Article.tex
intro modifs

@@
 This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
-As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk.
+As aggregators are bound to the market operator by a contract, they seek to maximize their offer while minimizing their risk.
 The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
@@
 Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition.
-Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages [5].
-In this paper, we focus on these "nodes" in the distribution network that can produce and consume electricity.
-More precisely, we consider agents owning both renewable distributed energy resources (DER) and electrical loads.
-The production of an agent can be used to meet its own demand, but in cases when it is over-producing, we consider that it has the possibility to inject, against financial compensation, its extra-production in the grid.
-Such an agent model is known as a "prosumer" [6].
-A key point of this work is to merge the interests of the grid operator with the prosumers in a single utility function.
-While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the injections.
+Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.
 
-%The coalitions that we wish to form should thus be both stable and productive.
-This compromise can be difficult to obtain in real situations, especially when renewable generation and complex spatiotemporal correlation happen.
+In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads.
+We will refer to these agents as prosumers \cite{Rathnayaka2012} in the future.
+If a prosumer produces more than he consumes, he might be willing to sell power.
+We thus assume the existence of an energy market where producers post a production capacity for an upcoming period, and buyers decide what bids to purchase.
+When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account.
+Rather than having a risk dependant price, we propose here a safer and simpler model.
+We assume the existence of a market operator who constrains the market entrance.
+A key point of this work is to merge the interests of the market operator with the prosumers in a single utility function.
+While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.
 
-%More precisely, in our model, any entity that wishes to sell its production on some energy market has to estimate and announce a production capacity for an upcoming period of time.
-If a contract is bought, the entity commits to injecting, at any time of the contracted period, the announced amount of power, under financial penalties if it fails.
-Since prosumers use exclusively renewables, contracts come naturally with some amount of risk due to the intermittent nature of most renewables.
-Using storage as buffers as to reduce risks is a popular approach.
-Without storage and proper control, the over-producing state of the agents, and thus the power they inject in the grid, might be rather unstable.
-This is unacceptable for the grid operator that cannot ensure system stability if it has to deal with numerous small unpredictable entities.
+If a buyer purchases a contract, the corresponding producer has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates.
+Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value.
+In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.
+Single prosumers might be authorized to enter such market, but it is unlikely that a single agent can propose enough power with enough reliability.
 
-We rely on a market representation of the interactions.
-Coalitions of prosumers estimate a production capacity for an upcoming period and announce it on the market.
-If some other entity buys the contract, the coalition has to inject exactly the contracted amount of power at any time.
-Obviously, deviations will cause financial penalties to the coalitions, such that respecting the contracts is paramount.
-Since prosumers use exclusively renewables, we are particularly concerned with the probability that the available capacity of a coalition falls below its contract value.
-In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.
-
-We focus on coalitions because it appears as a solution to both the number of prosumers communicating with the operator as well as their high variabilities.
-Indeed, it has been argued that using a multi-level aggregators architecture for the control could lead to better performances on the communication side \cite{Negeri2012}.
-On the other hand, it is known that diversification of the assets is a way of minimizing risks when building a portfolio.
-We thus expect more stable and predictable energy productions for coalitions than for single agents.
+Therefore, we use a coalitional framework in order to model aggregators of prosumers.
+It is known that diversification of the assets is a way of minimizing risks when building a portfolio.
+We thus expect more stable and predictable energy productions for aggreagtors than for single agents.
 Nevertheless, all coalitions are not equally stable or productive.
-Special attention should thus be paid to the aggregation step.
+Attention should thus be paid to the aggregation step.
 Recall that prosumers have both consumption and production components, each depending on location and time, meaning that there are complex underlying correlations between the agents.
@@
 
-We will see that the variability in the aggregated productions can be quantified to a certain extent by the correlation among the agents forming the coalitions.
-Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell.
+We will see that the variability in the production of the aggregators can be quantified to a certain extent by the correlation among the agents within the coalitions.
+Understanding the correlation relationships among the agents can thus give an indication about what coalitions to form and how much they should sell on the market.
 More precisely, we use a graph representation of the correlation structure to gain insight about the expected production to risk ratios of different coalitions.
-We build a framework in which the system operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to accept.
+We build a framework in which the market operator specifies both the minimum production acceptable to enter the market and restrain the amount of risk he is willing to accept.
 We then propose a graph heuristic that uses decorrelated cliques in order to form diversified productive and stable coalitions, and we compare the results with other formation strategies (see section \ref{sec:results}).
-Rather than maximizing the profit of each agent in the grid (prosumers and grid operator), our algorithm tries to form the most productive coalitions given a maximum amount of acceptable risk.
-In other words, it tries to maximize coalitions profits without considering individual retributions to single agents.
+Rather than maximizing the profit of each agent in the grid (prosumers and operator), our algorithm tries to form the most productive coalitions given a maximum amount of acceptable risk.
+In other words, it tries to maximize aggregators' profits without considering individual retributions to single agents.
 Furthermore, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.
@@
 
-Assuming a large penetration of DER, it is expected that the managment of the system would have to evolve in order to copy with the large amount of unpredictable small generators.
+Assuming a large penetration of DER, it is expected that the managment of the system would have to evolve as to copy with the large amount of unpredictable small generators.
 In \cite{Zhang2013}, the authors introduce a distributed energy management system with a high use of renewables such that power is scheduled in a distributed fashion.
@@
 Selecting the right contract value appears thus as an interesting problem on its own that we plan to investigate in future works.
-In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the grid operator.
-The role of the grid operator is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
-More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the grid operator as a maximum value for the probability of under-producing.
+In order to keep the present paper in a reasonable length, we simplify the contract value selection problem by giving some responsibilities to a third party named the market operator.
+His role is to constrain the market entry to coalitions able to propose both sufficiently high and sufficiently credible contract values.
+More formally, let $ \phi \in [0,1] $ be the reliability threshold fixed by the market operator as a maximum value for the probability of under-producing.
 The highest contract value that a coalition can propose is thus $ P_{S}^{CRCT \star} $ such that $ Pr[P_{ij} \leq P_{S}^{CRCT \star}] = \phi $.
@@
 
-This is the best contract value that the coalition S can afford for a given stability policy $\phi$ of the grid operator.
+This is the best contract value that the coalition S can afford for a given stability policy $\phi$ of the market operator.
 Figure \ref{fig:Gaussian} shows how $ P_{S}^{CRCT \star} $ evolves according to the reliability parameter $ \phi $.
@@
 As visible on figure \ref{fig:Gaussian}, coalitions with high expected productions but presenting a high unpredictability are penalized and can only afford small contracts.
-The grid operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions.
+The market operator also specifies a lower bound $ P^{MIN} $ on the contract values as not to overload the market with unrealistic small coalitions.
 We thus characterized a valid coalition as one satisfying the two conditions :
@@
 	\item \textbf{The agents :} historical series of available productions $P_{i}$, 
-	\item \textbf{The grid operator :} market entrance policy $ (P^{MIN},\phi) $,
+	\item \textbf{The market operator :} market entrance policy $ (P^{MIN},\phi) $,
 	\item \textbf{The "user" :} Number of desired coalitions $ N_{COAL} $ and size of starting cliques k.
@@
 Therefore, the notion of resilience we will use in the following can be seen as the ability of the coalition structures to inject stable power in the grid when some of its internal agents are removed.
-According to our model, the grid operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $.
+According to our model, the market operator specified two thresholds ($P^{MIN}$ and $ \phi $) such that the power injected by every coalition is constrained : $ P_{S}^{CRCT} \in [P^{MIN}, P_{S}^{CRCT \star}] $.
 As long as a coalition can propose a contract value higher than $ P^{MIN} $, it is valid and allowed to enter the energy market.
@@
 
-In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production to the grid operator.
+In this paper we studied how aggregations of prosumers could be authorized to sell their surplus of production on the energy market.
 By relying on the past values of the agents, we constrained the market entry to both sufficiently productive and stable coalitions.

### 2016-02-17 review.tex
respond change

@@
 		\textbf{Answer :} In this work, the grid operator is only responsible for selecting the minimum production threshold $P^{MIN}$ and the reliability threshold $ \phi $ whom coalitions should comply with.
-Furthermore, the consumption production balance over the grid is not addressed here.
-We indeed focus on the formation of productive coalitions that propose some production capacity over a period of time.
-In a more complex and maybe more realistic framework, we could imagine that these parameters ($P^{MIN}$ and $ \phi $) are selected by some market process, or that the electricity price depends on the uncertainty of the production, such that power from unreliable coalitions is cheaper than power from reliable ones.
+We agree that the terminology used could be misleading, so, as to avoid any confusion, we renamed this entity "market operator".
+Prosumers are grouped in coalitions managed by aggregators that whish to sell on the energy market.
+In order to do so, they have to respect the conditions fixed by the "market operator".
+The balance between production and consumption in the grid is not addressed here.
+However energy markets, as depicted in this work, could very well be some way, for the actual grid operator, of preserving this equilibrium.
+As to make sure that both the scenario and the limitations of our work are clear, we clarify these points in the begining of the introduction.
 

### 2016-02-17 Article.tex
modifs intro

@@
 
-The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system.
-The rigid top-down centralized architecture of today’s power grids is expected to evolve in more distributed and flexible systems.
-Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition.
+%The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system.
+
+%Bidirectional power flows, storage, and distributed production are characteristics of smart grid systems.
+Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads depending on weather conditions.
+
+%Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition.
 Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.
@@
 We assume the existence of a market operator who constrains the market entrance.
-A key point of this work is to merge the interests of the market operator with the prosumers in a single utility function.
-While the prosumers intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.
 
-If a buyer purchases a contract, the corresponding producer has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates.
+As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets.
+A key point of this work is to merge the interests of the market operator and aggregators in a single utility function.
+While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.
+
+If a buyer purchases a contract, the corresponding aggregator has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates.
 Since prosumers use renewables, we are concerned with the probability that the production falls below the contract value.
 In this case, storage or demand side managment might be solutions to avoid penalties, but this probability should still be maintained low.
-Single prosumers might be authorized to enter such market, but it is unlikely that a single agent can propose enough power with enough reliability.
 
-Therefore, we use a coalitional framework in order to model aggregators of prosumers.
+In this paper, we decided to use a coalitional framework in order to model aggregators.
 It is known that diversification of the assets is a way of minimizing risks when building a portfolio.

### 2016-02-17 Article.tex
intro modifs

@@
 \begin{abstract}
-Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
-It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system.
-All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
-In this paper, we propose an algorithm that forms coalitions of agents, called prosumers, that both produce and consume.
-It is designed to be used by aggregators that aim at selling the aggregated surplus of production of the prosumers they control.
+In the context of the smart grid, we propose in this paper an algorithm that forms coalitions of agents, called prosumers, that both produce and consume.
+It is designed to be used by aggregators that aim at selling aggregated surplus of production of the prosumers they control.
 We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each prosumer.
-This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
+This enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors.
 As aggregators are bound to the market operator by a contract, they seek to maximize their offer while minimizing their risk.
 The proposed graph based algorithm takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability.
-We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
+We show that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
 \end{abstract}
@@
 
+Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century.
+It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with the use of small renewable generators and a proper efficient information system.
+All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation.
+
 In this paper, we consider agents in the distribution network that own both renewable distributed energy resources (DER) and loads.
@@
 When the production is stochastic, as it can be with DER, this approach gets more complicated and risk should be quantified and taken into account.
-Rather than having a risk dependant price, we propose here a safer and simpler model.
-We assume the existence of a market operator who constrains the market entrance.
-
+We thus assume the existence of a market operator who constrains the market entrance.
 As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets.
-A key point of this work is to merge the interests of the market operator and aggregators in a single utility function.
-While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply.
-
+While there exists several models for fixing prices and organizing bids in such stochastic settings, the present work does not focus on transactions on the market.
+Instead, we study how aggregators can manage their portfolios as to maximize the production that they offer, given that such production is stochastic by nature.
+A key point of this work is then to merge the interests of the market operator and aggregators in a single utility function.
+While the aggregators intend to maximize their benefits with better contracts, the operator concerns are related to the quality and reliability of the market supply (more information about the scenario can be found in \cite{Europe} page 67-68).
 If a buyer purchases a contract, the corresponding aggregator has to inject the contracted amount of power at any time of the contract, exposing himself to financial penalties if he deviates.
@@
 	\begin{center}
-		\includegraphics[scale=.4]{Fig2}
+		\includegraphics[scale=.39]{Fig2}
 		\caption{Process}

### 2016-02-18 Article.tex
update
corrections Monique

@@
 As single agents might not be productive or reliable enough, aggregators managing large portfolios of prosumers are more likely to enter such markets.
-While there exists several models for fixing prices and organizing bids in such stochastic settings, the present work does not focus on transactions on the market.
+While there exist several models for fixing prices and organizing bids in such stochastic settings, the present work does not focus on transactions on the market.
 Instead, we study how aggregators can manage their portfolios as to maximize the production that they offer, given that such production is stochastic by nature.
@@
 
-In this paper, we decided to use a coalitional framework in order to model aggregators.
+We use a coalitional framework in order to model aggregators.
 It is known that diversification of the assets is a way of minimizing risks when building a portfolio.
-We thus expect more stable and predictable energy productions for aggreagtors than for single agents.
+We thus expect more stable and predictable energy productions for aggregators than for single agents.
 Nevertheless, all coalitions are not equally stable or productive.
@@
 Because agents are susceptible to fail for diverse reasons, the propensity of the system to undertake these failures is critical \cite{Pahwa}.
-We therefore investigate in this paper the resilience of the coalitions when prosumers fail.
+We also investigate in this paper the resilience of the coalitions when prosumers fail.
 Despite the fact that losing agents is usually detrimental to the coalitions, we will see that the coalitions formed with our algorithm tend to be less impacted by random failures.
@@
 We showed that, in this situation, the coalitions resulting from our algorithm better withstand losses of agents.
-A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators as to propose a prosumer aggregation model closer to real conditions.
+A possible direction for future work could be to use, in addition to real weather data, data from energy markets and aggregators' portfolios as to propose a prosumer aggregation model closer to real conditions.
 

### 2017-05-10 Appendix.tex
latest changes

@@
 
-%\onecolumn
 \appendix
 
-%\twocolumn
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%            APPENDIX A : ALGORITHMS
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Algorithms}
@@
 \begin{figure}[b]
-\includegraphics[scale=.5]{./figs/figure_7}
+\includegraphics[scale=.5]{figure_8.eps}
 \caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%            APPENDIX B : NOTATIONS
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Notations}
 
-Figure \ref{fig:notations} presents all the notations used in the article.
+\noindent Figure \ref{fig:notations} presents all the notations used in the article.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%            APPENDIX C : NET PRODUCTION SERIES
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Net production series}
@@
 Power curves are functions that, for a given type of generator, map some input quantity to the output power produced.
-For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007, Piedallu2008}.
+For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007}, \cite{Piedallu2008}.
 For the wind turbines, the power curve can be specified by 4 values :
@@
 The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data.
-As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007, Piedallu2008}.
+As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007}, \cite{Piedallu2008}.
 This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps.
-As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007, Piedallu2008} for more details) :
+As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007}, \cite{Piedallu2008} for more details) :
 
@@
 \end{equation}
-where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
+
+\noindent where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t.
 Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :
@@
 \end{equation}
-where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
+
+\noindent where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency.
 The very simple form of this power curve is due to some simplifications in order not to overload the model.
@@
 \end{equation}
-where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
+
+\noindent where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day.
 In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification.
@@
 \end{equation}
-where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
+
+\noindent where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance.
 We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers.
@@
 \end{equation}
-where $ \epsilon $ is a noise term.
+
+\noindent where $ \epsilon $ is a noise term.
 The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%            APPENDIX D : RESILIENCE ALGORITHM
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Resilience algorithm}
 \label{appendix_resilience}
+
 The following algorithm (see algorithm \ref{alg:algo4}) explains the process through which the resilience of a coalition structure is explored.
@@
 In this case, the coalition fails, and is removed from the market.
+
 \begin{algorithm}
@@
 \end{algorithm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%%            APPENDIX E : UTILITY FUNCTION
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 Ideally, we seek to express $ \alpha $ as a function of $ \bar{N} $ and the agents power distributions and correlations.
-\\
+
 In the general case, neither the power distributions nor the correlation structure have special forms which makes obtaining an analytical expression for $ \alpha $ difficult.
@@
 
-With :
+\noindent With :
 
@@
 
-Therefore, we can write the utility function as :
+\noindent Therefore, we can write the utility function as :
 
@@
 
-And the utility function becomes :
+\noindent And the utility function becomes :
 
@@
 
-$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $.
+\noindent $ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $.
 Depending on $ \alpha $, the utility function can have very different shapes.
-\\
-We select $\alpha^{\star} $ such that :
+
+\noindent We select $\alpha^{\star} $ such that :
 
@@
 
-Solving these equations yields the following expression :
+\noindent Solving these equations yields the following expression :
 
@@
 \bibliographystyle{IEEEtran}  
-\bibliography{Article}
+\bibliography{Appendix}
 \end{document}

### 2017-05-10 Article.tex
latest changes

@@
 \begin{document}
-%\pagestyle{plain}
-%\pagenumbering{arabic}
 
@@
 \author{Nicolas Gensollen, Vincent Gauthier, Monique Becker and Michel Marot
-%\IEEEauthorblockA{CNRS UMR 5157 SAMOVAR, \\
-%Telecom SudParis/Institut Mines Telecom\\
-%Email: \{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu}}
-\thanks{Authors are with CNRS UMR 5157 SAMOVAR, Telecom SudParis/Institut Mines Telecom, e-mail: (\{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu)}}%
+\thanks{Authors are with CNRS UMR 5157 SAMOVAR, Telecom SudParis/Institut Mines Telecom, e-mail: (\{nicolas.gensollen, vincent.gauthier, michel.marot, monique.becker\}@telecom-sudparis.eu)}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \begin{abstract}
@@
 {\bf Keywords:} prosumers, coalition formation, correlation, stability, clique expansion, electricity market
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:introduction}
-
-%The recent developments in the smart grid area, such as demand side management mechanisms, storage efficiency, or communication architectures to name a few, sketch a new power system.
-
-%Bidirectional power flows, storage, and distributed production are characteristics of smart grid systems.
-Nodes that were simply pure loads yesterday could behave tomorrow as generators or loads depending on weather conditions.
-
-%Bidirectional power flows, storage, and efficient use of information are some of the key principles that will enable this transition.
-Since production could be located down to the very end of the distribution networks, nodes that were simply pure loads yesterday could behave tomorrow as generators or loads, further complicating energy usages.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Related Work}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Generating realistic prosumer patterns}
@@
 	\begin{center}
-		\includegraphics[scale=.39]{Fig2}
+		\includegraphics[scale=.35]{figure_1.eps}
 		\caption{Process}
@@
 
-Where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
+\noindent where $ P_{i}^{P}(t) $ represents the total production of agent i at time t and $ P_{i}^{D}(t) $ its consumption at time t.
 In other words, $ P_{i}(t) $ represents the instantaneous surplus of power that agent i is willing to sell at time t.
@@
 \end{equation}
+
 \begin{equation}
@@
 We denote by $ \nu_{i}(t) $ and $ \Psi_{i}(t) $ the wind speed (in $ m.s^{-1} $) and the solar radiance ( in $ W.m^{-2} $) at the location of agent i and at time t, so that :
+
 \begin{equation}
@@
 \end{equation}
-Where $ \mathcal{F}_{WT} $ (resp.
+
+\noindent where $ \mathcal{F}_{WT} $ (resp.
 $ \mathcal{F}_{PV} $) is the power curve for the wind-turbines (resp.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Notations}
@@
 \end{equation}
-$ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
+
+\noindent $ \rho_{ij} $ being the Pearson's correlation coefficient between $ P_{i} $ and $ P_{j} $.
 If the coalition $ \{i,j\}$ proposes a contract value $ P_{S}^{CRCT} $, all instants when $ \{i,j\}$ will produce less than $ P_{S}^{CRCT} $ is critical.
@@
 A well-known result for normal distributions is that the cumulative distribution function can be written as :
+
 \begin{equation}
@@
 \end{equation}
-where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
+
+\noindent where $ erf $ is the error function : $ erf(x) = \dfrac{2}{\sqrt{\pi}}\int_{0}^{x} e^{-t^{2}} dt $.
 
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_1}
+		\includegraphics[scale=.46]{figure_2.eps}
 		\caption{{\footnotesize $ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for Gaussian distributions (see equation \ref{eq:P_star}).
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Utility Function}
@@
 \end{equation}
-(since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
+
+\noindent (since $ P_{S}^{CRCT \star} $ is supposed to be a constant rate over the contracted period).
 Unfortunately $ gain(S) $ is not a concave function of the coalitions' sizes, meaning that coalitions can grow as large as the number of agents allows it, without any counterbalance effects.
@@
 \end{equation}
-where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
+
+\noindent where parameter $ \alpha $ controls to what extent the size of a coalition impacts its utility, and $ P^{MAX} $ is a normalizing factor.
 Based on $ \mathcal{U} $, the marginal contribution of an agent i can be expressed as $ \delta_{S}(i) = \mathcal{U}(S+\{i\}) - \mathcal{U}(S) $.
@@
 
-Which leads to (see Appendix) :
+\noindent Which leads to (see Appendix) :
 
@@
 	\begin{center}
-		\includegraphics[scale=.46]{./figs/figure_2}
+		\includegraphics[scale=.46]{figure_3.eps}
 		\caption{{\footnotesize Gaussian mean approximation ($N=100,\ \phi=0.1,\ \bar{\mu}=5,\ \bar{\sigma}=.5,\ \bar{\rho}=0.2 $).
@@
 \label{sec:forming}
+
 We aim at forming $ N_{COAL} $ coalitions $ \{ S_1;S_2;...;S_{N_{COAL}} \} $ such that the global utility $ U = \sum_{i = 1}^{N_{COAL}} \mathcal{U}(S_i) $ is maximized.
@@
 \subsection{Representing the correlation structure}
+
 As seen in section \ref{sec:notations}, the variance of the aggregated production impacts directly the contract values, and depends on the covariances between the agents productions.
@@
 \end{equation}
-where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
+
+\noindent where $ G_{2}^{\epsilon} $ is the de-correlation graph $ G_{2} $ filtered by $ \epsilon $, and $ \Theta_{k}(G) $ is the set of non overlapping cliques of size k in a given graph G.
 In other words we select $ \epsilon^{\star} $ as the smallest threshold possible such that the filtered de-correlation graph contains at least $ N_{COAL} $ non overlapping cliques of size k.
@@
 	\begin{center}
-		\includegraphics[scale=.45]{./figs/figure_3}
+		\includegraphics[scale=.45]{figure_4.eps}
 		\caption{{\footnotesize Histograms of utility values for coalitions of size 3 ($\bar{N}=3$) in a $N=200$ prosumers example ($\phi = 0.1$, $ \alpha^{\star} = 0.08 $, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 \subsection{Algorithm}
-The algorithm takes inputs from :
+
+\noindent The algorithm takes inputs from :
 
@@
 	\begin{center}
-		\includegraphics[scale=.45]{./figs/figure_4}
+		\includegraphics[scale=.45]{figure_5.eps}
 		\caption{{\footnotesize Utility of random coalitions depending on their size in a $N=200$ prosumers example ($\phi = .1$, $ \bar{\mu}=3.9\ MW $, $\bar{\sigma} = 1.9 MW $, $ \bar{\rho} = 0.69 $).
@@
 	\begin{center}
-		\includegraphics[scale=.45]{./figs/figure_5}
+		\includegraphics[scale=.45]{figure_6.eps}
 		\caption{{\footnotesize Coalitions formed in the (contract value, volatility) space.
@@
 	\begin{center}
-		\includegraphics[scale=.45]{./figs/figure_6}
+		\includegraphics[scale=.45]{figure_7.eps}
 		\caption{{\footnotesize Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot)} }

### 2017-05-10 biography.tex
latest changes

\documentclass[journal]{IEEEtran}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Package
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ifpdf}
\ifCLASSINFOpdf
  \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[french, english]{babel}
\usepackage{cite}
\usepackage{caption}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{subcaption}
\usepackage{array}
\usepackage{color}
\usepackage{float}
\usepackage{multicol}
\usepackage[]{algorithm2e}
\usepackage{hyperref}
\usepackage{fancyhdr}
%
% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Nicolas_Gensollen.jpg}}]%
{Gensollen Nicolas}
was born in Paris in 1989. He received the telecommunication engineer degree from Telecom SudParis in 2012 with a specialization in networking. He is currently a PhD student at the Institut Mines-T\'el\'ecom, T\'el\'ecom SudParis, France. His current research interests are mainly on smart grid, complex systems, and agent based modeling. Other research interests include dynamics and control in networks and game theory. 
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Bio_Vincent.eps}}]%
{Gauthier Vincent}
was born in Paris in 1978. He received the B.S. in electrical engineering from University de Bretagne Occidentale in 2002 and M.S. degrees the Ph.D. degree in computer science and computer networks from University of Paris 6 in 2003 and 2006, respectively. He was a Guest Researcher at National Institute of Standards and Technology, MA, United States between 2006 and 2008. He joined the faculty of Telecom SudParis, and the lab CNRS SAMOVAR (UMR 5157), Evry, France, in 2008, where he is currently an Associate Professor of the Department of Computer Network. His current research interests are primarily on computer networks, self-organization and complex systems. His other research interests include mobility modeling, performance analysis, and queuing theory, sensor networks, ad-hoc networks, cross-layer design.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{Monique_Becker.jpg}}]%
{Becker Monique}
who studied Maths at ENSJF was a researcher at CNRS in Paris VI University from 1968 to 1987, and then a Professor at Telecom SudParis, till 2015. She is now an Emeritus Professor in Institut Mines-Telecom.
Her PhD concerned “validity of simulations” and her research work concerns mostly Performance Evaluation and Modelling of Networks and Distributed Systems. She designed aggregation methods for performance evaluation. She wrote a book, with her PhD students on Network simulations. 
She was the Director of the CNRS UMR5157: Samovar, since its starting and during 10 years.
She was the director of about 50 PhD students.
She started in 2001 a Master course taught in English: “Computer and Communication Networks” in Telecom SudParis. This Master is now taught in Paris Saclay University.
She was an expert for NSF, for NSERC (Canada) and for the European Commission.
She mostly collaborates with research groups in USA, in Austria, in Singapore and in India.

\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{michel_marot.jpg}}]%
{Marot Michel}
was born in Paris in 1973. He received the Ph.D. and habilitation degrees in computer networks from University of Paris 6 in 2001 and 2010. He is a professor in the Telecommunication Networks and services department at the Institut Mines-T\'el\'ecom, T\'el\'ecom SudParis, France. His current research interests are mainly on network performance and self-organization in wireless networks, ad-hoc and sensor networks, vehicular networks, context aware adaptation. His other research interests include mobility modeling, complex systems, and queuing theory and more generally performance of distributed systems.
\end{IEEEbiography}

\end{document}
